---
title: "TABR 2024 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to look at each logger individually to verify that the data is clean.

Then I will plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read_excel("24TABR_ibuttons.xlsx", 
    range = "A1:J51", na = "NA") |> 
  mutate(Trim_Date = ymd(Trim_Date)) |> 
  mutate(Removal_Date = ymd(Removal_Date))

glimpse(Metadata)

# Logger_Assignments <- read_excel("TemperatureLoggers/StarOddi_Temp_Logger_Assignments_Epla_2023.xlsx",                                 range = "A1:E74", col_types = c("text", "text", "text", "date", "text")) # not needed because logger number already in metadata
# Logger_Assignments <- Logger_Assignments %>%
#  mutate(DMR_ID = paste0("DMR", `DMR Number`)) %>% 
#  relocate(DMR_ID) # takes DMR Number and adds DMR to the beginning in a new column
# write_tsv(Logger_Assignments, 
        #  file = "TemperatureLoggers/StarOddi_Temp_Logger_Assignments_Epla_2023_updated.tsv")
# Logger_Assignments$DMR_ID %in% Epla2023MuniDataset$DMR_ID
# Epla2023MuniDataset$DMR_ID %in% Logger_Assignments$DMR_ID
# Metadata <- full_join(Logger_Assignments, Epla2023MuniDataset, by = "DMR_ID")
# Logger_Assignments %>%
#  filter(!DMR_ID %in% Epla2023MuniDataset$DMR_ID)
```

I do not need most of the code chunk above because I already created a metadata csv containing logger number, bat ID, and bat location.

```{r}
DATA_DIR <- "./BUTT Data/Identified" # load in each logger file

file_list <- as_tibble(list.files(DATA_DIR)) |> 
  mutate(Bat_ID = gsub("_.*", "", value)) |> # removes everything after bat ID
  rename(Filename = value) 

Metadata <- left_join(Metadata, file_list, by = join_by(`Bat ID` == Bat_ID)) # adds file name into metadata

batfilelist <- list.files(path = "./BUTT Data/Identified", 
                          pattern = "*.csv", full.names = T) # shows full file path name
batfilenames <- list.files(path = "./BUTT Data/Identified", 
                           pattern = "*.csv", full.names = F) # shows only file name
batfiles <- lapply(batfilelist, fread, 
                   col.names = c("DateTime","Unit", "Temperature"), sep = "," , 
                   skip = 20) # applies columns to each file, skips the first 20 rows

batnames <- gsub("_.*", "", batfilenames) # leaves only bat ID

names(batfiles) <- batnames # names each file by logger name
allbatresults <- bind_rows(batfiles, .id = "batnames") # creates a long dataset with each time point for every logger, 4 variables 

allbatresults <- allbatresults %>%
  mutate(DateTime = mdy_hms(DateTime)) # changes date to appropriate format
str(allbatresults) # similar to glimpse
summary(as.factor(allbatresults$batnames)) # display all bat names and their counts use to verify they ran correctly
md.pattern(allbatresults) # checking for NA's
Metadata %>%
  filter(!`Bat ID` %in% file_list$Bat_ID) %>% # only selects loggers with a matching file name
  drop_na(`Filename`) # removes all na's
Metadata %>%
  filter(is.na(`Filename`)) # shows which bats do not have a logger
```

The bats with no files no files represent loggers that failed to download, 

The logger numbers with no bat ID have not been identified

## Missing Data

```{r}
clean_met <- Metadata %>%
  drop_na(Filename) # removes any ibuttons with no file

head(clean_met) # look at the top of the dataset
md.pattern(clean_met, rotate.names = TRUE)
write_tsv(clean_met, file = "Tabr2024TempLoggerMetadata.tsv")
```

I am left with 21 identified ibuttons with files. 

## Making a plot function

```{r}
plot_fun <- function(arg1) {
  allbatresults %>%
    dplyr::filter(batnames == arg1) %>%
    ggplot(aes(x=DateTime, y=Temperature)) +
    geom_line() +
    ylab("Temperature (C)") + xlab("Date") + ylim(4,40) +
    theme_cowplot() 
}
```

```{r}
plotly_fun <- function(arg1) {
  plot_ly(data = allbatresults[batnames == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

## Looking at the data

```{r Bat MST60063}
ggplotly(plot_fun("MST60063"))
```

```{r Plotly MST60063}
plotly_fun("MST60063")
```

Here we go!

This works fine, but is not interactive: lapply(batnames2023, plot_fun)

```{r}
lapply(batnames, plotly_fun)
```

If you can't see the whole plot, use the "Show in New Window" icon above the output.


Downstairs boost was on Nov 18, 202
Upstairs boost was on 20 Nov 2024

MST60069 logger fell off 2 days (over 48 hours) after boost
MST60094, MST60101, logger fell off over 24 hours but less than 48 hours after boost
I think as long as they were on for the 24 hours after the boost we should keep the data

```{r Plotly MST60069}
plotly_fun("MST60069")
```


```{r Plotly MST60094}
plotly_fun("MST60094")
```

```{r Plotly MST60101}
plotly_fun("MST60101")
```

Additional things to consider:
MST60066 recorded temps over 70C during the warm bath cal. - temp data on the bat looks sound though
MST60085, MST60104, MST60157, and MST60158 recorded temps around 65C during the warm bath cal. - temp data on bat looks sound

No identified loggers fell off at any point that I can tell from the data. 

## Left Trimming the data

Removing all data before midnight on the day of the logger attachment. 

Now to trim the datalogger data.
Loggers were attached at least 4 days before the booster injection.
Note that there are two attachment dates, 2024-11-14 for downstairs and 2024-11-15 for upstairs.
I will split the data, filter it, and then recombine them.

```{r}
summary(as.factor(clean_met$`Attachment Date`))
first_attachment <- clean_met %>%
  filter(`Attachment Date` == ymd("2024-11-14"))
second_attachment <- clean_met %>%
  filter(`Attachment Date` == ymd("2024-11-15"))
first_results <- allbatresults %>%
  filter(batnames %in% first_attachment$'Bat ID')
second_results <- allbatresults %>%
  filter(batnames %in% second_attachment$'Bat ID')
first_results_trim <- first_results %>%
  filter(DateTime >= ymd("2024-11-14"))
second_results_trim <- second_results %>%
  filter(DateTime >= ymd("2024-11-15"))
left_trimmed_results <- bind_rows(first_results_trim, second_results_trim)
```


```{r}
lapply(batnames, plotly_fun)
```
## Removing bad data

No bad data to remove right now
```{r}
# left_trimmed_results <- left_trimmed_results %>%
  # filter(!batnames %in% c("1N8482", "1N8519", "1N8524", "2N8247"))
```

## Right trimming the data

I do not have the date removed in the metadata how the EPLA data set up, so instead I want to filter all results for bat IDs in the first attachment and then trim it for November 25th. This was the day loggers were removed from downstairs bats. 
I will do the same for upstairs bats for November 27th. 

After that, I will look at which loggers fell off before then and trim the data more for each logger.

For both kinds of right trimming, I will remove all data for the partial day and the datastream will stop at the previous midnight.
```{r}
summary(as.factor(clean_met$`Attachment Date`))



first_results_trimmed <- first_results_trim %>%
  filter(DateTime <= ymd("2024-11-24"))
second_results_trimmed <- second_results_trim %>%
  filter(DateTime <= ymd("2024-11-26"))
allbatresults2 <- bind_rows(first_results_trimmed, second_results_trimmed)

summary(as.factor(allbatresults2$batnames))
#TempLoggerMetadata <- read_delim("Tabr2024TempLoggerMetadata.tsv", 
#    delim = "\t", escape_double = FALSE, 
 #  trim_ws = TRUE)
#str(TempLoggerMetadata)
#head(TempLoggerMetadata)
#TempLoggerMetadata$Id %in% left_trimmed_results$batnames
#TempLoggerMetadata$Id[!TempLoggerMetadata$Id %in% left_trimmed_results$batnames]
#unique(left_trimmed_results$batnames) %in% TempLoggerMetadata$Id
#Merged_results <- left_join(left_trimmed_results, TempLoggerMetadata, by = join_by(batnames == Id)) %>%
 # arrange(batnames, Timepoint)
#Merged_trimmed_results <- Merged_results %>%
#  filter(DateTime < RIP_date)
#allbatresults2 <- Merged_trimmed_results %>%
#  select(batnames, Timepoint, DateTime, Temperature)
cleanbatnames2024 <- unique(allbatresults2$batnames)
```

Making a new plotly function for these results. 
Yes, I should try to make the whole dataframe one of the arguments for the function, but I will save that for later.

```{r}
plotly_fun2 <- function(arg1) {
  plot_ly(data = allbatresults2[batnames == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

I went through every plot generated below and recorded "Logger_fail_date" which was the date that
either the logger appeared to fall off, the bat died, or (for one logger) the logger started recording > 50C.

```{r}
lapply(cleanbatnames2024, plotly_fun2)
```

Notes for each logger:
MST60063 -- fell off 16:45 on Nov 21st
MST60069 -- fell off 23:15 on Nov 20th
MST60076 -- stayed on duration of project
MST60085 -- fell off 10:03 on Nov 23rd
MST60101 -- Fell off 02:32 on Nov 20th
MST60110 -- stayed on duration of project
MST60136 -- stayed on duration of project
MST60142 -- stayed on duration of project
MST60149 -- stayed on duration of project
MST60157 -- stayed on duration of project
MST60066 -- stayed on duration of project; shows very clear deep torpor within 2-3 hours of lights off getting into the mid to low 20s - cGAMP bat
MST60070 -- stayed on duration of project; going into deep torpor 3 days after injection - LPS bat
MST60075 -- stayed on duration of project; going into deep torpor 3 dats after injection - cGAMP bat
MST60089 -- fell off 17:57 on Nov 24th
MST60090 -- stayed on duration of project
MST60094 -- fell off 10:48 on Nov 22nd
MST60096 -- stayed on duration of project; torpor deeper for next few days after injection
MST60104 -- stayed on duration of project; shallower fluctuation in body temperature after injection 
MST60126 -- stayed on duration of project; lower average body temp after injection?
MST60139 -- stayed on duration of project
MST60158 -- stayed on duration of project; shallower fluctuation in body temp for few days after injection



Double checking the ones that were missing: Ask ken about this - not sure what was missing
```{r}
#plotly_fun2("1N8482")
#plotly_fun("1N8482")
#plotly_fun2("1N8519")
#plotly_fun("1N8519")
#plotly_fun2("1N8524")
#plotly_fun("1N8524")
```

Okay, that is all as expected.
Time to retrim that data.

Added the dates each logger fell off or was removed in the excel file. 

```{r}
# Moves to top NewTempLoggerMetadata <- read_excel("24TABR_ibuttons.xlsx") %>%
 # mutate(Trim_Date = ymd(Trim_Date)) |> 
  # mutate(Removal_Date = ymd(Removal_Date))

# str(NewTempLoggerMetadata)
Clean_results <- left_join(allbatresults2, clean_met, by = join_by(batnames == 'Bat ID')) %>%
  arrange(batnames, DateTime)
Clean_trimmed_results <- Clean_results %>%
  filter(DateTime < Removal_Date)
```


Before I move on and separate the results from the metadata, I want to make sure D21s are lining up correctly. To do this, I am going to timeshift Group2 by 2 days. Run this by Ken to determine if this was the best course of action.
```{r Timesift}
# Clean_trimmed_results$time_aligned <- Clean_trimmed_results$DateTime

# Clean_trimmed_results$time_aligned[Clean_trimmed_results$Location == "Upstairs"] <- Clean_trimmed_results$time_aligned[Clean_trimmed_results$Location == "Upstairs"] - lubridate::days(2)

```

Okay, I am about to go very off track here.. I want to convert date into day of experiment because date does not matter to us. I also want days to start at 11AM since that is when the lights go off. I used chatGPT for the following code chunk.. we'll see what Ken thinks about it. 

```{r}
Clean_trimmed_results <- Clean_trimmed_results |> 
  group_by(batnames) |> 
  mutate(Logger_start = min(DateTime)) |> 
  ungroup()

Clean_trimmed_results <- Clean_trimmed_results |> 
  mutate(Logger_11AM = floor_date(Logger_start, unit = "days") + hours(11),
         Logger_11AM = if_else(Logger_start > Logger_11AM,
                               Logger_11AM + days(1), Logger_11AM))

Cleaner_results <- Clean_trimmed_results |> 
  filter(DateTime >= Logger_11AM) # Removes data from before 11Am on attachment day

Cleaner_results <- Cleaner_results %>%
  mutate(
    Minutes_Since_11AM = as.numeric(difftime(DateTime, Logger_11AM, units = "mins")),
    Days_Since_11AM = Minutes_Since_11AM / (60 * 24))

Cleaner_results <- Cleaner_results %>%
  mutate(
    Logger_Experiment_Day = case_when(Location == "Downstairs" ~ 17,
                                      Location == "Upstairs" ~ 16),
    Experiment_Day = Logger_Experiment_Day + Days_Since_11AM,
    Days_From_D21 = Experiment_Day - 21)
```


```{r}
allbatresults3 <- Cleaner_results %>%
  select(batnames, Experiment_Day, Temperature) |> 
  rename(id = batnames)
cleanbatnames2024.1 <- unique(allbatresults3$id)
summary(as.factor(allbatresults3$id))
```

```{r}
plotly_fun3 <- function(arg1) {
  plot_ly(data = allbatresults3 |>  filter(id == arg1),
          x = ~Experiment_Day, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

```{r}
lapply(cleanbatnames2024.1, plotly_fun3)
```

Notes (were made before switching to experimental day format)
cGAMP:
  Downstairs:
  MST60063 -- Slightly higher body temperature after injection - higher maximum temps and no drastic drops in body temp - body temp between 32 and 34
  MST60101 -- only 1 day of data after injection; nothing noticeable - not much variation in body temp compared to others
  MST60110 -- no noticeable trends - not much variation in body temp compared to others
  MST60136 -- higher averages after injection?
  Upstairs:
  MST60066 -- shows very clear deep torpor within 2-3 hours of lights off getting into the mid to low 20s; average temps look a bit lower after injection
  MST60075 -- going into deep torpor 3 days after injection body temp dropping to 22
  MST60126 -- Body temp lower the last 5 days after injection

LPS:
  Downstairs: 
  MST60157 -- Body temp lower for the few hours after injection; higher average body temp 24-96 hours after the injection - looks like less fluctuation in body temp after injection
  Upstairs:   
  MST60070 -- going into deep torpor 3 days after injection body temp dropping to 22
  MST60090 -- extended torpor 3 days after injection 2:30 to 10 am - maybe lower average temp after injection?
  

Poly I:C:
  Downstairs:
  MST60069: No drop in body temp before lights off 1 day after injection maybe not going into torpor - not much variation compared to others
  MST60085 -- No noticeable trends
  Upstairs:
  MST60089 -- No noticeable trends
  MST60158 -- body temperature not dropping as low the two days post injection - reaching 28 and 29.5 lows compared to 25 - large drop in body temp 4 days after injection reaching 21.5 - note that body temp was also not very low the day of the injection before handling only reached 28  - check ambient temps

eVLP 
  Downstairs:
  MST60096 -- lower low body temp (maybe torpor deeper?) for next 5 days after injection
  MST60142 -- 36.5 at 16:21 on injection day and 36 at 17:01 on D22 (could be handled at that time or in bag?); maybe more inconsistent after injection
  MST60149 -- no noticeable trends
  Upstairs:
  MST60104 -- Torpor somewhat more shallow for two days after injection - usually drops to around 24-26 around 10 am but was around 28.5-30.5 at the same time the two days after injection
  MST60139 -- no noticeable trends

Sham:
  Downstairs:
  MST60076 -- Large drop in body temp morning after injection - utilized torpor more after injection? - not much variation compared to others
  Upstairs: 
  MST60094 -- No noticeable trends - only 1 days worth of data after injection
  


Saving the trimmed data!

Look at MST600126- fell off right before midnight will this drop effect results? Should I trim the last two recordings?


```{r}
# write_tsv(allbatresults3, file = "TrimmedTemperatureLoggerData.long.tsv")

write_tsv(allbatresults3, file = "TempDataDays.long.tsv")

```


## Behavr Analysis

I am going to move ahead with the timepoints as before, 
but I need to verify that the data is lining up properly.

Ask Ken - Does this need to be done for my datasets? No timepoints recorded 

This code chunk is not needed if we are keeping the above time shift and experimental days

```{r}
# select_bat_results <- allbatresults3 %>%
#  mutate(Timepoint = Timepoint*10*60) %>%
 # select(id = batnames, t = Timepoint, Temperature) 
# head(select_bat_results)

#  <- clean_met %>%
#  rename(id = Id) %>%
 # filter(id %in% cleanbatnames2023)
  
# bat_data <- clean_met3 %>%
#  select(id) %>%
 # left_join(select_bat_results, by = join_by(id))
# str(bat_data)

# bat_met2 <- data.table::data.table(clean_met3, key = "id")
# bat_data2 <- data.table::data.table(bat_data, key = "id")

# bat_behavr <- behavr(bat_data2, bat_met2)
# summary(bat_behavr)

## Ashlyn Attempt Below
  # Need to change t into a number or timepoint?
  
# str(allbatresults3)  

#select_bat_results <- allbatresults3 %>%
 # select(id = batnames, t = DateTime, Temperature) 
  #head(select_bat_results)

#  clean_met3 <- clean_met %>%
#    rename(id = 'Bat ID') %>%
 #  filter(id %in% cleanbatnames2024.1)
  
#  bat_data <- clean_met3 %>%
 #   select(id) %>%
  #  left_join(select_bat_results, by = join_by(id))
  #str(bat_data)

#  bat_data$t <- as.numeric(bat_data$t)

 # str(bat_data)

#code to line up all times on 0
 # bat_data |> group_by(id) |> 
  #  mutate(Min_time = min(t),
        #   Time = as.numeric(t - Min_time)) |> 
  #  ungroup() -> shifted_data
        
# code to delete all times starting after 11:00 am - did up above with experimental days.



 # bat_met2 <- data.table::data.table(clean_met3, key = "id")
 # bat_data2 <- data.table::data.table(shifted_data, key = "id")

 # bat_behavr <- behavr(bat_data2, bat_met2)
 # summary(bat_behavr)
 # str(bat_behavr)
```

```{r}
  clean_met3 <- clean_met %>%
    rename(id = 'Bat ID') %>%
   filter(batnames %in% cleanbatnames2024.1) 
bat_met2 <- data.table::data.table(clean_met3, key = "id")
  bat_data2 <- data.table::data.table(allbatresults3, key = "id")

  bat_behavr <- behavr(bat_data2, bat_met2)
  summary(bat_behavr)
  str(bat_behavr)

 # bat_behavr[, t:= Experiment_Day]  # only use if using experiment day
```


```{r}
# ggetho(bat_behavr, aes(x=Time, y=id, z=Temperature)) + stat_tile_etho() # use with commented out code chunk 

# ggetho(bat_behavr, aes(x= t, y=id, z=Temperature)) + stat_tile_etho() # use with experiment day - not working though

```

```{r}
ggetho(bat_behavr, aes(x=Time, z = Temperature, y = Group)) + stat_tile_etho()
```

```{r, warning=FALSE}
summary(bat_behavr)
bat_behavr %>%
  ggetho(aes(x = Time, y=Temperature, colour=Group)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(Group ~ .)
bat_behavr %>%
  ggetho(aes(x = Time, y=Temperature)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(Group ~ .)
```

```{r, warning=FALSE}
#ggetho(bat_behavr, aes(x = Time, y=Temperature, colour=Treatment), time_wrap = hours(24)) +
 #     stat_pop_etho() +
  #    stat_ld_annotations() +
   #   facet_grid(Treatment ~ .) +
    #  scale_y_continuous(name= "Temperature")
```

```{r Sham, warning=FALSE}
bat_met_sham <- bat_met2 %>%
  filter(Treatment == "SHAM")
bat_data_sham <- bat_data2 %>%
  filter(id %in% bat_met_sham$id)
bat_behavr_sham <- behavr(bat_data_sham, bat_met_sham)

bat_behavr_sham %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP, warning=FALSE}
bat_met_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP")
bat_data_eVLP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP$id)
bat_behavr_eVLP <- behavr(bat_data_eVLP, bat_met_eVLP)

bat_behavr_eVLP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_LPS, warning=FALSE}
bat_met_eVLP_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS")
bat_data_eVLP_LPS <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_LPS$id)
bat_behavr_eVLP_LPS <- behavr(bat_data_eVLP_LPS, bat_met_eVLP_LPS)

bat_behavr_eVLP_LPS %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_POLY, warning=FALSE}
bat_met_eVLP_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY")
bat_data_eVLP_POLY <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_POLY$id)
bat_behavr_eVLP_POLY <- behavr(bat_data_eVLP_POLY, bat_met_eVLP_POLY)

bat_behavr_eVLP_POLY %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_cGAMP, warning=FALSE}
bat_met_eVLP_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP")
bat_data_eVLP_cGAMP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_cGAMP$id)
bat_behavr_eVLP_cGAMP <- behavr(bat_data_eVLP_cGAMP, bat_met_eVLP_cGAMP)

bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r, warning=FALSE}
ggetho(bat_behavr, aes(y=Temperature, colour=Treatment)) + 
  stat_pop_etho()
```

```{r, warning=FALSE}
bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(x=t, z=Temperature), multiplot = 2) +
      stat_tile_etho() +
      facet_wrap( ~ id)
```

## Subsetting the data to pre-boost and post-boost

Before I proceed with subsetting, I am going to make sure that I understand the data.

```{r}
allbatresults3 %>%
  group_by(batnames) %>%
  summarize(Min_timepoint = min(Timepoint), Min_DateTime = min(DateTime), Max_DateTime = max(DateTime),
            Min_Temp = min(Temperature), Max_Temp = max(Temperature))
```

We can tell that the loggers on 2023-04-16 were all programmed uniformly and collected 42 timepoints before the new start time of midnight on 2023-04-16. 
While the 2023-04-16 loggers have two different programming regimens. Some collected 186 and others collected 330 earlier datapoints. 

Bottom line, there are three different starting timepoints: 43, 187, and 331.
I will substract the appropriate value from the Timepoint column after I separate the data.


```{r}
left_trimmed_results %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint))
allbatresults %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint), maxTimepoint = max(Timepoint), 
            minTemp = min(Temperature), maxTemp = max(Temperature)) %>%
  arrange(maxTimepoint)
```


There are three different batches of treatment, as expected:
2023-03-27 2023-03-29 2023-03-31 were the initial immunization dates (prime)
Day 21 for each batch was, therefore:
2023-04-17 2023-04-19 2023-04-21

```{r}
summary(as.factor(bat_met2$Treatment_date))
first_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-27"))
second_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-29"))
third_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-31"))

first_results <- allbatresults3 %>%
  filter(batnames %in% first_treatment$id)
second_results <- allbatresults3 %>%
  filter(batnames %in% second_treatment$id)
third_results <- allbatresults3 %>%
  filter(batnames %in% third_treatment$id)

first_results_trim <- first_results %>%
  filter(DateTime <= ymd("2023-04-17")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_trim <- second_results %>%
  filter(DateTime <= ymd("2023-04-19")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_trim <- third_results %>%
  filter(DateTime <= ymd("2023-04-21")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

pretreatment_results <- bind_rows(first_results_trim, second_results_trim) %>%
  bind_rows(third_results_trim)

select_bat_results <- pretreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(select_bat_results)
  
bat_data_preboost <- clean_met3 %>%
  select(id) %>%
  left_join(select_bat_results, by = join_by(id))
str(bat_data_preboost)

bat_data_preboost2 <- data.table::data.table(bat_data_preboost, key = "id")

bat_behavr_preboost <- behavr(bat_data_preboost2, bat_met2)
```

Note that I think that the last day shown on these plots is the booster immunization day.

```{r, warning=FALSE}
# summary(bat_behavr_preboost)
# bat_behavr_preboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_preboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

And now for the post-boost results. 
For starters, I will trim the 4 days after the injections.
That didn't work because ggetho kept saying that there was less than 3 days of data :/

Secondly just kept the above format but used more data.

```{r}
# first_results_post <- first_results %>%
#   filter(DateTime >= ymd("2023-04-17")) %>%
#   filter(DateTime <= ymd("2023-04-22")) %>%
#   mutate(Timepoint = Timepoint - 474) ## My attempt to standardize the start time. 
# second_results_post <- second_results %>%
#   filter(DateTime >= ymd("2023-04-19")) %>%
#   filter(DateTime <= ymd("2023-04-24")) %>%
#   mutate(Timepoint = Timepoint - 762)
# third_results_post <- third_results %>%
#   filter(DateTime >= ymd("2023-04-21")) %>%
#   filter(DateTime <= ymd("2023-04-26")) %>%
#   mutate(Timepoint = Timepoint - 906)

first_results_post <- first_results %>%
  filter(DateTime <= ymd("2023-04-22")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_post <- second_results %>%
  filter(DateTime <= ymd("2023-04-24")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_post <- third_results %>%
  filter(DateTime <= ymd("2023-04-26")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

posttreatment_results <- bind_rows(first_results_post, second_results_post) %>%
  bind_rows(third_results_post)

post_bat_results <- posttreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(post_bat_results)
  
bat_data_postboost <- clean_met3 %>%
  select(id) %>%
  left_join(post_bat_results, by = join_by(id))
str(bat_data_postboost)

bat_data_postboost2 <- data.table::data.table(bat_data_postboost, key = "id")

bat_behavr_postboost <- behavr(bat_data_postboost2, bat_met2)
summary(bat_behavr_postboost)
```
```{r Postboost_Summary, warning=FALSE}
# summary(bat_behavr_postboost)
# bat_behavr_postboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_postboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

So I think that the booster happened on day 4 on these graphs, but I need to verify.

## Booster Analysis

To be certain of using the correct timestamps, I am going to analyze with the openair package, as I did for the squirrels.

This is done one bat at a time...

https://bookdown.org/david_carslaw/openair/sections/trend-analysis/time-variation.html

First, comparing all pre-boost data to (boost day + 1 day following)

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

The results look very similar.


```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

To keep the comparison cleaner, I will continue with pre versus post. 
The only downside is that the confidence interval is a little wider on the "pre" data than the "other" data.

## Comparison of all bats

```{r}
bat_met2 %>%
  arrange(Treatment) %>%
  select(id, Treatment)
```

Let's make a function for the timeVariation plot

```{r}
timeVariation_fun <- function(arg1) {
  mybatID <- arg1

  
  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(2)

  singlebatresults <- allbatresults3 %>%
    filter(batnames == mybatID)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", 
                           group = "Boosted",  
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

```{r}
timeVariation_fun("1N8476")
```

Now let's look at the groups one at a time.

```{r SHAM}
bat_sham <- bat_met2 %>%
  filter(Treatment == "SHAM") %>%
  select(id) %>%
  pull
lapply(bat_sham, timeVariation_fun)
```

```{r eVLP only}
bat_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP") %>%
  select(id) %>%
  pull
lapply(bat_eVLP, timeVariation_fun)
```


```{r eVLP LPS}
bat_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS") %>%
  select(id) %>%
  pull
lapply(bat_LPS, timeVariation_fun)
```

```{r eVLP POLY}
bat_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY") %>%
  select(id) %>%
  pull
lapply(bat_POLY, timeVariation_fun)
```

```{r eVLP cGAMP}
bat_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP") %>%
  select(id) %>%
  pull
lapply(bat_cGAMP, timeVariation_fun)
```

## Tent Logger Data

Two possible ways to incorporate this data. 
First, I will add it as a separate "pollutant".
Second, I will show it as a separate "feature".

```{r}
alltentresults <- fread("TemperatureLoggers/tentloggerdata_long.tsv")
alltentresults <- alltentresults %>%
  mutate(Tent = gsub("Tent ", "", Tent))
str(alltentresults)

tent_met <- Metadata %>%
  drop_na(Filename) %>%
  select(Filename, DMR_ID, Tag, Treatment, Age, Sex,
         Attachment_date = "Date Attached", 
         Treatment_date = Day_00, 
         RIP_date,
         Tent = Tent.x) %>%
  mutate(id = gsub(".DAT", "", Filename)) %>%
  relocate(id)

tentbatresults <- tent_met %>%
  select(id, Tent) %>%
  right_join(allbatresults3, by = join_by(id == batnames))

tentbatresults2 <- tentbatresults %>%
  left_join(alltentresults, by = join_by(Tent, DateTime)) %>%
  rename(`Bat T` = Temperature.x) %>%
  rename(`Tent T` = Temperature.y)

write_csv(tentbatresults2, file = "allbatandtentresults.csv")
```

A new plot function:

```{r}
timeVariation_fun2 <- function(arg1) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  # Immunization <- bat_met2 %>%
  #   filter(id == mybatID) %>%
  #   select(Treatment_date) %>%
  #   pull
  # Boost <- Immunization + days(21)
  # BoostWindow <- days(2)

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID)

  # singlebatresults2 <- singlebatresults %>%
  #   filter(DateTime <= Boost + BoostWindow) %>%
  #   mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  #   mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  #   rename(date = DateTime)
  
  singlebatresults2 <- singlebatresults %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant =c("Tent T", "Bat T"), 
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun2("1N8476")
```
That looks good. Let's plot each group at a time.

```{r Sham-Tent}
lapply(bat_sham, timeVariation_fun2)
```

```{r eVLP-Tent}
lapply(bat_eVLP, timeVariation_fun2)
```

```{r LPS-Tent}
lapply(bat_LPS, timeVariation_fun2)
```

```{r POLY-Tent}
lapply(bat_POLY, timeVariation_fun2)
```

```{r cGAMP-Tent}
lapply(bat_cGAMP, timeVariation_fun2)
```

Okay, because the "difference" function worked fine, I don't think that I need to reorganize the data. 
Let's see if I can also use the boost "group" at the same time.

Dammit: "Can only have one pollutant with a grouping variable, or several pollutants and no grouping variable."

Okay, I think I will make a new variable called delta T and use that as the only pollutant.

I will also test different "postboost" windows.

A new plot function:

```{r}
timeVariation_fun3 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant = "delta T", 
                           group = "Boosted",
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun3("1N8476", 5)
```

```{r Sham-deltaT}
lapply(bat_sham, timeVariation_fun3, 3)
```

```{r eVLP-deltaT}
lapply(bat_eVLP, timeVariation_fun3, 3)
```

```{r LPS-deltaT}
lapply(bat_LPS, timeVariation_fun3, 3)
```

```{r POLY-deltaT}
lapply(bat_POLY, timeVariation_fun3, 3)
```

```{r cGAMP-deltaT}
lapply(bat_cGAMP, timeVariation_fun3, 3)
```

## Comparing by Treatment Group

Using the "group" parameter in timeVariation and to make separate plots for pre-boost and then each day post-boost.


```{r Group Comparisons}
tentbatresults3 <- tentbatresults2 %>%
  mutate(`delta T` = `Bat T` - `Tent T`) %>%
  rename(date = DateTime)

bat_boost_date <- bat_met2 %>%
  select(id, Treatment, Treatment_date) %>%
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP",
                                                  "eVLP LPS", "eVLP POLY", "eVLP cGAMP")))

tentbatresults4 <- tentbatresults3 %>%
  left_join(bat_boost_date, by = join_by(id))

str(tentbatresults4)

preboost_tentbatresults <- tentbatresults4 %>%
  filter(date < (Treatment_date + days(21)))

boostday_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(21)) & date <= (Treatment_date + days(22)))

postboostday1_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(22)) & date <= (Treatment_date + days(23)))

postboostday2_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(23)) & date <= (Treatment_date + days(24)))

postboostday3_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday4_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday5_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(25)) & date <= (Treatment_date + days(26)))


preboost_plot <- timeVariation(preboost_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Preboost", "Preboost", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(preboost_plot, subset="hour")

boostday_plot <- timeVariation(boostday_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Boost Day", "Boost Day", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(boostday_plot, subset="hour")

postboostday1_plot <- timeVariation(postboostday1_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 1", "Postboost Day 1", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday1_plot, subset="hour")

postboostday2_plot <- timeVariation(postboostday2_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 2", "Postboost Day 2", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday2_plot, subset="hour")

postboostday3_plot <- timeVariation(postboostday3_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 3", "Postboost Day 3", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday3_plot, subset="hour")

postboostday4_plot <- timeVariation(postboostday4_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 4", "Postboost Day 4", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday4_plot, subset="hour")

postboostday5_plot <- timeVariation(postboostday5_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 5", "Postboost Day 5", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday5_plot, subset="hour")

```


## Quantifying Temperature Differences

Designing a function to measure the daily difference in Bat temperature compared to the tent temperature.

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
```


A function to quantify this delta T by day:

```{r}
tempDifference_fun <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  DMR_Id <- NewTempLoggerMetadata %>%
    filter(Id == mybatID) %>%
    select(DMR_ID) %>%
    pull
  
  Tent <- Metadata %>%
    filter(DMR_ID == DMR_Id) %>%
    select(Tent.x) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`)/144, 
                      daily_batT = mean(`Bat T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 Tent = Tent,
                                 `DeltaT d-3` = daily_summary$daily_delta[1],
                                 `DeltaT d-2` = daily_summary$daily_delta[2],
                                 `DeltaT d-1` = daily_summary$daily_delta[3],
                                 `DeltaT d0` = daily_summary$daily_delta[4],
                                 `DeltaT d1` = daily_summary$daily_delta[5],
                                 `DeltaT d2` = daily_summary$daily_delta[6],
                                 `BatT d-3` = daily_summary$daily_batT[1],
                                 `BatT d-2` = daily_summary$daily_batT[2],
                                 `BatT d-1` = daily_summary$daily_batT[3],
                                 `BatT d0` = daily_summary$daily_batT[4],
                                 `BatT d1` = daily_summary$daily_batT[5],
                                 `BatT d2` = daily_summary$daily_batT[6])
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
tempDifference_fun("1N8476", 3)
bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
bat_delta <- bind_rows(bat_delta_list)
bat_delta_full <- bat_delta |>
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
  rowwise() |>
  mutate(`Preboost Mean dT` = mean(c(`DeltaT d-3`,`DeltaT d-2`,`DeltaT d-1`))) |>
  mutate(`Postboost Mean dT` = mean(c(`DeltaT d0`,`DeltaT d1`,`DeltaT d2`))) |>
  mutate(`Preboost Bat T` = mean (c(`BatT d-3`,`BatT d-2`,`BatT d-1`))) |>
  mutate(`Postboost Bat T` = mean(c(`BatT d0`,`BatT d1`,`BatT d2`)))
bat_delta_full
```

And let's try to look at the various possible models.

```{r Delta T Models}
first_model <- glm(`Postboost Mean dT` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

```{r Delta T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Mean dT`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day Postboost ',Delta,'T'))) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)
bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Mean dT`:`Postboost Mean dT`, names_to = "Boost")
ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day ',Delta,'T')))

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


Well that is quite disappointing. 
Tent is clearly the biggest factor driving this.
Delta T is not the best way to analyze this phenomenon.

First, we need to look at Bat T instead of delta T.


```{r Bat T Models}
first_model <- glm(`Postboost Bat T` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

This is the ideal situation for a mixed effects model!

Let's see how it works.

Following guidance at https://m-clark.github.io/mixed-models-with-R/random_intercepts.html

```{r}
bat_delta_T <- bat_delta_full |>
  mutate(`Boost dT` = `Postboost Bat T` - `Preboost Bat T`)

mixed_model <- lmer(`Boost dT` ~ Treatment + (1|Tent), data = bat_delta_T)
summary(mixed_model)
confint(mixed_model)
ranef(mixed_model)$Tent
coef(mixed_model)$Tent

predictInterval(mixed_model)   # for various model predictions, possibly with new data

REsim(mixed_model)             # mean, median and sd of the random effect estimates

plotREsim(REsim(mixed_model))  # plot the interval estimates

```

It looks like even after incorporating Tent as a random effect, the Treatment effect is not significant.
This may be because I am looking at the 3-day average and I need a more focused approach.

```{r Bat T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Bat T`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)

bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Bat T`:`Postboost Bat T`, names_to = "Boost") |>
  mutate(Boost = factor(Boost, levels = c("Preboost Bat T", "Postboost Bat T")))

ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_pivot, aes(x = Boost, y = value)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5, position = position_nudge()) +
  geom_line(aes(group=BatID), colour="red", linetype="11") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Treatment) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


### Day 1 Post-boost

Reviewing some of the data from the sham group, I think it would be better to look only at post-boost day 1 or day 2.
This should avoid some of the impact of handling. 

```{r}
bat_delta_T_day1 <- bat_delta_full |>
  mutate(`Boost dT day1` = `BatT d1` - `Preboost Bat T`) |>
  mutate(`Boost dT day2` = `BatT d2` - `Preboost Bat T`)

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d1`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d2`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

mixed_model_d1 <- lmer(`Boost dT day1` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d1)
confint(mixed_model_d1)
ranef(mixed_model_d1)$Tent
coef(mixed_model_d1)$Tent
predictInterval(mixed_model_d1)   # for various model predictions, possibly with new data
REsim(mixed_model_d1)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d1))  # plot the interval estimates

mixed_model_d2 <- lmer(`Boost dT day2` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d2)
confint(mixed_model_d2)
ranef(mixed_model_d2)$Tent
coef(mixed_model_d2)$Tent
predictInterval(mixed_model_d2)   # for various model predictions, possibly with new data
REsim(mixed_model_d2)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d2))  # plot the interval estimates
```

The d2 model might show something interesting for the Poly(I:C) group, but it is a small effect.


## Hourly Temperature Differences

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults3 <- singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                    .by = "hour", 
                    hourly_delta = sum(`delta T`)/6)


ggplot(singlebatresults3, aes(x = date, y = hourly_delta, color = Boosted)) + 
  geom_line() +
  theme_cowplot()
```
A function to quantify this delta T by hour:

```{r}
tempDifference_fun2 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 `Preboost d-3` = daily_summary$daily_delta[1]/144,
                                 `Preboost d-2` = daily_summary$daily_delta[2]/144,
                                 `Preboost d-1` = daily_summary$daily_delta[3]/144,
                                 `Boost d0` = daily_summary$daily_delta[4]/144,
                                 `Boost d1` = daily_summary$daily_delta[5]/144,
                                 `Boost d2` = daily_summary$daily_delta[6]/144)
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
# tempDifference_fun2("1N8476", 3)
# bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
# bat_delta <- bind_rows(bat_delta_list)
# bat_delta_full <- bat_delta |>
#   mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
#   rowwise() |>
#   mutate(`Preboost Mean dT` = mean(c(`Preboost d-3`,`Preboost d-2`,`Preboost d-1`))) |>
#   mutate(`Postboost Mean dT` = mean(c(`Boost d0`,`Boost d1`,`Boost d2`)))
# bat_delta_full
```
