---
title: "MOCO 2025 Temperature Loggers Analysis"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
#if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("zoo")) install.packages("zoo"); library(zoo)
if (!require("emmeans")) install.packages("emmeans"); library(emmeans)
if (!require("multcomp")) install.packages("multcomp"); library(multcomp)
if (!require("interactions")) install.packages("interactions"); library(interactions)
if (!require("colorspace")) install.packages("colorspace"); library(colorspace)
if (!require("broom")) install.packages("broom"); library(broom)
if (!require("performance")) install.packages("performance"); library(performance)
if (!require("MASS")) install.packages("MASS"); library(MASS)
if (!require("lmerTest")) install.packages("lmerTest"); library(lmerTest)
if (!require("patchwork")) install.packages("patchwork"); library(patchwork)
if (!require("ggsignif")) install.packages("ggsignif"); library(ggsignif)
if (!require("car")) install.packages("car"); library(car)
if (!require("Hmisc")) install.packages("Hmisc"); library(Hmisc)
#if (!require("ggeffects")) install.packages("ggeffects"); library(ggeffects)



conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

## Metadata

```{r}
Primary_met <- read.delim("MOCO2025Primary_Met.tsv", # read in metadata for primary 
     na = "NA") |> 
  mutate(Bat_ID = as.character(Bat_ID))

glimpse(Primary_met) # check formatting

Primary_Results <- read.delim("MOCO2025PrimaryTemperature.tsv", # read in results for primary 
     na = "NA") |> 
  mutate(Bat_ID = as.character(Bat_ID))

glimpse(Primary_Results)
```

```{r}
Secondary_met <- read.delim("MOCO2025Secondary_Met.tsv", # read in metadata for primary 
     na = "NA") |> 
  mutate(Bat_ID = as.character(Bat_ID))

glimpse(Secondary_met) # check formatting

Secondary_Results <- read.delim("MOCO2025SecondaryTemperature.tsv", # read in results for primary 
     na = "NA") |> 
  mutate(Bat_ID = as.character(Bat_ID))

glimpse(Secondary_Results)
```


I am going to load in my data and metadata for both immunizations

Make values for bat names and assign colors for groups
```{r}
PrimaryNames <- unique(Primary_met$Bat_ID)
SecondaryNames <- unique(Secondary_met$Bat_ID)


treatment_colors <- c("Cold" = "#9EB0FFFF",
                    "NoCo" = "#8A3F2A",
                    "SHAM" = "#11181DFF" )

facet_colors <- c("Cold" = "#9EB0FFFF",
                    "NoCo - C1" = "#8A3F2A",
                  "NoCo - C2" = "#8A3F2A",
                    "Sham - C1" = "#11181DFF",
                  "Sham - C2" = "#11181DFF")

primary_colors <- c("64" = "#7891FF", "68" = "#4D69E8","74" = "#1E3BB3")
```



## Traces and Visualizations 

```{r}
Primary_met <- Primary_met |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id))

Secondary_met <- Secondary_met |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id))
```


#### None Traces

```{r}
None_data_primary <- Primary_Results |> 
  filter(DateType == "Baseline") |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

none_p_met <- data.table::data.table(Primary_met, key = "id")
none_p_data <- data.table::data.table(None_data_primary, key = "id")

 none_p_behavr <- behavr(none_p_data, none_p_met)
 
 nonestartp <- min(none_p_behavr$t)
  noneendp <- max(none_p_behavr$t)
```
  
```{r}
none_summary <- None_data_primary %>%
  # STAGE 1: Calculate each bat's metrics per day
  group_by(Bat_ID, DateType, ExperimentDay) %>%
  summarise(
    bat_min  = min(Temperature, na.rm = TRUE),
    bat_max  = max(Temperature, na.rm = TRUE),
    bat_mean = mean(Temperature, na.rm = TRUE),
    .groups  = "drop")
    
    none_summary2 <- none_summary |> 
  group_by(Bat_ID, DateType) %>%
  summarise(
    n         = n(),
    # Daily Minimums
    Avg_Min   = mean(bat_min, na.rm = TRUE),
    sd_min    = sd(bat_min, na.rm = TRUE),
    # Daily Maximums
    Avg_Max   = mean(bat_max, na.rm = TRUE),
    sd_max    = sd(bat_max, na.rm = TRUE) ,
    # Daily Means
    Avg_Mean  = mean(bat_mean, na.rm = TRUE),
    sd_mean   = sd(bat_mean, na.rm = TRUE),
    .groups   = "drop"
  )

write.csv(none_summary2, "MOCO2025_none_summary.csv")

summary_df <- none_p_behavr %>%
  mutate(t_day = t %% 86400) %>% 
  group_by(Bat_ID, id, t_day) %>%
  summarise(
    # Bat Temperature (Mean ± SD)
    mean_temp  = mean(Temperature, na.rm = TRUE),
    sd_temp    = sd(Temperature, na.rm = TRUE),
    sd_lower   = mean_temp - sd_temp,
    sd_upper   = mean_temp + sd_temp,
    # Ambient Temperature (Mean ± SD)
    Amean_temp = mean(AmbientT, na.rm = TRUE),
    Asd_temp   = sd(AmbientT, na.rm = TRUE),
    Asd_lower  = Amean_temp - Asd_temp,
    Asd_upper  = Amean_temp + Asd_temp,
    .groups    = "drop"
  )

```

```{r}
MOCO_Nonep <- ggplot(summary_df, aes(x = t_day)) +
  # Light/Dark Bar
  stat_ld_annotations(
    period = 86400, phase = 25200, l_duration = 43200, 
    ld_colours = c("white", "black")
  ) +
  # Ambient Ribbon & Line
  geom_ribbon(data = summary_df, aes(ymin = Asd_lower, ymax = Asd_upper), fill = "grey80", color = NA, alpha = 0.5) +
  geom_line(data = summary_df, aes(y = Amean_temp), colour = "grey40", linewidth = 0.5) +
  
  # Bat Skin Temp Ribbon & Line
  geom_ribbon(data = summary_df, aes(ymin = sd_lower, ymax = sd_upper, fill = id), color = NA, alpha = 0.35) +
  geom_line(data = summary_df, aes(y = mean_temp, color = id), linewidth = 0.8) +
  
  scale_colour_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_grid(Bat_ID ~ .) +
  scale_x_continuous(
    name = "Zeitgeber Time (HH:MM)",
    breaks = seq(0, 86400, by = 10800),
    labels = function(x) sprintf("%02d:00", (x / 3600) %% 24),
    expand = c(0, 0)
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, vjust = 0.8),
    text = element_text(size = 15),
    panel.grid = element_blank()
  ) +
  labs(y = "Skin Temperature (°C)")

MOCO_Nonep


ggsave("MOCO_none.jpeg", plot = MOCO_Nonep, device = jpeg(), width = 7, height = 5.5)
```

```{r}
summary_df2 <- none_p_behavr %>%
  mutate(t_day = t %% 86400) %>% 
  group_by(Bat_ID, id, t_day) %>%
  summarise(
    # delta t (Mean ± SD)
    mean_DT  = mean(deltaT, na.rm = TRUE),
    sd_DT    = sd(deltaT, na.rm = TRUE),
    sd_lower   = mean_DT - sd_DT,
    sd_upper   = mean_DT + sd_DT,
    .groups    = "drop"
  )
```

### Primary Traces

```{r}
Primary_days <- Primary_Results |> 
  filter(Day_Number >= -4 & Day_Number <= 3) |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

none_p_met <- data.table::data.table(Primary_met, key = "id")
p_data <- data.table::data.table(Primary_days, key = "id")

p_behavr <- behavr(p_data, none_p_met)
 
 startp <- min(p_behavr$t)
 endp <- max(p_behavr$t)

p_summary <- p_behavr %>%
  group_by(Bat_ID, DateType) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE),
    n = n(), 
    sd_temp = sd(Temperature),
    se_temp  = sd_temp / sqrt(n),) |> 
  ungroup() 

write.csv(p_summary, "MOCO2025_primary_summary.csv")
```

```{r}
shaded <- data.frame(
  xmin = c(1748001900, 1748277000), 
  xmax = c(1748049000, 1748306400),
  ymin = -Inf,
  ymax = Inf)

injection <- data.frame(
  xmin = c(1748274300), 
  xmax = c(1748277000),
  ymin = -Inf,
  ymax = Inf)

day_ticks <- p_behavr[!duplicated(p_behavr$ExperimentDay), ]

  ggetho(p_behavr, aes(x = t, y = Temperature, colour = id, fill = id)) +
  # 1. Shaded region
  geom_rect(data = shaded, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
    geom_line(
      data = p_behavr,
    aes(x = t, y = AmbientT), # Map ambient temp column here
    colour = "grey50",    # Fixed grey color
    linewidth = 0.2,
    linetype = "solid",   # Use "dashed" if you want to distinguish it further
    inherit.aes = FALSE   # Ignores global 'colour = id' mapping
  ) +
    geom_rect(data = injection, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 1, inherit.aes = FALSE) +
    geom_line(linewidth = 0.8) +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,      
    l_duration = 43200, 
    ld_colours = c("white", "black")) +
  facet_grid(id ~ .) +
     scale_colour_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  scale_x_continuous(
    name = "Experiment Day",
    breaks = day_ticks$t,             
    labels = day_ticks$DateType,
    limits = c(min(p_behavr$t), max(p_behavr$t)),
    expand = c(0,0) ) +
  theme_bw() +
  theme(legend.position = "none",
    axis.text.x = element_text(angle = -25, vjust = 1, hjust = 0),
    text = element_text(size = 16),
    panel.grid = element_blank()) +
  labs(y = "Skin Temperature (°C)") -> MOCO_Primary

print(MOCO_Primary)

ggsave("MOCO_Primary.jpeg", plot = MOCO_Primary, device = jpeg(), width = 7, height = 5.5)

```

```{r}
Post_prim <- Primary_Results |> 
  filter(Day_Number >= 1 & Day_Number <= 3) |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

none_p_met <- data.table::data.table(Primary_met, key = "id")
p_data2 <- data.table::data.table(Post_prim, key = "id")

p_behavr2 <- behavr(p_data2, none_p_met)
 
 startp2 <- min(p_behavr$t)
 endp2 <- max(p_behavr$t)
```

```{r}
 MOCO_deltat <- ggplot(p_behavr2, aes(x = t %% 86400, y = deltaT, colour = Bat_ID)) +
  
  # 1. Light/Dark Bar Annotations
  stat_ld_annotations(
    period = 86400,
    phase = 25200,      
    l_duration = 43200, 
    ld_colours = c("white", "black")
  ) +
  
  # 2. Mean Delta T Line (from summary_df2)
  geom_line(
    data = summary_df2,
    aes(x = t_day, y = mean_DT),
    colour = "grey50",
    linewidth = 0.5,
    linetype = "solid",
    inherit.aes = FALSE
  ) +
  
  # 3. Skin Temperature Line
  geom_line(linewidth = 0.8) +
  
  # 4. Facet Grid (3 Bat Rows x 3 Day Columns)
  facet_grid(Bat_ID ~ DateType) +
  
  # 5. Colors and Aesthetics
  scale_colour_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  scale_x_continuous(
    name = "Zeitgeber Time (HH:MM)",
    breaks = seq(14400, 86400, by = 14400),
    labels = function(x) sprintf("%02d:00", (x / 3600) %% 24),
    expand = c(0, 0)
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, vjust = 0.8),
    text = element_text(size = 14),
    panel.grid = element_blank(),
    strip.text.y = element_text(angle = 0)
  ) +
  labs(y = "Delta T (°C)")

print(MOCO_deltat)

ggsave("MOCO_deltaT.jpeg", plot = MOCO_deltat, device = jpeg(), width = 7, height = 5.5)

```
```{r} 
Primary_Results_delta <- Primary_Results |> 
  filter(DateType %in% c("Baseline", "Primary + 1", "Primary + 2", "Primary + 3"))

p_summary2 <- Primary_Results_delta %>%
  group_by(Bat_ID, DateType) |> 
  summarise(
    Minimum = min(deltaT, na.rm = TRUE),
    Maximum = max(deltaT, na.rm = TRUE),
    Average = mean(deltaT, na.rm = TRUE),
    n = n(), 
    sd_dtemp = sd(deltaT),
    se_temp  = sd_dtemp / sqrt(n),) |> 
  ungroup() 

write.csv(p_summary2, "MOCO2025_deltat_summary.csv")
```


## Secondary Traces

```{r}
Secondary_Results <- Secondary_Results %>%
  mutate(
    Facet_Group = case_when(
      Treatment == "Cold" ~ "Cold",
      TRUE ~ paste0(Treatment, " - C", Group) # e.g. "NoCo - 1", "NoCo - 2"
    ),
    # Lock in exact vertical order: Cold at top, then NoCo 1/2, then SHAM 1/2
    Facet_Group = factor(
      Facet_Group, 
      levels = c("Cold", "NoCo - C1", "NoCo - C2", "Sham - C1", "Sham - C2")))

Secondary_met <- Secondary_met |> 
  mutate(
    Facet_Group = case_when(
      Treatment == "Cold" ~ "Cold",
      TRUE ~ paste0(Treatment, " - C", Group) # e.g. "NoCo - 1", "NoCo - 2"
    ),
    # Lock in exact vertical order: Cold at top, then NoCo 1/2, then SHAM 1/2
    Facet_Group = factor(
      Facet_Group, 
      levels = c("Cold", "NoCo - C1", "NoCo - C2", "Sham - C1", "Sham - C2")
    )
  )

Secondary_Results_filter <- Secondary_Results |> 
  filter(DateType %in% c("Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3"))

Secondary_Results_filter <- Secondary_Results_filter |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

Secondary_met <- Secondary_met |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id))

 Secondary_met2 <- data.table::data.table(Secondary_met, key = "id")
 Secondary_results2 <- data.table::data.table(Secondary_Results_filter, key = "id")

 secondary_behavr <- behavr(Secondary_results2, Secondary_met2)
 summary(secondary_behavr)
 
s_summary <- Secondary_Results %>%
  group_by(Bat_ID, DateType) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE),
    n = n(), 
    sd_temp = sd(Temperature),
    se_temp  = sd_temp / sqrt(n),) |> 
  ungroup() 

s_summary_treatment <- Secondary_Results %>%
  # STAGE 1: Calculate each bat's metrics per day
  group_by(Treatment, Bat_ID, DateType) %>%
  summarise(
    bat_min  = min(Temperature, na.rm = TRUE),
    bat_max  = max(Temperature, na.rm = TRUE),
    bat_mean = mean(Temperature, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  # STAGE 2: Means and SEs across bats per treatment per day
  group_by(Treatment, DateType) %>%
  summarise(
    n         = n(),
    
    # Daily Minimums
    Avg_Min   = mean(bat_min, na.rm = TRUE),
    se_min    = sd(bat_min, na.rm = TRUE) / sqrt(n),
    
    # Daily Maximums
    Avg_Max   = mean(bat_max, na.rm = TRUE),
    se_max    = sd(bat_max, na.rm = TRUE) / sqrt(n),
    
    # Daily Means
    Avg_Mean  = mean(bat_mean, na.rm = TRUE),
    se_mean   = sd(bat_mean, na.rm = TRUE) / sqrt(n),
    
    .groups   = "drop"
  )

write.csv(s_summary_treatment, "MOCO2025_secondary_summary.csv")
```

```{r}
secondary_etho  <- ggetho(secondary_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho() +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Temp (°C)",
       y = "Bat ID")

secondary_etho
ggsave("EPFU_Sec_Etho.jpeg", plot = secondary_etho, device = jpeg(), width = 12, height = 6.67)
```

```{r}
secondary_etho2  <- ggetho(secondary_behavr, aes(x=t, y=id, z=deltaT)) + stat_tile_etho() +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Delta T (°C)",
       y = "Bat ID")

secondary_etho2

ggsave("EPFU_Sec_Etho2.jpeg", plot = secondary_etho, device = jpeg(), width = 12, height = 6.67)
```

```{r}
ggetho(secondary_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r}

shaded2 <- data.frame(
  xmin = c(1749137400 ), 
  xmax = c(1749144600),
  ymin = -Inf,
  ymax = Inf)

sday_ticks <- secondary_behavr[!duplicated(secondary_behavr$ExperimentDay), ]

  secondary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Facet_Group)) +
    geom_rect(data = shaded2, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
    geom_line(
      data = secondary_behavr,
    aes(x = t, y = AmbientT), # Map ambient temp column here
    colour = "grey50",    # Fixed grey color
    linewidth = 0.2,
    linetype = "solid",   # Use "dashed" if you want to distinguish it further
    inherit.aes = FALSE   # Ignores global 'colour = id' mapping
  ) +
  stat_pop_etho(method = mean_cl_boot) +
     stat_ld_annotations(
    period = 86400,
    phase = 25200,      
    l_duration = 43200, 
    ld_colours = c("white", "black")) +
  facet_grid(Facet_Group ~ . ) +
  scale_x_continuous(
    name = "Experiment Day",
     breaks = sday_ticks$t,             
    labels = sday_ticks$DateType,
    limits = c(min(secondary_behavr$t), max(secondary_behavr$t)), 
    expand = c(0,0) )+
  scale_colour_manual(values = facet_colors) +
  scale_fill_manual(values = facet_colors) +
  theme_bw() +
  theme(legend.position = "null",
    axis.text.x = element_text(angle = -25, vjust = 1, hjust = 0)) +
  theme(text = element_text(size = 16), panel.grid = element_blank(),
        strip.text.y = element_text(
      size = 10,                                # Smaller font size for strip text
      margin = margin(l = 5, r = 5)            # Adds horizontal padding
    ))+
  labs(y = "Skin Temperature (°C)") -> MOCO_Secondary

print(MOCO_Secondary)

ggsave("MOCO_Secondary.jpeg", plot = MOCO_Secondary, device = jpeg(), width = 7, height = 5.5)
```

## Basal Model

```{r}
BasalvNone <- Analysis_results |> 
  filter(Treatment %in% c("CP", "CP+S", "CS"),
         DateType %in% c("Basal TC", "None"))
```

Comparing none to basal thermal challenge NTD
```{r}
Basal_none_mixed <- lmerTest::lmer(Total_AUC ~ DateType * Treatment + Weight + (1 | Group / Bat_ID),  data = BasalvNone)

# Check the summary
summary(Basal_none_mixed)

basalnonemixed_emmeans <- emmeans(Basal_none_mixed, ~ DateType | Treatment)

summary(basalnonemixed_emmeans)

basal_emmeans_df <- as.data.frame(basalnonemixed_emmeans)
basal_pvalues <- as.data.frame(pairs(basalnonemixed_emmeans, adjust = "none"))

write.csv(basal_pvalues, "EPFU2025_basalnone_Output.csv")

```

```{r}
Basal_TC <- Analysis_results |> 
  filter(Treatment %in% c("CP", "CP+S", "CS"),
         DateType == "Basal TC")
```


```{r}
options(contrasts = c("contr.sum", "contr.poly"))

Basal_lm <- lm(Total_AUC ~ Weight + Group + Treatment, data = Basal_TC)

# 3. Type III ANOVA: Check overall effect of Treatment
# Look at the p-value in the 'Treatment' row
Anova(Basal_lm, type = "3")

# 4. Extract weight-adjusted means for the 3 treatments
basal_emmeans_trt <- emmeans(Basal_lm, ~ Treatment)
summary(basal_emmeans_trt)

# 5. Run all 3 pairwise contrasts with P-value adjustment
# Use "fdr" or "tukey" (Tukey HSD is standard for all pairwise contrasts)
basal_pvalues <- pairs(basal_emmeans_trt, adjust = "tukey")
summary(basal_pvalues)

# 6. Save results to data frames
basal_emmeans_df <- as.data.frame(basal_emmeans_trt)
basal_pvalues_df <- as.data.frame(basal_pvalues)

write.csv(basal_pvalues_df, "EPFU2025_basaltreatment_Output.csv")

```

## Primary Models

Creating a dataframe for just primary data

```{r}
Primary_Results <- Analysis_results |> 
  filter(!ExperimentDay %in% c("D20", "D21", "D22", "D23", "D24", "D25", "D26", "D27", "D28", "D29", "D30", "D31", "D32", "D33", "D34", "D35", "D36"),
         !DateType %in% c("Basal TC-1", "Basal TC", "Basal TC + 1", "Pre"))

#drops experiment day levels that are no longer in the primary data frame after filtering
Primary_Results <- droplevels(Primary_Results)
levels(Primary_Results$Primary_Combo)

```

Okay, I am going to run some primary models now.

```{r}
primary_glm <- glm(Total_AUC ~ Bat_ID + Group + Weight + Primary_Combo, data = Primary_Results)

primary_emmeans <- emmeans(primary_glm, ~ Primary_Combo)

primary_subset <- list(
  "SvN_P" = c(0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0),
  "SvN_1" = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0),
  "SvN_2" = c(0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0),
  "SvN_3" = c(0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1),
  "NvC_P" = c(0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
  "NvC_1" = c(0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
  "NvC_2" = c(0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0),
  "NvC_3" = c(0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0),
  "SvN_None" = c(0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0),
  "NvC_None" = c(1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0),
  "SvN_-1" = c(0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0),
  "NvC_-1" = c(0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0))

# 3. Run the contrast. 
# R will automatically apply these exact vectors to Males, and then to Females.
contrast(primary_emmeans, method = primary_subset, adjust = "fdr")
```
Okay, I am getting NAs and I think it is because of Bat_ID
I am going to rerun the model without Bat_ID
```{r}
primary_glm2 <- glm(Total_AUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

primary_emmeans2 <- emmeans(primary_glm2, ~ Primary_Combo)

contrast(primary_emmeans2, method = primary_subset, adjust = "fdr")
```
Okay this ran and the significant values make sense; however, after asking some questions with gemini I don't think I can ignore Bat_ID because the glm assumes that every row is independent of other rows. Since bat_id is found multiple times across rows they are not independent of each other. It wants me to run a mixed effects model. I know Ken and I talked about this and he doesnt want me to run one, but I am going to anyways.

```{r}
primary_mixed <- lmerTest::lmer(Total_AUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(primary_mixed)

primarymixed_emmeans <- emmeans(primary_mixed, ~ Primary_Combo)
primary_emmeans_df <- as.data.frame(primarymixed_emmeans)

# 2. Run your custom contrasts
primary_contrasts <- as.data.frame(contrast(primarymixed_emmeans, method = primary_subset, adjust = "fdr"))

write.csv(primary_contrasts, "EPFU2025_primary_group_Output.csv")

```

redoing all my models without group becaus
Okay, this shows that there is no statistically significant difference between sham and noco for any primary days including non-handled days. Bats that were cold at primary were significantly different than noco bats on primary which makes sense, but also on Primary +2 which is interesting.

#### Figure for primary model

```{r}
em_df_split <- primary_emmeans_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

em_df_split <- em_df_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```

Makes the divider between top and bottom figures
```{r}
p_divider <- ggplot() + 
  theme_void() + 
  geom_hline(yintercept = 0.5, color = "gray70", linewidth = 0.8)
```

Renaming some things in raw data so it matches my dataframe
```{r}
Primary_raw <- Primary_Results |> 
  rename("Assigned_treatment" = "Treatment") |> 
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Treatment, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Treatment, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Treatment, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ))

Primary_raw <- Primary_raw |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
```

Also adding significant bars
```{r}
prim_sig_bars <- primary_contrasts
prim_sig_bars <- prim_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_P", contrast) ~ "Primary",
    grepl("_1", contrast) ~ "Primary + 1",
    grepl("_2", contrast) ~ "Primary + 2",
    grepl("_3", contrast) ~ "Primary + 3",
    grepl("_-1", contrast) ~ "Primary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

prim_sig_bars <- prim_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvC", contrast) ~ "NoCo",
    TRUE ~ "other"))

prim_sig_bars <- prim_sig_bars |> 
  mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvC", contrast) ~ "Cold",
    TRUE ~ "other"),
    p.value2 = sprintf("%.2e", p.value))

comparison_bars_primary <- prim_sig_bars 

prim_sig_bars <- prim_sig_bars |> 
  filter(p.value < 0.05)

prim_sig_bars <- prim_sig_bars |> 
 mutate(y_position = case_when(
    contrast == "NvC_P" ~ 125,
    contrast == "NvC_2" ~ 125  ))

```

Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_primary <- comparison_bars_primary |> 
  filter(DateType == "Primary + 3")

comparison_bars_primary <- comparison_bars_primary |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -220,
    contrast == "NvC_3" ~ -180  ))

comparison_bars_primary <- comparison_bars_primary |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))

prim_sig_bars <- prim_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))
```

```{r}
p_bottom2 <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = em_df_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = em_df_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3
  ) +
   geom_signif(data = prim_sig_bars,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = p.value2, 
      y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 5,
    vjust = -0.1) +
  geom_signif(data = comparison_bars_primary,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.05,
    textsize = 16,
    vjust = -0.3) +
    coord_cartesian(ylim = c(-250, 135)) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = "NTD (Degree Hours) EMM", color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 14, face = "bold")
  )
print(p_bottom2)

ggsave(
  filename = "primary_bottom_plot.png", 
  plot = p_bottom2, 
  width = 8,  # Narrower width (inches)
  height = 7,  # Much taller height (inches)
  dpi = 300)
```

### Longitudinal Primary

```{r}
primary_long_mixed <- lmerTest::lmer(Total_AUC ~ DateType * Primary_Treatment + Weight + (1 | Group / Bat_ID),  data = Primary_Results)

# Check the summary
summary(primary_long_mixed)

primary_long_emmeans <- emmeans(primary_long_mixed, ~ DateType | Primary_Treatment)

summary(primary_long_emmeans)

primary_long_em_df <- as.data.frame(primary_long_emmeans)

primary_long_pval <- as.data.frame(pairs(primary_long_emmeans, adjust = "fdr"))

write.csv(primary_long_em_df, "EPFU2025_primary_emm.csv")

write.csv(primary_long_pval, "EPFU2025_primary_long_Output.csv")
```

##### Longitudinal figure

```{r}
plong_em_split <- primary_long_em_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Treatment, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Treatment, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Treatment, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ))

plong_em_split <- plong_em_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make a longitudinal figure that shows EMM for each primary treatment group over the primary period
```{r}
pd <- position_dodge(width = 0.2) # used to shift groups so they are not overlapping

p_longitudinal2 <- ggplot() +
    geom_line(
    data = em_df_split,
    aes(x = DateType, y = emmean, color = Treatment, group = Treatment),
    linewidth = 1,
    position = pd
  ) +
  geom_point(
    data = em_df_split,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3,
    position = pd
  ) +
  geom_errorbar(
    data = em_df_split,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.3, # changes width of t bars 
    linewidth = 0.7, # changes width of the line
    position = pd
  ) +
  scale_color_manual(values = primary_colors) +
  labs(x = "Experiment Day", y = "NTD (Degree Hours) EMM", color = "Primary Treatment",
       fill = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold")  )
print(p_longitudinal2)

ggsave(
  filename = "primary_longitudianl2_plot.png", 
  plot = p_longitudinal2, 
  width = 7,  # Make narrower (inches)
  height = 6,  # Make taller  (inches)
  dpi = 300)
```

### Hypo NTD
I am going to run the same models with Hypo and Hyper AUC. I am doing this because I want to see if immunized bats had a febrile response and if they did, if the thermal challenge affected that response. The way AUC is calculated currently 'cancels' the extreme values out.

```{r}
primary_hypo <- glm(HypoAUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

p_hypo_emmeans <- emmeans(primary_hypo, ~ Primary_Combo)

contrast(p_hypo_emmeans, method = primary_subset, adjust = "fdr")
```

```{r}
hypo_mixed <- lmerTest::lmer(HypoAUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(hypo_mixed)

hypomixed_emmeans <- emmeans(hypo_mixed, ~ Primary_Combo)

levels(hypomixed_emmeans)

# 2. Run your custom contrasts
p_hypo_contrast <- as.data.frame(contrast(hypomixed_emmeans, method = primary_subset, adjust = "fdr"))
```

##### Hypo Figure

```{r}
hypo_df_p <- as.data.frame(emmeans(hypo_mixed, ~ Primary_Combo))

hypo_split <- hypo_df_p %>%
  mutate(
    # Extract Factor A using exact prefix matching
        Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

hypo_split <- hypo_split |> 
  mutate(SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

# Interaction plot with connecting lines and capped error bars
ggplot(hypo_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hypo NTD)",
    color = "Primary Treatment",
    title = "Interaction Effect on Hypo NTD"
  ) +
  theme_bw()
```

```{r}

p_hypo_sig_bars <- p_hypo_contrast
p_hypo_sig_bars <- p_hypo_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_P", contrast) ~ "Primary",
    grepl("_1", contrast) ~ "Primary + 1",
    grepl("_2", contrast) ~ "Primary + 2",
    grepl("_3", contrast) ~ "Primary + 3",
    grepl("_-1", contrast) ~ "Primary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

p_hypo_sig_bars <- p_hypo_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvC_", contrast) ~ "NoCo",
    TRUE ~ "other"))

p_hypo_sig_bars <- p_hypo_sig_bars |> 
   mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvC_", contrast) ~ "Cold",
    TRUE ~ "other"))

p_hypo_sig_bars <- p_hypo_sig_bars |> 
  mutate(p.value2 = sprintf("%.2e", p.value),
         annotation = paste("P-Value =", p.value2))

comparison_bars_prim_hypo <- p_hypo_sig_bars 

p_hypo_sig_bars <- p_hypo_sig_bars |> 
  filter(p.value < 0.05)

p_hypo_sig_bars <- p_hypo_sig_bars |> 
 mutate(y_position = case_when(
    contrast == "NvC_P" ~ 22,
    contrast == "NvC_2" ~ 18))
```


Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_prim_hypo <- comparison_bars_prim_hypo |> 
  filter(DateType == "Primary + 3")

comparison_bars_prim_hypo <- comparison_bars_prim_hypo |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -200,
    contrast == "NvC_3" ~ -150))

p_hypo_sig_bars <- p_hypo_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))
comparison_bars_prim_hypo <- comparison_bars_prim_hypo |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))
Primary_raw <- Primary_raw |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
          Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
hypo_split <- hypo_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
```


```{r}
p_hypo_top <- ggplot() +
  geom_jitter(
    data = subset(Primary_raw, Treatment == "NoCo"),
    aes(x = DateType, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(hypo_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(hypo_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(hypo_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-275, 30)) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_hypo_top)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
p_hypo_bottom <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = hypo_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = hypo_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3) +
  geom_signif(data = p_hypo_sig_bars,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = annotation, 
      y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 3,
    vjust = -0.2) +
   geom_signif(data = comparison_bars_prim_hypo,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.02) +
  coord_cartesian(ylim = c(-275, 30)) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = NULL, color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(p_hypo_bottom)

primary_hypo_plot <- (p_hypo_top / p_divider / p_hypo_bottom) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "Hypo TD (Degree Hours)") 

print(primary_hypo_plot)

ggsave(
  filename = "primary_HypoNTDplot.png", 
  plot = primary_hypo_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Hypo Longitudinal Primary

```{r}
p_hypo_long_mixed <- lmerTest::lmer(HypoAUC ~ DateType * Primary_Treatment + Weight + (1 | Group / Bat_ID),  data = Primary_Results)

# Check the summary
summary(p_hypo_long_mixed)

p_hypo_long_emmeans <- emmeans(p_hypo_long_mixed, ~ DateType | Primary_Treatment)

summary(p_hypo_long_emmeans)

p_hypo_long_em_df <- as.data.frame(p_hypo_long_emmeans)

p_hypo_long_pval <- as.data.frame(pairs(p_hypo_long_emmeans, adjust = "fdr"))
```

##### Longitudinal figure

```{r}
plong_hypo_em_split <- p_hypo_long_em_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Treatment, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Treatment, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Treatment, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ))

plong_hypo_em_split <- plong_hypo_em_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make the top part of the figure for all 3 groups and it be 3 facets

```{r}
p_hypo_long <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = DateType, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = hypo_split,
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = hypo_split,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = hypo_split,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-275, 30)) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = "Hypo TD (Degree Hours") +
  theme_bw() +
     facet_wrap(~ Treatment, ncol = 1) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_hypo_long)

ggsave(
  filename = "primary_hypo_longitudinal.png", 
  plot = p_hypo_long, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```


### Hyper NTD

```{r}
primary_hyper <- glm(HyperAUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

p_hyper_emmeans <- emmeans(primary_hyper, ~ Primary_Combo)

contrast(p_hyper_emmeans, method = primary_subset, adjust = "fdr")
```
Okay, this is very interesting. I am going to run this as a mixed effects model just so I am including Bat_ID still

```{r}
hyper_mixed <- lmerTest::lmer(HyperAUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(hyper_mixed)

hypermixed_emmeans <- emmeans(hyper_mixed, ~ Primary_Combo)

levels(hypermixed_emmeans)



# 2. Run your custom contrasts

contrast(hypermixed_emmeans, method = primary_subset, adjust = "fdr")
p_contr_hyper <- as.data.frame(contrast(hypermixed_emmeans, method = primary_subset, adjust = "fdr"))
```

#### Hyper Figure 

```{r}
hyper_df_p <- as.data.frame(emmeans(hyper_mixed, ~ Primary_Combo))

hyper_split <- hyper_df_p %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

hyper_split <- hyper_split |> 
  mutate(SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

# Interaction plot with connecting lines and capped error bars
ggplot(hyper_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hyper NTD)",
    color = "Primary Treatment",
    title = "Interaction Effect on Hyper NTD"
  ) +
  theme_bw()
```

```{r}
prim_hyper_sig_bars <- p_contr_hyper
prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_P", contrast) ~ "Primary",
    grepl("_1", contrast) ~ "Primary + 1",
    grepl("_2", contrast) ~ "Primary + 2",
    grepl("_3", contrast) ~ "Primary + 3",
    grepl("_-1", contrast) ~ "Primary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvC_", contrast) ~ "NoCo",
    TRUE ~ "other"))

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
   mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvC_", contrast) ~ "Cold",
    TRUE ~ "other"))

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
  mutate(p.value2 = sprintf("%.2e", p.value),
         annotation = paste("P-Value =", p.value2))

comparison_bars_prim_hyper <- prim_hyper_sig_bars 

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
  filter(p.value < 0.05)

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
 mutate(y_position = case_when(
    contrast == "SvN_P" ~ 95,
    contrast == "NvC_P" ~ 83,
    contrast == "NvC_1" ~ 65,
    contrast == "NvC_2" ~ 65,
    contrast == "NvC_3" ~ 65))
```


Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_prim_hyper <- comparison_bars_prim_hyper |> 
  filter(DateType == "Primary + 3")

comparison_bars_prim_hyper <- comparison_bars_prim_hyper |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -12,
    contrast == "NvC_3" ~ -6))

prim_hyper_sig_bars <- prim_hyper_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))
comparison_bars_prim_hyper <- comparison_bars_prim_hyper |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))
Primary_raw <- Primary_raw |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
          Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
hyper_split <- hyper_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
```


```{r}
p_hyper_top <- ggplot() +
  geom_jitter(
    data = subset(Primary_raw, Treatment == "NoCo"),
    aes(x = DateType, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(hyper_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(hyper_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(hyper_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
   coord_cartesian(ylim = c(-10, 95)) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_hyper_top)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
p_hyper_bottom <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = hyper_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = hyper_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3) +
  geom_signif(data = prim_hyper_sig_bars,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = annotation, 
      y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 3,
    vjust = -0.2) +
   geom_signif(data = comparison_bars_prim_hyper,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.02) +
  coord_cartesian(ylim = c(-10, 95)) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = NULL, color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(p_hyper_bottom)

primary_hyper_plot <- (p_hyper_top / p_divider / p_hyper_bottom) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "Hyper TD (Degree Hours)") 

print(primary_hyper_plot)

ggsave(
  filename = "primary_HyperNTDplot.png", 
  plot = primary_hyper_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Hyper Longitudinal Primary

```{r}
p_hyper_long_mixed <- lmerTest::lmer(HyperAUC ~ DateType  * Primary_Treatment + Weight + (1 | Group / Bat_ID),  data = Primary_Results)

# Check the summary
summary(p_hyper_long_mixed)

p_hyper_long_emmeans <- emmeans(p_hyper_long_mixed, ~ DateType | Primary_Treatment)

summary(p_hyper_long_emmeans)

p_hyper_long_em_df <- as.data.frame(p_hyper_long_emmeans)

p_hyper_long_pval <- as.data.frame(pairs(p_hyper_long_emmeans, adjust = "fdr"))
```

##### Longitudinal figure

```{r}
plong_hyper_em_split <- p_hyper_long_em_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Treatment, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Treatment, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Treatment, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ))

plong_hyper_em_split <- plong_hyper_em_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make the top part of the figure for all 3 groups and it be 3 facets

```{r}
p_hyper_long <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = DateType, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = hyper_split,
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = hyper_split,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = hyper_split,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = "Hyper TD (Degree Hours") +
  theme_bw() +
  coord_cartesian(ylim = c(-10, 95)) +
     facet_wrap(~ Treatment, ncol = 1) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_hyper_long)

ggsave(
  filename = "primary_hyper_longitudinal.png", 
  plot = p_hyper_long, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Torpor effects 
This code is to test is NTD on one day effects NTD on the next day.
```{r}
Exploratory_primary <- Primary_Results |> 
  filter(!DateType == "None")

Exploratory_primary <- Exploratory_primary %>%
  # Sort by Bat_ID first, then numerically by date
  arrange(Bat_ID, Day_Num) %>% 
  # Group by Bat_ID so lag() doesn't accidentally pull data from a different bat
  group_by(Bat_ID) %>% 
  # Create the lag column using the current day's AUC
  mutate(Lagged_NTD = lag(Total_AUC)) %>% 
  # Always ungroup when finished to prevent downstream coding issues
  ungroup()
```

```{r}
primary_lagged <- lmerTest::lmer(
  Total_AUC ~ Lagged_NTD * Primary_Treatment + DateType + Weight + (1 | Group / Bat_ID),
  data = Exploratory_primary)

summary(primary_lagged) # 0 variance for bat_id but keeping in the model because on secondary there is still variance

p_lag_slope <- emtrends(
  primary_lagged, 
  ~ Primary_Treatment, 
  var = "Lagged_NTD")

# View individual treatment slopes (are they significantly different from 0?)
summary(p_lag_slope, infer = c(TRUE, TRUE))

levels(p_lag_slope) # checking order to make list in the right order
plag_contrast <- list("NoCo vs Cold" = c(1, -1, 0),
                      "SHAM vs NoCo" = c(0, 1, -1))

# Step B: Pairwise contrasts between treatment slopes (post-hoc comparisons)
contrast(p_lag_slope, method = plag_contrast, adjust = "holm")
```


## Secondary Models

Okay, now I am going to run the same things but for secondary models. I need to decide if I should run this model with their original treatment groups or their 'this is what they actually did groups' since I have some bats that did not go cold in the primary thermal challenge.

I am going to look at a histogram and see if I can tell which bats used torpor or not. I am going to look at both total AUC and hypoAUC. Just in case there are some bats that used torpor, but if they remained euthermic or even had a fever at some point during the day, their total AUC might not be as low.

```{r}
# 1. Visualize the distribution to find the "cut point"
Primary_only_results <- Analysis_results |> 
  filter(DateType == "Primary")

hist(Primary_only_results$Total_AUC, breaks = 20, main = "Distribution of Primary Torpor")
hist(Primary_only_results$HypoAUC, breaks = 20, main = "Hypo AUC")

```
I talked to Ken about this and the bats did not go cold during the primary thermal challenge still experienced different energetic requirements that bats that were not cold at primary. For that reason, we should not group them together for the secondary analysis even if they were not cold at the primary.

```{r}
Secondary_Results <- Analysis_results |> 
  filter(!ExperimentDay %in% c("D0", "D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14", "D15", "D16", "D17", "D18", "D19", "D-1", "D27", "D28"),
         !DateType %in% c("Basal TC-1", "Basal TC", "Basal TC + 1", "Pre"))

secondary_bats <- Secondary_Results %>%
  filter(ExperimentDay == "D21") %>%
  pull(Bat_ID) %>%
  unique()

Secondary_Results_filtered <- Secondary_Results %>%
  filter(Bat_ID %in% secondary_bats)

Secondary_Results <- droplevels(Secondary_Results)
levels(Secondary_Results$Group_Combo)
```

I am going to run the same models for secondary as I did for primary
```{r}
secondary_glm <- glm(Total_AUC ~ Bat_ID + Group + Weight + Group_Combo, data = Secondary_Results_filtered)

secondary_emmeans <- emmeans(secondary_glm, ~ Group_Combo)

secondary_subset <- list(
  "SvN_S" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0),
"SvN_1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0),
"SvN_2"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0),
"SvN_3"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1),
"SvN_-1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0),
"SvN_None"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0),
"NvCP_S" = c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
"NvCP_1"= c(0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
"NvCP_2"= c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0),
"NvCP_3"= c(0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0),
"NvCP_-1"= c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0),
"NvCP_None"= c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_S"= c(0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_1"= c(0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_2"= c(0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_3"= c(0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_-1"= c(0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_None"= c(0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_S"= c(0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_1"= c(0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_2"= c(0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_3" = c(0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_-1"= c(0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_None" = c(-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"NvCS_S"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
"NvCS_1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
"NvCS_2" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0),
"NvCS_3" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0),
"NvCS_-1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0),
"NvCS_None" = c(0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0))

# 3. Run the contrast. 
# R will automatically apply these exact vectors to Males, and then to Females.
contrast(secondary_emmeans, method = secondary_subset, adjust = "fdr")
```
Okay, I am getting NAs and I think it is because of Bat_ID
I am going to rerun the model without Bat_ID
```{r}
secondary_glm2 <- glm(Total_AUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

secondary_emmeans2 <- emmeans(secondary_glm2, ~ Group_Combo)

contrast(secondary_emmeans2, method = secondary_subset, adjust = "fdr")
```

Running mixed effects model.. 
```{r}
secondary_mixed <- lmerTest::lmer(Total_AUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(secondary_mixed)

secondarymixed_emmeans <- emmeans(secondary_mixed, ~ Group_Combo)
secondarymixed_emmeansdf <- as.data.frame(emmeans(secondary_mixed, ~ Group_Combo))


# 2. Run your custom contrasts
secondary_contrasts <- as.data.frame(contrast(secondarymixed_emmeans, method = secondary_subset, adjust = "fdr"))

write.csv(secondary_contrasts, "EPFU2025_secondary_Output.csv")
write.csv(secondarymixed_emmeansdf, "EPFU2025_secondary_emm.csv")
\

```
So there is no statistically difference in AUC between SHAM and NoCo, CP+S and CS, or NoCo and CP on any day. CP+S had a significantly lower AUC than CP on D21 and CS had a significantly lower AUC than Noco on D21. 

#### Secondary Figure

```{r}
Sec_df <- as.data.frame(secondarymixed_emmeans, level = 0.95)

sec_df_split <- Sec_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

sec_df_split <- sec_df_split |> 
  mutate(SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

```

Okay rerunning with the same as primary.

```{r}
sec_df_split <- sec_df_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

sec_df_split <- sec_df_split |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")))

Secondary_Raw <- Secondary_Results_filtered |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")))
```

I want to add significant bars
```{r}
Secondary_sig_bars <- secondary_contrasts
Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_S", contrast) ~ "Secondary",
    grepl("_1", contrast) ~ "Secondary + 1",
    grepl("_2", contrast) ~ "Secondary + 2",
    grepl("_3", contrast) ~ "Secondary + 3",
    grepl("_-1", contrast) ~ "Secondary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvCP_", contrast) ~ "NoCo",
    grepl("CPSvCS_", contrast) ~ "CP+S",
    grepl("CPvCPS_", contrast) ~ "CP",
    grepl("NvCS_", contrast) ~ "NoCo",
    TRUE ~ "other"))

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvCP_", contrast) ~ "CP",
    grepl("CPSvCS_", contrast) ~ "CS",
    grepl("CPvCPS_", contrast) ~ "CP+S",
    grepl("NvCS_", contrast) ~ "CS",
    TRUE ~ "other"))

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(annotation = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*" ),
      p.value2 = sprintf("%.2e", p.value))

comparison_bars_secondary <- Secondary_sig_bars 

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(annotation = paste("P-Value =", p.value2))

Secondary_sig_bars <- Secondary_sig_bars |> 
  filter(p.value < 0.05)

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(y_position = case_when(
    contrast == "CPvCPS_S" ~ 22,
    contrast == "NvCS_S" ~ 62  ))

Secondary_Raw <- Secondary_Raw |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

Secondary_sig_bars <- Secondary_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

comparison_bars_secondary <- comparison_bars_secondary |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))
```

Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_secondary <- comparison_bars_secondary |> 
  filter(DateType == "Secondary + 3")

comparison_bars_secondary <- comparison_bars_secondary |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -240,
    contrast == "NvCP_3" ~ -140,
     contrast == "CPSvCS_3" ~ -190,
    contrast == "CPvCPS_3" ~ -165,
     contrast == "NvCS_3" ~ -215))

```


```{r}
s_bottom <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = sec_df_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = sec_df_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3) +
  geom_signif(data = Secondary_sig_bars,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = p.value2, 
      y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 4.2,
    vjust = -0.5) +
   geom_signif(data = comparison_bars_secondary,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.02,
    textsize = 3,
    vjust = -0.6) +
  coord_cartesian(ylim = c(-270, 80)) +
  scale_color_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Secondary Treatment", y = "NTD (Degree Hours) EMM", color = "Secondary Treatment") +
  theme_bw() +
  theme( legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 14, face = "bold"))
print(s_bottom)

ggsave(
  filename = "secondary_NTD_bottom_plot.png", 
  plot = s_bottom, 
  width = 8,  # Narrower width (inches)
  height = 7,  # Much taller height (inches)
  dpi = 300)
```

### Longitudinal Secondary

```{r}
s_long_mixed <- lmerTest::lmer(Total_AUC ~ DateType * Treatment + Weight + (1 | Group / Bat_ID),  data = Secondary_Results_filtered)

# Check the summary
summary(s_long_mixed)

s_long_emmeans <- emmeans(s_long_mixed, ~ DateType | Treatment)

summary(s_long_emmeans)

s_long_em_df <- as.data.frame(s_long_emmeans)

s_long_pval <- as.data.frame(pairs(s_long_emmeans, adjust = "fdr"))

write.csv(s_long_pval, "EPFU2025_secondary_long_Output.csv")

```

##### Longitudinal figure

```{r}
s_long_em_df <- s_long_em_df |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

s_long_em_df <- s_long_em_df |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make the top part of the figure for all 3 groups and it be 3 facets

```{r}
pd <- position_dodge(width = 0.4) # used to shift groups so they are not overlapping

s_long <- ggplot() +
    geom_line(
    data = s_long_em_df,
    aes(x = DateType, y = emmean, color = Treatment, group = Treatment),
    linewidth = 1,
    position = pd
  ) +
  geom_point(
    data = s_long_em_df,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3,
    position = pd
  ) +
  geom_errorbar(
    data = s_long_em_df,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.3, # changes width of t bars 
    linewidth = 0.7, # changes width of the line
    position = pd
  ) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "Experiment Day", y = "NTD (Degree Hours) EMM", color = "Treatment",
       fill = "Treatment") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1),
    legend.position = "bottom",
    legend.direction = "horizontal",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold")  )
print(s_long)

ggsave(
  filename = "secondary_longitudinal_plot.png", 
  plot = s_long, 
  width = 6.5,  # Make narrower (inches)
  height = 5,  # Make taller  (inches)
  dpi = 300)
```


### Hypo Models
I am going to run the same models with Hypo and Hyper AUC. I am doing this because I want to see if immunized bats had a febrile response and if they did, if the thermal challenge affected that response. The way AUC is calculated currently 'cancels' the extreme values out.

```{r}
secondary_hypo <- glm(HypoAUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

s_hypo_emmeans <- emmeans(secondary_hypo, ~ Group_Combo)

contrast(s_hypo_emmeans, method = secondary_subset, adjust = "fdr")
```
**I tried running this on hyper and i was getting an error. I think the error was because variance for hyper was almost 0, so they model was not able to factor in group and bat_id as a random effect.**
```{r}
s_hypo_mixed <- lmerTest::lmer(HypoAUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hypo_mixed)

s_hypo_emmeans <- emmeans(s_hypo_mixed, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(s_hypo_emmeans, method = secondary_subset, adjust = "fdr")
```

```{r}
s_hypo_mixed2 <- lmerTest::lmer(HypoAUC ~ Group_Combo + Weight + (1 | Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hypo_mixed2)

s_hypo_emmeans2 <- emmeans(s_hypo_mixed2, ~ Group_Combo)

# 2. Run your custom contrasts
contrast_sec_hypo <- as.data.frame(contrast(s_hypo_emmeans2, method = secondary_subset, adjust = "fdr"))
```


#### Hypo Figure

```{r}
S_hypo_df <- as.data.frame(s_hypo_emmeans)

s_hypo_split <- S_hypo_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

s_hypo_split <- s_hypo_split |> 
  mutate(SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

# Interaction plot with connecting lines and capped error bars
ggplot(s_hypo_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hypo NTD)",
    color = "Secondary Treatment",
    title = "Interaction Effect on Hypo NTD"
  ) +
  theme_bw()
```

```{r}
s_hypo_split <- s_hypo_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

s_hypo_split <- s_hypo_split |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")))
```

I want to add significant bars
```{r}
Sec_hypo_sig_bars <- contrast_sec_hypo
Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_S", contrast) ~ "Secondary",
    grepl("_1", contrast) ~ "Secondary + 1",
    grepl("_2", contrast) ~ "Secondary + 2",
    grepl("_3", contrast) ~ "Secondary + 3",
    grepl("_-1", contrast) ~ "Secondary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvCP_", contrast) ~ "NoCo",
    grepl("CPSvCS_", contrast) ~ "CP+S",
    grepl("CPvCPS_", contrast) ~ "CP",
    grepl("NvCS_", contrast) ~ "NoCo",
    TRUE ~ "other"))

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvCP_", contrast) ~ "CP",
    grepl("CPSvCS_", contrast) ~ "CS",
    grepl("CPvCPS_", contrast) ~ "CP+S",
    grepl("NvCS_", contrast) ~ "CS",
    TRUE ~ "other"))

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(p.value2 = sprintf("%.2e", p.value),
         annotation = paste("P-Value =", p.value2))

comparison_bars_secondary_hypo <- Sec_hypo_sig_bars 

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  filter(p.value < 0.05)

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(y_position = case_when(
    contrast == "CPvCPS_S" ~ 22,
    contrast == "NvCS_S" ~ 62  ))
```

Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_secondary_hypo <- comparison_bars_secondary_hypo |> 
  filter(DateType == "Secondary + 3")

comparison_bars_secondary_hypo <- comparison_bars_secondary_hypo |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -330,
    contrast == "NvCP_3" ~ -180,
     contrast == "CPSvCS_3" ~ -250,
    contrast == "CPvCPS_3" ~ -210,
     contrast == "NvCS_3" ~ -290))

comparison_bars_secondary_hypo <- comparison_bars_secondary_hypo |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

Sec_hypo_sig_bars <- Sec_hypo_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))
```


```{r}
s_hypo_top <- ggplot() +
  geom_jitter(
    data = subset(Secondary_Raw, Treatment == "NoCo"),
    aes(x = DateType, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(s_hypo_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(s_hypo_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(s_hypo_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-335, 80)) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(s_hypo_top)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
s_hypo_bottom <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = Treatment, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = s_hypo_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = s_hypo_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3) +
  geom_signif(data = Sec_hypo_sig_bars,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = annotation, 
      y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 3,
    vjust = -0.5) +
   geom_signif(data = comparison_bars_secondary_hypo,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.05,
    textsize = 3,
    vjust = -0.6) +
  coord_cartesian(ylim = c(-335, 80)) +
  scale_color_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Secondary Treatment", y = NULL, color = "Secondary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(s_hypo_bottom)

secondary_hypo_plot <- (s_hypo_top / p_divider / s_hypo_bottom) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "Hypo TD (Degree Hours)") 

print(secondary_hypo_plot)

ggsave(
  filename = "secondary_HypoNTDplot.png", 
  plot = secondary_hypo_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Longitudinal Hypo Secondary

```{r}
s_long_hypo_mixed <- lmerTest::lmer(HypoAUC ~ DateType * Treatment + Weight + (1 | Group / Bat_ID),  data = Secondary_Results_filtered)

# Check the summary
summary(s_long_hypo_mixed)

s_long_hypo_emmeans <- emmeans(s_long_hypo_mixed, ~ DateType | Treatment)

summary(s_long_hypo_emmeans)

s_long_hypo_em_df <- as.data.frame(s_long_hypo_emmeans)

s_long__hypo_pval <- as.data.frame(pairs(s_long_hypo_emmeans, adjust = "fdr"))
```

##### Longitudinal Hypo figure

```{r}
s_long_hypo_emmeans <- s_long_hypo_em_df |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make the top part of the figure for all 3 groups and it be 3 facets

```{r}
s_long_hypo <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = DateType, y = HypoAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = s_long_hypo_emmeans,
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = s_long_hypo_emmeans,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = s_long_hypo_emmeans,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-300, 80)) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "DateType", y = "Hypo TD (Degree Hours") +
  theme_bw() +
     facet_wrap(~ Treatment, ncol = 1) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(s_long)

ggsave(
  filename = "secondary_longitudinal_hypo.png", 
  plot = s_long_hypo, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Hyper Models
```{r}
secondary_hyper <- glm(HyperAUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

s_hyper_emmeans <- emmeans(secondary_hyper, ~ Group_Combo)

contrast(s_hyper_emmeans, method = secondary_subset, adjust = "fdr")
```

```{r}
s_hyper_mixed <- lmerTest::lmer(HyperAUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hyper_mixed)

s_hypermixed_emmeans <- emmeans(s_hyper_mixed, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(s_hypermixed_emmeans, method = secondary_subset, adjust = "fdr")
```

```{r}
s_hyper_mixed2 <- lmerTest::lmer(HyperAUC ~ Group_Combo + Weight + (1 | Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hyper_mixed2)

s_hypermixed_emmeans2 <- emmeans(s_hyper_mixed2, ~ Group_Combo)

# 2. Run your custom contrasts
s_hyper_constrast <- as.data.frame(contrast(s_hypermixed_emmeans2, method = secondary_subset, adjust = "fdr"))
```

#### Hyper Figure

```{r}
S_hyper_df <- as.data.frame(s_hypermixed_emmeans2)

s_hyper_split <- S_hyper_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

s_hyper_split <- s_hyper_split |> 
  mutate(SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 

# Interaction plot with connecting lines and capped error bars
ggplot(s_hyper_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hyper NTD)",
    color = "Secondary Treatment",
    title = "Interaction Effect on Hyper NTD"
  ) +
  theme_bw()
```

```{r}
Sec_hyper_sig_bars <- s_hyper_constrast
Sec_hyper_sig_bars <- Sec_hyper_sig_bars |> 
  mutate(DateType = case_when(
    grepl("_S", contrast) ~ "Secondary",
    grepl("_1", contrast) ~ "Secondary + 1",
    grepl("_2", contrast) ~ "Secondary + 2",
    grepl("_3", contrast) ~ "Secondary + 3",
    grepl("_-1", contrast) ~ "Secondary - 1",
    grepl("_None", contrast) ~ "None",
    TRUE ~ "other"))

Sec_hyper_sig_bars <- Sec_hyper_sig_bars |> 
  mutate(Left = case_when(
    grepl("SvN_", contrast) ~ "NoCo",
    grepl("NvCP_", contrast) ~ "NoCo",
    grepl("CPSvCS_", contrast) ~ "CP+S",
    grepl("CPvCPS_", contrast) ~ "CP",
    grepl("NvCS_", contrast) ~ "NoCo",
    TRUE ~ "other"))

Sec_hyper_sig_bars <- Sec_hyper_sig_bars |> 
  mutate(Right = case_when(
    grepl("SvN_", contrast) ~ "SHAM",
    grepl("NvCP_", contrast) ~ "CP",
    grepl("CPSvCS_", contrast) ~ "CS",
    grepl("CPvCPS_", contrast) ~ "CP+S",
    grepl("NvCS_", contrast) ~ "CS",
    TRUE ~ "other"))

Sec_hyper_sig_bars <- Sec_hyper_sig_bars |> 
  mutate(p.value2 = sprintf("%.2e", p.value),
         annotation = paste("P-Value =", p.value2))

comparison_bars_secondary_hyper <- Sec_hyper_sig_bars 

#no significance so no bars will be in the figure
```

Making bars for my comparison to add in the primary +3 facet
```{r}
comparison_bars_secondary_hyper <- comparison_bars_secondary_hyper |> 
  filter(DateType == "Secondary + 3")

comparison_bars_secondary_hyper <- comparison_bars_secondary_hyper |> 
 mutate(y_position = case_when(
    contrast == "SvN_3" ~ -26,
    contrast == "NvCP_3" ~ -4,
     contrast == "CPSvCS_3" ~ -8,
    contrast == "CPvCPS_3" ~ -14,
     contrast == "NvCS_3" ~ -20))

s_hyper_split <- s_hyper_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

comparison_bars_secondary_hyper <- comparison_bars_secondary_hyper |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

Sec_hyper_sig_bars <- Sec_hyper_sig_bars |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))
```

```{r}
s_hyper_top <- ggplot() +
  geom_jitter(
    data = subset(Secondary_Raw, Treatment == "NoCo"),
    aes(x = DateType, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(s_hyper_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(s_hyper_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(s_hyper_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-28, 44)) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(s_hyper_top)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
s_hyper_bottom <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = Treatment, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = s_hyper_split,
    aes(x = Treatment, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = s_hyper_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3) +
    geom_signif(data = comparison_bars_secondary_hyper,
    aes(xmin = Left, 
      xmax = Right, 
      annotations = "", 
      y_position = y_position),
    manual = TRUE,
    tip_length = -0.03,
    textsize = 3,
    vjust = -0.6) +
  coord_cartesian(ylim = c(-28, 44)) +
  scale_color_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Secondary Treatment", y = NULL, color = "Secondary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(s_hyper_bottom)

secondary_hyper_plot <- (s_hyper_top / p_divider / s_hyper_bottom) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "Hyper TD (Degree Hours)") 

print(secondary_hyper_plot)

ggsave(
  filename = "secondary_HyperNTDplot.png", 
  plot = secondary_hyper_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Longitudinal Hyper Secondary

```{r}
s_long_hyper_mixed <- lmerTest::lmer(HyperAUC ~ DateType  * Treatment + Weight + (1 | Group / Bat_ID),  data = Secondary_Results_filtered)

# Check the summary
summary(s_long_hyper_mixed)

s_long_hyper_emmeans <- emmeans(s_long_hyper_mixed, ~ DateType | Treatment)

summary(s_long_hyper_emmeans)

s_long_hyper_em_df <- as.data.frame(s_long_hyper_emmeans)

s_long__hyper_pval <- as.data.frame(pairs(s_long_hyper_emmeans, adjust = "fdr"))
```

##### Longitudinal Hyper figure

```{r}
s_long_hyper_emmeans <- s_long_hyper_em_df |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")),
         Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")),
         SE_lower = emmean - SE,
         SE_Upper = emmean + SE) 
```


I am going to make the top part of the figure for all 3 groups and it be 3 facets

```{r}
s_long_hyper <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = DateType, y = HyperAUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = s_long_hyper_emmeans,
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = s_long_hyper_emmeans,
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = s_long_hyper_emmeans,
    aes(x = DateType, ymin = SE_lower, ymax = SE_Upper, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  coord_cartesian(ylim = c(-5, 44)) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "DateType", y = "Hyper TD (Degree Hours") +
  theme_bw() +
     facet_wrap(~ Treatment, ncol = 1) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())
print(s_long_hyper)

ggsave(
  filename = "secondary_longitudinal_hyper.png", 
  plot = s_long_hyper, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Torpor effects 
This code is to test is NTD on one day effects NTD on the next day.
```{r}
Exploratory_sec <- Secondary_Results_filtered |> 
  filter(!DateType %in% c("None", "Secondary - 1"))

Exploratory_sec <- Exploratory_sec %>%
  # Sort by Bat_ID first, then numerically by date
  arrange(Bat_ID, Day_Num) %>% 
  # Group by Bat_ID so lag() doesn't accidentally pull data from a different bat
  group_by(Bat_ID) %>% 
  # Create the lag column using the current day's AUC
  mutate(Lagged_NTD = lag(Total_AUC)) %>% 
  # Always ungroup when finished to prevent downstream coding issues
  ungroup()
```

```{r}
lagged_sec <- lmerTest::lmer(
  Total_AUC ~ Lagged_NTD * Treatment + DateType + Weight + (1 | Group / Bat_ID),
  data = Exploratory_sec)

summary(lagged_sec) 

lag_slope_s <- emtrends(
  lagged_sec, 
  ~ Treatment, 
  var = "Lagged_NTD")

# View individual treatment slopes (are they significantly different from 0?)
summary(lag_slope_s, infer = c(TRUE, TRUE))

levels(lag_slope_s) # checking order to make list in the right order
slag_contrast <- list("NoCo vs CS" = c(0, -1, 0, 1, 0),
                      "SHAM vs NoCo" = c(-1, 1, 0, 0, 0),
                      "CS vs CP+S" = c(0, 0, 0, -1, 1),
                      "NoCo vs CP" = c(0, -1, 1, 0, 0))

# Step B: Pairwise contrasts between treatment slopes (post-hoc comparisons)
contrast(lag_slope_s, method = slag_contrast, adjust = "holm")
```

```{r}
#later on get this to work to visualize
#pred_interaction <- predict_response(
#  lagged_sec, 
#  terms = c("Lagged_NTD", "Primary_Treatment")
#)
#
# 2. Plot lines with confidence intervals
#plot(pred_interaction, show_data = TRUE) +
#  labs(
#    x = "Yesterday's AUC (Lagged NTD)",
#    y = "Today's Total AUC",
#    title = "Carryover Effect Moderated by Treatment Group",
#    color = "Treatment"
#  ) +
#  theme_classic()
```


### Basal vs Primary Model

```{r}
TC_Basal <- Analysis_results |> 
  filter(Treatment %in% c("CP", "CP+S"),
         DateType %in% c("Basal TC", "Primary"))
```


```{r}
Basal_mixed <- lmerTest::lmer(Total_AUC ~ DateType + Weight + Group + (1 | Bat_ID),  data = TC_Basal)

# Check the summary
summary(Basal_mixed) #variance for group is 0

basalmixed_emmeans <- emmeans(Basal_mixed, ~ DateType)

summary(basalmixed_emmeans)

pairs(basalmixed_emmeans) # only comparing two groups so need for multiple correction 

basal_emmeans_df <- as.data.frame(basalmixed_emmeans)
basal_pvalues <- as.data.frame(pairs(basalmixed_emmeans))

```


```{r}
# 2. Build the paired trajectory plot
ggplot() +
  # Raw data lines connecting individual bats across time points
  geom_line(
    data = TC_Basal, 
    aes(x = DateType, y = Total_AUC, group = Bat_ID), 
    color = "gray70", alpha = 0.5, size = 0.6
  ) +
  # Raw data points for individual bats
  geom_point(
    data = TC_Basal, 
    aes(x = DateType, y = Total_AUC), 
    color = "gray50", alpha = 0.5, size = 2
  ) +
  
  # Model-estimated population mean line
  geom_line(
    data = basal_emmeans_df, 
    aes(x = DateType, y = emmean, group = 1), 
    color = "#9EB0FFFF", size = 1.3
  ) +
  # Model-estimated mean points
  geom_point(
    data = basal_emmeans_df, 
    aes(x = DateType, y = emmean), 
    color = "#9EB0FFFF", size = 4
  ) +
  # Model 95% Confidence Intervals
  geom_errorbar(
    data = basal_emmeans_df, 
    aes(x = DateType, ymin = lower.CL, ymax = upper.CL), 
    color = "#9EB0FFFF", width = 0.08, size = 1
  ) +
  
  # Axis labels and formatting
  labs(
    x = "Experiment Day",
    y = expression("NTD (Degree Minutes) EMM")) +
  scale_x_discrete(labels = c("Basal TC" = "Basal TC", "Primary" = "Primary TC")) +
  theme_classic(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text = element_text(color = "black")
  ) -> b_v_p 


ggsave(
  filename = "EPFU2025_BasalvPrimary.png", 
  plot = b_v_p, 
  width = 5,  # Make narrower (inches)
  height = 5,  # Make taller  (inches)
  dpi = 300)
```

#Primary effect secondary

```{r}
P_vs_S <- Analysis_results |> 
  filter(DateType %in% c("Primary", "Secondary")) |> 
  group_by(Bat_ID) |> 
  filter(n() == 2) |> 
  ungroup()

P_vs_S <- P_vs_S %>%
  # Sort by Bat_ID first, then numerically by date
  arrange(Bat_ID, Day_Num) %>% 
  # Group by Bat_ID so lag() doesn't accidentally pull data from a different bat
  group_by(Bat_ID) %>% 
  # Create the lag column using the current day's AUC
  mutate(Lagged_NTD = lag(Total_AUC)) %>% 
  # Always ungroup when finished to prevent downstream coding issues
  ungroup()
```

```{r}
lag_ps <- lm(
  Total_AUC ~ Lagged_NTD * Treatment + Weight + factor(Group),
  data = P_vs_S)

summary(lag_ps)

lag_slope_ps <- emtrends(
  lag_ps, 
  ~ Treatment, 
  var = "Lagged_NTD")

# View individual treatment slopes (are they significantly different from 0?)
summary(lag_slope_ps, infer = c(TRUE, TRUE))

levels(lag_slope_ps) # checking order to make list in the right order
slag_contrast <- list("NoCo vs CS" = c(0, -1, 0, 1, 0),
                      "SHAM vs NoCo" = c(-1, 1, 0, 0, 0),
                      "CS vs CP+S" = c(0, 0, 0, -1, 1),
                      "NoCo vs CP" = c(0, -1, 1, 0, 0))

# Step B: Pairwise contrasts between treatment slopes (post-hoc comparisons)
contrast(lag_slope_ps, method = slag_contrast, adjust = "holm")
```


## Weight Models

I am interested in if there was a difference in weight change between groups. So I am going to make a column that is delta Weight. 

```{r}
Analysis_results <- Analysis_results |> 
  arrange(Bat_ID, Day_Num) |> 
  group_by(Bat_ID) |> 
  mutate(Day_gap = lead(Day_Num) - Day_Num,
         weight_diff = if_else(Day_gap == 1,
                               lead(Weight) - Weight,
                               NA_real_)) |> 
  ungroup()

Weight_analysis <- Analysis_results |> 
  filter(ExperimentDay %in% c("D0", "D1", "D2", "D3", "D21", "D22", "D23", "D24")) |> mutate(Phase = case_when(ExperimentDay %in% c("D0", "D1", "D2", "D3") ~ "Primary",
                        ExperimentDay %in% c("D21", "D22", "D23", "D24") ~ "Secondary",
                        TRUE ~ NA_character_),
      Post_inject = case_when(
      ExperimentDay %in% c("D0", "D21") ~ 0,
      ExperimentDay %in% c("D1", "D22") ~ 1,
      ExperimentDay %in% c("D2", "D23") ~ 2,
      ExperimentDay %in% c("D3", "D24") ~ 3,
      TRUE ~ NA_real_ ))
```

```{r}
weight_model1 <- lmerTest::lmer(weight_diff ~ Treatment * HypoAUC + Treatment * HyperAUC + (1| Bat_ID), data = Weight_analysis)

summary(weight_model1)
```
Variance from bat_id is 0, so once I have accounted for treatment and hyper and hypo auc there is no difference left for bat_id to account for. Going to run as a regular lm first 

```{r}
weight_model_lm <- lm(
  weight_diff ~ Treatment * HyperAUC + Treatment * HypoAUC + ExperimentDay,
  data = Weight_analysis)

summary(weight_model_lm)
```
I ran these results through gemini and because my experiment days are different from each other, it recommended I add a days post injection and phase (primary or secondary). I'll run it again with this.. not sure if that is really allowed. 

```{r}
weight_model_lm2 <- lm(
  weight_diff ~ Treatment * HyperAUC + Treatment * HypoAUC + Phase + Post_inject,
  data = Weight_analysis)

summary(weight_model_lm2)
```
So there is no significant effect on weight including treatment. I originally asked gemini to include treatment but I'm not sure this was the best thing to do since I had bats that went hot even in cold chambers. However, dropping treatment groups sham bats with all other immunized bats so... I'll talk to Ken about this. 

```{r}
weight_model_lm3 <- lm(
  weight_diff ~ HyperAUC + HypoAUC + Phase + Post_inject,
  data = Weight_analysis)

summary(weight_model_lm3)
```


```{r}
Primary_Secondary_hist <- Analysis_results |> 
  filter(DateType %in% c("Primary", "Primary + 1", "Primary + 2", "Primary + 3", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")) |> 
  mutate(DateType = fct_relevel(DateType, "Primary", "Secondary", "Primary + 1", "Secondary + 1", "Primary + 2", "Secondary + 2", "Primary + 3",  "Secondary + 3")) |> 
  mutate(Treatment = fct_relevel(Treatment, "CP", "CP+S", "CS", "NoCo", "SHAM"))

Hist <- ggplot(Primary_Secondary_hist, aes(x= Total_AUC, fill = Treatment)) +
  geom_histogram(binwidth = 200) +
  facet_wrap(~DateType, ncol = 2) +
  labs(x ="Daily AUC", y = "Count") +
  scale_fill_manual(values = treatmentcolors) +
  theme(text = element_text(size = 18), panel.spacing.x = unit(2, "lines"))


ggsave("EPFU_Histogram.jpeg", plot = Hist, device = jpeg(), width = 8, height = 6.67)

```

## Traces and Visualizations 

```{r}
Metadata <- read.delim("EPFU2025Metadata.tsv", # read in metadata
     na = "NA") 

Metadata <- Metadata |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id)) |> 
    mutate(Treatment = fct_relevel(Treatment, "CP", "CP+S", "CS", "NoCo", "SHAM"))

Metadata2 <- Metadata |> 
  left_join(
    Analysis_results |> select(Bat_ID, Tnorm) |> distinct(), 
    by = "Bat_ID")

Results <- read.delim("EPFU2025FullResults.tsv", # read in results
     na = "NA") 

Results <- Results %>% 
  filter(valid_counts >= 137) |> 
  group_by(Treatment) %>%
  mutate(Treatment_Tnorm = mean(Tnorm, na.rm = TRUE)) |> 
  ungroup()
```


#### None Traces

```{r}
None_data <- Results |> 
  filter(DateType == "None") |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

none_met <- data.table::data.table(Metadata2, key = "id")
none_data2 <- data.table::data.table(None_data, key = "id")

 none_behavr <- behavr(none_data2, none_met)
 
 nonestart <- min(none_behavr$t)
  noneend <- max(none_behavr$t)


ggetho(
  none_behavr, 
  aes(y = Temperature, colour = Treatment, fill = Treatment), 
  time_wrap = 86400,
  time_offset = 50400 # 14:00 offset
) +
  geom_hline(
    aes(yintercept = Tnorm), 
    linetype = "dashed", 
    color = "black", 
    linewidth = 0.8
  ) +
  stat_pop_etho(method = mean_cl_boot) + 
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c("black", "white")
  ) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Time (Hours)",
    breaks = seq(-50400, 86400 - 50400, by = 7200),
    labels = function(x) (x + 50400) / 3600,
    expand = c(0, 0)
  ) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 0, vjust = 1),
    text = element_text(size = 18),
    panel.grid = element_blank()
  ) +
  labs(y = "Temp (°C)") -> EPFU_none

print(EPFU_none)


ggsave("EPFU_none.jpeg", plot = EPFU_none, device = jpeg(), width = 7, height = 7)
```


```{r}
none_summary <- None_data %>%
  group_by(Bat_ID, DateType, Treatment) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE)) |> 
  ungroup() 
 

none_summary2 <- none_summary |> 
  group_by(DateType, Treatment) |> 
  summarise(
    AVGMinimum = mean(Minimum , na.rm = TRUE),
    AVGMaximum = mean(Maximum, na.rm = TRUE),
    AVGAverage = mean(Average, na.rm = TRUE),
    sd_min_temp  = sd(Minimum, na.rm = TRUE),
    sd_max_temp  = sd(Maximum, na.rm = TRUE),
    sd_avg_temp  = sd(Average, na.rm = TRUE),
    n_bats       = n(),) |> 
  ungroup() 
```


### Basal Traces

```{r}
Basal_TC_Traces <- read.delim("EPFU2025FullBasal.tsv", # read in results
     na = "NA")

Basal_TC_Names <- unique(Basal_TC_Traces$Bat_ID)
  
 Basal_TC_Traces <- Basal_TC_Traces |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate_shifted) |> 
  mutate(id = as.character(id)) 
 
Basal_Metadata <- Metadata2 |> 
  filter(Bat_ID %in% Basal_TC_Names) |> 
    mutate(id = Bat_ID) |> 
    mutate(id = as.character(id)) 

BTC_summary <- Basal_TC_Traces %>%
  group_by(Bat_ID, DateType, Treatment) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE)) |> 
  ungroup() 
 

BTC_summary2 <- BTC_summary |> 
  group_by(DateType, Treatment) |> 
  summarise(
    AVGMinimum = mean(Minimum , na.rm = TRUE),
    AVGMaximum = mean(Maximum, na.rm = TRUE),
    AVGAverage = mean(Average, na.rm = TRUE),
    sd_min_temp  = sd(Minimum, na.rm = TRUE),
    sd_max_temp  = sd(Maximum, na.rm = TRUE),
    sd_avg_temp  = sd(Average, na.rm = TRUE),
    n_bats       = n(),) |> 
  ungroup() 
```


```{r}
Basal_met <- data.table::data.table(Basal_Metadata, key = "id")
Basal_Traces <- data.table::data.table(Basal_TC_Traces, key = "id")

basal_behavr <- behavr(Basal_Traces, Basal_met)
```

```{r}
t_start <- min(basal_behavr$t)
t_end   <- max(basal_behavr$t)

shade_start <- 1757002201  
shade_end   <- 1757079001

basal_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment, fill = Treatment)) +
  annotate("rect", 
    xmin = shade_start, 
    xmax = shade_end, 
    ymin = -Inf, 
    ymax = Inf,
    fill = "grey90", 
    alpha = 0.5 ) +
  geom_hline(aes(yintercept = Tnorm), 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.8) +
 stat_pop_etho(method = mean_cl_boot) +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")
  ) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Time (hours)",
    # Breaks MUST start at min(t) in the dataset!
    breaks = seq(t_start, t_end, by = 7200),
    # Labels subtract start time and convert seconds to hours
    labels = function(x) (x - t_start) / 3600,
    expand = c(0, 0)
  ) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(legend.position = "none",
        axis.ticks.x = element_line(color = "black", linewidth = 0.8),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, color = "black"),
    text = element_text(size = 18), 
    panel.grid = element_blank()
  ) +
  labs(y = "Temp (°C)") -> EPFU_Basal


print(EPFU_Basal)

ggsave("EPFU_Basal.jpeg", plot = EPFU_Basal, device = jpeg(), width = 10, height = 6.67)
```

## Basal Primary Traces

```{r}
Primary_only <- Results |> 
  filter(DateType == "Primary",
         Treatment %in% c("CP", "CP+S")) |> 
  select(Bat_ID, DateTime, NumberDate, Temperature, DateType) 

Primary_only$NumberDate_shifted <- Primary_only$NumberDate

Primary_only$NumberDate_shifted<- Primary_only$NumberDate_shifted - (4 * 86400) 

Basal_TC_Traces_short <- Basal_TC_Traces |> 
  select(Bat_ID, DateTime, NumberDate_shifted, Temperature, DateType)

BP_traces <- bind_rows(Primary_only, Basal_TC_Traces_short)

 BP_traces <- BP_traces |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate_shifted) |> 
  mutate(id = as.character(id)) 
 
Basal_Metadata <- Metadata2 |> 
  filter(Bat_ID %in% Basal_TC_Names) |> 
    mutate(id = Bat_ID) |> 
    mutate(id = as.character(id)) 
```


```{r}
BP_traces2 <- data.table::data.table(BP_traces, key = "id")

Bp_behavr <- behavr(BP_traces2, Basal_met)
```

```{r}
bp_start <- min(Bp_behavr$t)
bp_end   <- max(Bp_behavr$t)

#shade_start <- 1757002201  
#shade_end   <- 1757079001

#Bp_behavr |>
#  ggetho(aes(x = t, y = Temperature, colour = DateType, fill = DateType)) +
  #annotate("rect", 
   # xmin = shade_start, 
    #xmax = shade_end, 
    #ymin = -Inf, 
  #  ymax = Inf,
  #  fill = "grey90", 
  #  alpha = 0.5 ) +
 # geom_hline(aes(yintercept = Tnorm), 
  #           linetype = "dashed", 
   #          color = "black", 
    #         linewidth = 0.8) +
# stat_pop_etho(method = mean_cl_boot) +
#  stat_ld_annotations(
#    period = 86400,
#    phase = 39600,
#    l_duration = 43200, 
#    ld_colours = c("black", "white")
#  ) +
#  scale_x_continuous(
#    name = "Time (hours)",
    # Breaks MUST start at min(t) in the dataset!
#    breaks = seq(t_start, t_end, by = 7200),
    # Labels subtract start time and convert seconds to hours
#    labels = function(x) (x - t_start) / 3600,
 #   expand = c(0, 0)
  #) +
#  scale_colour_manual("#9EB0FFFF") +
#  scale_fill_manual("#9EB0FFFF") +
#  theme_bw() +
#  theme(legend.position = "none",
#        axis.ticks.x = element_line(color = "black", linewidth = 0.8),
#    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, color = "black"),
#    text = element_text(size = 18), 
#    panel.grid = element_blank()
#  ) +
#  labs(y = "Temp (°C)") -> EPFU_BP


#print(EPFU_BP)

#ggsave("EPFU_Basal.jpeg", plot = EPFU_Basal, device = jpeg(), width = 10, height = 6.67)
```


## Primary Traces

```{r}
Primary_Results_Full <- Results |> 
  filter(DateType %in% c("Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3"))

Primary_Results_Full <- Primary_Results_Full |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

Primary_Results_Full <- Primary_Results_Full |> 
  mutate(Primary_Treatment = case_when(Treatment %in% c("CP+S" , "CP") ~  "Primary_Cold",
         Treatment %in% c("CS" , "NoCo") ~ "Primary_NoCo",
         TRUE ~ "SHAM"))

 Primary_met <- data.table::data.table(Metadata2, key = "id")
 Primary_results2 <- data.table::data.table(Primary_Results_Full, key = "id")

 primary_behavr <- behavr(Primary_results2, Primary_met) 
 summary(primary_behavr)
 
 glimpse(primary_behavr)
 
 primary_summary <- Primary_Results_Full %>%
  group_by(Bat_ID, DateType, Primary_Treatment) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE)) |> 
  ungroup() 
 

summary_primary <- primary_summary |> 
  group_by(DateType, Primary_Treatment) |> 
  summarise(
    AVGMinimum = mean(Minimum , na.rm = TRUE),
    AVGMaximum = mean(Maximum, na.rm = TRUE),
    AVGAverage = mean(Average, na.rm = TRUE),
    sd_min_temp  = sd(Minimum, na.rm = TRUE),
    sd_max_temp  = sd(Maximum, na.rm = TRUE),
    sd_avg_temp  = sd(Average, na.rm = TRUE),
    n_bats       = n(),) |> 
  ungroup() 

write.csv(summary_primary, "EPFU2025_primary_summary.csv")



Primary_Results_Full |> 
  group_by(Primary_Treatment) |> 
  summarise(
    primarygroup_tnorm = mean(Temp_norm),
    sd_tnorm  = sd(Temp_norm))
```

### Primary Behavr 

```{r}
shaded <- data.frame(Treatment = c("CP", "CP+S"), xmin = 1757344201 , 
  xmax = 1757422801,#adds shaded regions for both the basal and primary thermal #challenges
                     label = c("Primary TC"))

day_ticks <- primary_behavr[!duplicated(primary_behavr$ExperimentDay), ]

primary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment, fill = Treatment)) +
  
  # 1. Shaded region
  geom_rect(data = shaded, 
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
            
  # 2. Baseline line (Moved colour OUTSIDE aes)
  geom_hline(aes(yintercept = Tnorm), 
             colour = "black",        # Static color, won't show in legend
             linetype = "dashed", 
             linewidth = 0.8) +
             
  stat_pop_etho(method = mean_cl_boot) +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Experiment Day",
    breaks = day_ticks$t,             
    labels = day_ticks$DateType,
    expand = c(0,0) ) +
  # 3. Use your original treatmentcolors without the "Tnorm" entry
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(legend.position = "none",
    axis.text.x = element_text(angle = -25, vjust = 1, hjust = 0),
    text = element_text(size = 16),
    panel.grid = element_blank()) +
  labs(y = "Skin Temperature (°C)", colour = "Treatment", fill = "Treatment") -> EPFU_Primary

print(EPFU_Primary)
# dropped rows from figure have NAs
 
ggsave("EPFU_Primary.jpeg", plot = EPFU_Primary, device = jpeg(), width = 10, height = 6.67)

```

## Secondary Traces

```{r}
Secondary_Results <- Results |> 
  filter(DateType %in% c("Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3"))

Secondary_Results <- Secondary_Results |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 Secondary_met <- data.table::data.table(Metadata2, key = "id")
 Secondary_results2 <- data.table::data.table(Secondary_Results, key = "id")

 secondary_behavr <- behavr(Secondary_results2, Secondary_met)
 summary(secondary_behavr)
 
 secondary_summary <- Secondary_Results %>%
  group_by(Bat_ID, DateType, Treatment) |> 
  summarise(
    Minimum = min(Temperature, na.rm = TRUE),
    Maximum = max(Temperature, na.rm = TRUE),
    Average = mean(Temperature, na.rm = TRUE)) |> 
  ungroup() 
 

summary_secondary <- secondary_summary |> 
  group_by(DateType, Treatment) |> 
  summarise(
    AVGMinimum = mean(Minimum , na.rm = TRUE),
    AVGMaximum = mean(Maximum, na.rm = TRUE),
    AVGAverage = mean(Average, na.rm = TRUE),
    sd_min_temp  = sd(Minimum, na.rm = TRUE),
    sd_max_temp  = sd(Maximum, na.rm = TRUE),
    sd_avg_temp  = sd(Average, na.rm = TRUE),
    n_bats       = n(),) |> 
  ungroup() 

write.csv(summary_secondary, "EPFU2025_secondary_summary.csv")
```

```{r}
secondary_etho  <- ggetho(secondary_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho() +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Temp (°C)",
       y = "Bat ID")

ggsave("EPFU_Sec_Etho.jpeg", plot = secondary_etho, device = jpeg(), width = 12, height = 6.67)
```

```{r}
ggetho(secondary_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r}
shaded2 <- data.frame(Treatment = c("CS", "CP+S"), xmin = 1759155601 , 
  xmax = 1759234201,#adds shaded regions for both the basal and primary thermal #challenges
  label = c("Secondary TC"))

s_day_ticks <- secondary_behavr[!duplicated(secondary_behavr$ExperimentDay), ]

secondary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
   geom_hline(aes(yintercept = Tnorm), 
             colour = "black",        # Static color, won't show in legend
             linetype = "dashed", 
             linewidth = 0.8) +
  geom_rect(data = shaded2, 
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
  stat_pop_etho(method = mean_cl_boot) +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Experiment Day",
    breaks = s_day_ticks$t,             
    labels = s_day_ticks$DateType,
    expand = c(0,0)) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(legend.position = "null",
    axis.text.x = element_text(angle = -25, vjust = 1, hjust = 0)) +
  theme(text = element_text(size = 16), panel.grid = element_blank())+
  labs(y = "Skin Temperature (°C)") -> EPFU_Secondary

print(EPFU_Secondary)
 
ggsave("EPFU_Secondary.jpeg", plot = EPFU_Secondary, device = jpeg(), width = 10, height = 6.67)
```

