---
title: "MOCO 2025 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read.delim("MOCO2025TempLoggerMetadata.tsv", # read in metadata for ibuttons
     na = "NA")

glimpse(Metadata) # check formatting

Metadata$Bat_ID = as.character(Metadata$Bat_ID) # change Bat_ID to a character

glimpse(Metadata) # check formatting again
```

```{r}
G1_Ambient <- read.delim("MOCO2025_G1_Ambient.tsv", # read in ambient temp fo G1
     na = "NA")
G2_Ambient <- read.delim("MOCO2025_G2_Ambient.tsv", # read in ambient temp fo G2
     na = "NA")
G3_Ambient <- read.delim("MOCO2025_G3_Ambient.tsv", # read in ambient temp fo G3
     na = "NA")
```


```{r}
G1_Ambient |> 
  summarise(Avg = mean(Value),
            Min = min(Value),
            Max = max(Value))
```

### Shifting Data

We had two cohorts of bats for non thermally challenged bats. They followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. Because of this, ambient temperature is different for group 1 D21 bats and group 2 D21 bats. 

```{r}
SecondaryBatNames <- unique(Secondary_results$Bat_ID)

Secondary_Met <- no_hot_met %>%
  filter(Bat_ID %in% SecondaryBatNames) |> 
  select(! "Logger location") |> 
  select(! "Notes")

summary(as.factor(Secondary_Met$Treatment))
# shows 6 TC bats but I thought we had 7.. look into this.

SecondaryResultsMet <- Secondary_results |> 
left_join(Secondary_Met, by = join_by(Bat_ID))

SecondaryResultsMet %>%
  group_by(Bat_ID) %>%
  summarize(Min_DateTime = min(DateTime), Max_DateTime = max(DateTime))
```

```{r}
SecondaryResultsMet$time_aligned <- SecondaryResultsMet$DateTime

md.pattern(SecondaryResultsMet, rotate.names = TRUE)

SecondaryResultsMet <- SecondaryResultsMet |> 
  select(! "Serial Number") |> 
  select(! "Logger Name") |> 
  select(! "Filename")

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] - lubridate::days(2)

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] - lubridate::days(11)

boostresultsmet$shift_time_num <-SecondaryResultsMetboostresultsmet$shift_time_num <- boostresultsmet$time_aligned

boostresultsmet$shift_time_num <- as.numeric(boostresultsmet$shift_time_num)

str(boostresultsmet)

boostresultsmet |> 
  group_by(Bat_ID) |> 
  summarize(Min_date = min(DateTime), Min_shift = min(time_aligned),
            min_daten =min(NumberDate, Min_shiftn = min(shift_time_num)))

boosted <- boostresultsmet %>%
  select(id = Bat_ID, t = shift_time_num, Temperature, )
head(boosted)

boosted %>%
  group_by(id) %>%
  summarize(Min_DateTime = min(t), Max_DateTime = max(t))

```

I shifted the data so now data all shows d21 on June 5th, the same day as D21 for Group 1. 


##Cold bats initial

```{r}
 Primary_Temp <- clean_results |> 
   filter(Bat_ID %in% c("64.1", "68.1", "74.1"))  |> 
  mutate(t = NumberDate) |>  
mutate(id = Bat_ID)  
  

Primary_Results <- Primary_Temp |> 
  filter(DateTime <= as.POSIXct("2025-06-01 00:00:00", tz = "UTC"))
   
 
 PrimaryBatNames <- unique(Primary_Results$id)
 
 Cold_met <- clean_met |> 
   filter(Bat_ID %in% PrimaryBatNames) |> 
   mutate(id = Bat_ID)
  
 Cold_data <- Cold_met %>%
   select(Bat_ID) %>%
   left_join(Primary_Results, by = join_by(Bat_ID))
 str(Cold_data)

 cold_met2 <- data.table::data.table(Cold_met, key = "id")
 cold_data2 <- data.table::data.table(Cold_data, key = "id")

 cold_behavr <- behavr(cold_data2, cold_met2)
 summary(cold_behavr)
```

### Cold Bat Behavr 

```{r}

 ticks_7am <- seq(
   from = floor_date(as.POSIXct(min(cold_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
   to   = ceiling_date(as.POSIXct(max(cold_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
   by   = "1 day"
 )
 
 MOCO_Cold <- cold_behavr |>
   ggetho(aes(x = t, y = Temperature, colour = Bat_ID)) +
   stat_pop_etho() +
   # Use phase instead of offset to shift the LD bar
   stat_ld_annotations(
     ld_colours = c("white", "black"),
     period = hours(24), 
     phase = hours(7) 
   ) +
   facet_grid(id ~ .) +
   # To avoid the 'Scale for x is already present' warning, 
   # we modify the existing scale rather than adding a new one.
   # Alternatively, keep your scale but ignore the warning, 
   # or define time settings inside ggetho()
   scale_x_continuous(
     name = "Date/time",
     breaks = as.numeric(ticks_7am),
     labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")
   )
 MOCO_Cold <- cold_behavr |>
   ggplot(aes(x = t, y = Temperature, colour = id)) +
   stat_pop_etho() +
   stat_ld_annotations(
     ld_colours = c("white", "black"),
     period = 86400,
     phase = 7 * 3600
   ) +
   facet_grid(id ~ .) +
   scale_x_continuous(
     name = "Date/time",
     breaks = as.numeric(ticks_7am),
     labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")
   )
 
print(MOCO_Cold)
 
cold_behavr |>
   ggetho(aes(x = t, y = Temperature, colour = Bat_ID)) +
   stat_pop_etho() +
   stat_ld_annotations(
     ld_colours = c("white", "black"),
        # 12h light : 12h dark, in seconds
     offset = 7 * 3600) +           # <<< shift start by 7 hours (7 AM)
   facet_grid(id ~ .) +
     scale_x_continuous(
     name = "Date/time",
     breaks = as.numeric(ticks_7am),
     labels = function(x)
       format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) 
 
  ggsave("MOCO_Cold.jpeg", plot = MOCO_Cold, device = jpeg(), width = 12, height = 6.67)
  
#  
#  cold_behavr |>
#   ggetho(aes(x = t, y = Temperature, colour = id)) +
#   stat_pop_etho() +
#   stat_ld_annotations(
#     ld_colours = c( "black", "white")) +
#   facet_grid(id ~ .) +
#     scale_x_continuous(
#     name = "Date/time",
#     breaks = as.numeric(ticks_7am),
#     labels = function(x)
#       format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) +  theme(
#     axis.text = element_text(size = 14),       # axis tick labels
#     axis.title = element_text(size = 16),      # axis titles
#     legend.text = element_text(size = 13),     # legend labels
#     legend.title = element_text(size = 15),    # legend title
#     strip.text = element_text(size = 12)       # facet labels
#   )  -> MOCO_Cold
# 
# ggsave("MOCO_Cold.jpeg", plot = MOCO_Cold, device = jpeg(), width = 12, height = 6.67)
```

## Behavr Analysis

```{r}

BoostBatNames <- unique(boosted$id)

Boost_met <- clean_met %>%
  rename(id = Bat_ID) %>%
  filter(id %in% BoostBatNames)


Boost_data <- Boost_met %>%
  select(id) %>%
  left_join(boosted, by = join_by(id))
str(Boost_data)

boost_met2 <- data.table::data.table(Boost_met, key = "id")
boost_data2 <- data.table::data.table(Boost_data, key = "id")

boost_behavr <- behavr(boost_data2, boost_met2)
summary(boost_behavr)

```

```{r}
ggetho(boost_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho()
```

```{r}
ggetho(boost_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r, warning=FALSE}

ticks_7am <- seq(
  from = floor_date(as.POSIXct(min(boost_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
  to   = ceiling_date(as.POSIXct(max(boost_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
  by   = "1 day"
)

boost_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
  stat_pop_etho() +
  stat_ld_annotations(
    ld_colours = c( "white", "black"),
    ratio = c(12, 12)* 3600) +
  facet_grid(Treatment ~ .) +
    scale_x_continuous(
    name = "Date/time",
    labels = function(x)
      format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M"))


cols <- c(
  "eVLP alone" = "#F8776D",
  "eVLP+cGAMP" = "#ABA300",
  "eVLP+LPS"    = "#00C19A",
  "eVLP+POLY I:C" = "#00A9FF",
  "SHAM"        = "#ED68ED",
  "TC"          = "#E68613")

boost_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
  stat_pop_etho() +
  stat_ld_annotations(
    ld_colours = c("white", "black"),
       # 12h light : 12h dark, in seconds
    offset = 7 * 3600) +           # <<< shift start by 7 hours (7 AM)
  facet_grid(Treatment ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = as.numeric(ticks_7am),
    labels = function(x)
      format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) +
   scale_colour_manual(name = "Treatment", values = cols) +
  scale_fill_manual(name = "Treatment", values = cols) +
  theme(
    axis.text = element_text(size = 14),       # axis tick labels
    axis.title = element_text(size = 16),      # axis titles
    legend.text = element_text(size = 13),     # legend labels
    legend.title = element_text(size = 15),    # legend title
    strip.text = element_text(size = 12)       # facet labels
  ) 

unique(boost_behavr$Treatment)

# ggsave("MOCO_Boost.jpeg", plot = MOCO_Boost, device = jpeg(), width = 16.4, height = 9.5)

```

```{r, warning=FALSE}
#ggetho(bat_behavr, aes(y=Temperature, colour=Treatment), time_wrap = hours(24)) +
#      stat_pop_etho() +
#      stat_ld_annotations() +
#      facet_grid(Treatment ~ .) +
#      scale_y_continuous(name= "Temperature")
```

```{r Sham, warning=FALSE}
bat_met_sham <- bat_met2 %>%
  filter(Treatment == "SHAM")
bat_data_sham <- bat_data2 %>%
  filter(id %in% bat_met_sham$id)
bat_behavr_sham <- behavr(bat_data_sham, bat_met_sham)

bat_behavr_sham %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP, warning=FALSE}
bat_met_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP")
bat_data_eVLP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP$id)
bat_behavr_eVLP <- behavr(bat_data_eVLP, bat_met_eVLP)

bat_behavr_eVLP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_LPS, warning=FALSE}
bat_met_eVLP_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS")
bat_data_eVLP_LPS <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_LPS$id)
bat_behavr_eVLP_LPS <- behavr(bat_data_eVLP_LPS, bat_met_eVLP_LPS)

bat_behavr_eVLP_LPS %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_POLY, warning=FALSE}
bat_met_eVLP_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY")
bat_data_eVLP_POLY <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_POLY$id)
bat_behavr_eVLP_POLY <- behavr(bat_data_eVLP_POLY, bat_met_eVLP_POLY)

bat_behavr_eVLP_POLY %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_cGAMP, warning=FALSE}
bat_met_eVLP_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP")
bat_data_eVLP_cGAMP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_cGAMP$id)
bat_behavr_eVLP_cGAMP <- behavr(bat_data_eVLP_cGAMP, bat_met_eVLP_cGAMP)

bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r, warning=FALSE}
ggetho(bat_behavr, aes(y=Temperature, colour=Treatment)) + 
  stat_pop_etho()
```

```{r, warning=FALSE}
bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(x=t, z=Temperature), multiplot = 2) +
      stat_tile_etho() +
      facet_wrap( ~ id)
```

## Subsetting the data to pre-boost and post-boost

Before I proceed with subsetting, I am going to make sure that I understand the data.

```{r}
allbatresults3 %>%
  group_by(batnames) %>%
  summarize(Min_timepoint = min(Timepoint), Min_DateTime = min(DateTime), Max_DateTime = max(DateTime),
            Min_Temp = min(Temperature), Max_Temp = max(Temperature))
```

We can tell that the loggers on 2023-04-16 were all programmed uniformly and collected 42 timepoints before the new start time of midnight on 2023-04-16. 
While the 2023-04-16 loggers have two different programming regimens. Some collected 186 and others collected 330 earlier datapoints. 

Bottom line, there are three different starting timepoints: 43, 187, and 331.
I will substract the appropriate value from the Timepoint column after I separate the data.


```{r}
left_trimmed_results %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint))
allbatresults %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint), maxTimepoint = max(Timepoint), 
            minTemp = min(Temperature), maxTemp = max(Temperature)) %>%
  arrange(maxTimepoint)
```


There are three different batches of treatment, as expected:
2023-03-27 2023-03-29 2023-03-31 were the initial immunization dates (prime)
Day 21 for each batch was, therefore:
2023-04-17 2023-04-19 2023-04-21

```{r}
summary(as.factor(bat_met2$Treatment_date))
first_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-27"))
second_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-29"))
third_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-31"))

first_results <- allbatresults3 %>%
  filter(batnames %in% first_treatment$id)
second_results <- allbatresults3 %>%
  filter(batnames %in% second_treatment$id)
third_results <- allbatresults3 %>%
  filter(batnames %in% third_treatment$id)

first_results_trim <- first_results %>%
  filter(DateTime <= ymd("2023-04-17")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_trim <- second_results %>%
  filter(DateTime <= ymd("2023-04-19")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_trim <- third_results %>%
  filter(DateTime <= ymd("2023-04-21")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

pretreatment_results <- bind_rows(first_results_trim, second_results_trim) %>%
  bind_rows(third_results_trim)

select_bat_results <- pretreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(select_bat_results)
  
bat_data_preboost <- clean_met3 %>%
  select(id) %>%
  left_join(select_bat_results, by = join_by(id))
str(bat_data_preboost)

bat_data_preboost2 <- data.table::data.table(bat_data_preboost, key = "id")

bat_behavr_preboost <- behavr(bat_data_preboost2, bat_met2)
```

Note that I think that the last day shown on these plots is the booster immunization day.

```{r, warning=FALSE}
# summary(bat_behavr_preboost)
# bat_behavr_preboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_preboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

And now for the post-boost results. 
For starters, I will trim the 4 days after the injections.
That didn't work because ggetho kept saying that there was less than 3 days of data :/

Secondly just kept the above format but used more data.

```{r}
# first_results_post <- first_results %>%
#   filter(DateTime >= ymd("2023-04-17")) %>%
#   filter(DateTime <= ymd("2023-04-22")) %>%
#   mutate(Timepoint = Timepoint - 474) ## My attempt to standardize the start time. 
# second_results_post <- second_results %>%
#   filter(DateTime >= ymd("2023-04-19")) %>%
#   filter(DateTime <= ymd("2023-04-24")) %>%
#   mutate(Timepoint = Timepoint - 762)
# third_results_post <- third_results %>%
#   filter(DateTime >= ymd("2023-04-21")) %>%
#   filter(DateTime <= ymd("2023-04-26")) %>%
#   mutate(Timepoint = Timepoint - 906)

first_results_post <- first_results %>%
  filter(DateTime <= ymd("2023-04-22")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_post <- second_results %>%
  filter(DateTime <= ymd("2023-04-24")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_post <- third_results %>%
  filter(DateTime <= ymd("2023-04-26")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

posttreatment_results <- bind_rows(first_results_post, second_results_post) %>%
  bind_rows(third_results_post)

post_bat_results <- posttreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(post_bat_results)
  
bat_data_postboost <- clean_met3 %>%
  select(id) %>%
  left_join(post_bat_results, by = join_by(id))
str(bat_data_postboost)

bat_data_postboost2 <- data.table::data.table(bat_data_postboost, key = "id")

bat_behavr_postboost <- behavr(bat_data_postboost2, bat_met2)
summary(bat_behavr_postboost)
```
```{r Postboost_Summary, warning=FALSE}
# summary(bat_behavr_postboost)
# bat_behavr_postboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_postboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

So I think that the booster happened on day 4 on these graphs, but I need to verify.

## Booster Analysis

To be certain of using the correct timestamps, I am going to analyze with the openair package, as I did for the squirrels.

This is done one bat at a time...

https://bookdown.org/david_carslaw/openair/sections/trend-analysis/time-variation.html

First, comparing all pre-boost data to (boost day + 1 day following)

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

The results look very similar.


```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

To keep the comparison cleaner, I will continue with pre versus post. 
The only downside is that the confidence interval is a little wider on the "pre" data than the "other" data.

## Comparison of all bats

```{r}
bat_met2 %>%
  arrange(Treatment) %>%
  select(id, Treatment)
```

Let's make a function for the timeVariation plot

```{r}
timeVariation_fun <- function(arg1) {
  mybatID <- arg1

  
  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(2)

  singlebatresults <- allbatresults3 %>%
    filter(batnames == mybatID)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", 
                           group = "Boosted",  
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

```{r}
timeVariation_fun("1N8476")
```

Now let's look at the groups one at a time.

```{r SHAM}
bat_sham <- bat_met2 %>%
  filter(Treatment == "SHAM") %>%
  select(id) %>%
  pull
lapply(bat_sham, timeVariation_fun)
```

```{r eVLP only}
bat_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP") %>%
  select(id) %>%
  pull
lapply(bat_eVLP, timeVariation_fun)
```


```{r eVLP LPS}
bat_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS") %>%
  select(id) %>%
  pull
lapply(bat_LPS, timeVariation_fun)
```

```{r eVLP POLY}
bat_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY") %>%
  select(id) %>%
  pull
lapply(bat_POLY, timeVariation_fun)
```

```{r eVLP cGAMP}
bat_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP") %>%
  select(id) %>%
  pull
lapply(bat_cGAMP, timeVariation_fun)
```

## Tent Logger Data

Two possible ways to incorporate this data. 
First, I will add it as a separate "pollutant".
Second, I will show it as a separate "feature".

```{r}
alltentresults <- fread("TemperatureLoggers/tentloggerdata_long.tsv")
alltentresults <- alltentresults %>%
  mutate(Tent = gsub("Tent ", "", Tent))
str(alltentresults)

tent_met <- Metadata %>%
  drop_na(Filename) %>%
  select(Filename, DMR_ID, Tag, Treatment, Age, Sex,
         Attachment_date = "Date Attached", 
         Treatment_date = Day_00, 
         RIP_date,
         Tent = Tent.x) %>%
  mutate(id = gsub(".DAT", "", Filename)) %>%
  relocate(id)

tentbatresults <- tent_met %>%
  select(id, Tent) %>%
  right_join(allbatresults3, by = join_by(id == batnames))

tentbatresults2 <- tentbatresults %>%
  left_join(alltentresults, by = join_by(Tent, DateTime)) %>%
  rename(`Bat T` = Temperature.x) %>%
  rename(`Tent T` = Temperature.y)

write_csv(tentbatresults2, file = "allbatandtentresults.csv")
```

A new plot function:

```{r}
timeVariation_fun2 <- function(arg1) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  # Immunization <- bat_met2 %>%
  #   filter(id == mybatID) %>%
  #   select(Treatment_date) %>%
  #   pull
  # Boost <- Immunization + days(21)
  # BoostWindow <- days(2)

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID)

  # singlebatresults2 <- singlebatresults %>%
  #   filter(DateTime <= Boost + BoostWindow) %>%
  #   mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  #   mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  #   rename(date = DateTime)
  
  singlebatresults2 <- singlebatresults %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant =c("Tent T", "Bat T"), 
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun2("1N8476")
```
That looks good. Let's plot each group at a time.

```{r Sham-Tent}
lapply(bat_sham, timeVariation_fun2)
```

```{r eVLP-Tent}
lapply(bat_eVLP, timeVariation_fun2)
```

```{r LPS-Tent}
lapply(bat_LPS, timeVariation_fun2)
```

```{r POLY-Tent}
lapply(bat_POLY, timeVariation_fun2)
```

```{r cGAMP-Tent}
lapply(bat_cGAMP, timeVariation_fun2)
```

Okay, because the "difference" function worked fine, I don't think that I need to reorganize the data. 
Let's see if I can also use the boost "group" at the same time.

Dammit: "Can only have one pollutant with a grouping variable, or several pollutants and no grouping variable."

Okay, I think I will make a new variable called delta T and use that as the only pollutant.

I will also test different "postboost" windows.

A new plot function:

```{r}
timeVariation_fun3 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant = "delta T", 
                           group = "Boosted",
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun3("1N8476", 5)
```

```{r Sham-deltaT}
lapply(bat_sham, timeVariation_fun3, 3)
```

```{r eVLP-deltaT}
lapply(bat_eVLP, timeVariation_fun3, 3)
```

```{r LPS-deltaT}
lapply(bat_LPS, timeVariation_fun3, 3)
```

```{r POLY-deltaT}
lapply(bat_POLY, timeVariation_fun3, 3)
```

```{r cGAMP-deltaT}
lapply(bat_cGAMP, timeVariation_fun3, 3)
```

## Comparing by Treatment Group

Using the "group" parameter in timeVariation and to make separate plots for pre-boost and then each day post-boost.


```{r Group Comparisons}
tentbatresults3 <- tentbatresults2 %>%
  mutate(`delta T` = `Bat T` - `Tent T`) %>%
  rename(date = DateTime)

bat_boost_date <- bat_met2 %>%
  select(id, Treatment, Treatment_date) %>%
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP",
                                                  "eVLP LPS", "eVLP POLY", "eVLP cGAMP")))

tentbatresults4 <- tentbatresults3 %>%
  left_join(bat_boost_date, by = join_by(id))

str(tentbatresults4)

preboost_tentbatresults <- tentbatresults4 %>%
  filter(date < (Treatment_date + days(21)))

boostday_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(21)) & date <= (Treatment_date + days(22)))

postboostday1_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(22)) & date <= (Treatment_date + days(23)))

postboostday2_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(23)) & date <= (Treatment_date + days(24)))

postboostday3_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday4_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday5_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(25)) & date <= (Treatment_date + days(26)))


preboost_plot <- timeVariation(preboost_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Preboost", "Preboost", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(preboost_plot, subset="hour")

boostday_plot <- timeVariation(boostday_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Boost Day", "Boost Day", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(boostday_plot, subset="hour")

postboostday1_plot <- timeVariation(postboostday1_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 1", "Postboost Day 1", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday1_plot, subset="hour")

postboostday2_plot <- timeVariation(postboostday2_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 2", "Postboost Day 2", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday2_plot, subset="hour")

postboostday3_plot <- timeVariation(postboostday3_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 3", "Postboost Day 3", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday3_plot, subset="hour")

postboostday4_plot <- timeVariation(postboostday4_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 4", "Postboost Day 4", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday4_plot, subset="hour")

postboostday5_plot <- timeVariation(postboostday5_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 5", "Postboost Day 5", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday5_plot, subset="hour")

```


## Quantifying Temperature Differences

Designing a function to measure the daily difference in Bat temperature compared to the tent temperature.

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
```


A function to quantify this delta T by day:

```{r}
tempDifference_fun <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  DMR_Id <- NewTempLoggerMetadata %>%
    filter(Id == mybatID) %>%
    select(DMR_ID) %>%
    pull
  
  Tent <- Metadata %>%
    filter(DMR_ID == DMR_Id) %>%
    select(Tent.x) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`)/144, 
                      daily_batT = mean(`Bat T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 Tent = Tent,
                                 `DeltaT d-3` = daily_summary$daily_delta[1],
                                 `DeltaT d-2` = daily_summary$daily_delta[2],
                                 `DeltaT d-1` = daily_summary$daily_delta[3],
                                 `DeltaT d0` = daily_summary$daily_delta[4],
                                 `DeltaT d1` = daily_summary$daily_delta[5],
                                 `DeltaT d2` = daily_summary$daily_delta[6],
                                 `BatT d-3` = daily_summary$daily_batT[1],
                                 `BatT d-2` = daily_summary$daily_batT[2],
                                 `BatT d-1` = daily_summary$daily_batT[3],
                                 `BatT d0` = daily_summary$daily_batT[4],
                                 `BatT d1` = daily_summary$daily_batT[5],
                                 `BatT d2` = daily_summary$daily_batT[6])
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
tempDifference_fun("1N8476", 3)
bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
bat_delta <- bind_rows(bat_delta_list)
bat_delta_full <- bat_delta |>
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
  rowwise() |>
  mutate(`Preboost Mean dT` = mean(c(`DeltaT d-3`,`DeltaT d-2`,`DeltaT d-1`))) |>
  mutate(`Postboost Mean dT` = mean(c(`DeltaT d0`,`DeltaT d1`,`DeltaT d2`))) |>
  mutate(`Preboost Bat T` = mean (c(`BatT d-3`,`BatT d-2`,`BatT d-1`))) |>
  mutate(`Postboost Bat T` = mean(c(`BatT d0`,`BatT d1`,`BatT d2`)))
bat_delta_full
```

And let's try to look at the various possible models.

```{r Delta T Models}
first_model <- glm(`Postboost Mean dT` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

```{r Delta T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Mean dT`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day Postboost ',Delta,'T'))) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)
bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Mean dT`:`Postboost Mean dT`, names_to = "Boost")
ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day ',Delta,'T')))

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


Well that is quite disappointing. 
Tent is clearly the biggest factor driving this.
Delta T is not the best way to analyze this phenomenon.

First, we need to look at Bat T instead of delta T.


```{r Bat T Models}
first_model <- glm(`Postboost Bat T` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

This is the ideal situation for a mixed effects model!

Let's see how it works.

Following guidance at https://m-clark.github.io/mixed-models-with-R/random_intercepts.html

```{r}
bat_delta_T <- bat_delta_full |>
  mutate(`Boost dT` = `Postboost Bat T` - `Preboost Bat T`)

mixed_model <- lmer(`Boost dT` ~ Treatment + (1|Tent), data = bat_delta_T)
summary(mixed_model)
confint(mixed_model)
ranef(mixed_model)$Tent
coef(mixed_model)$Tent

predictInterval(mixed_model)   # for various model predictions, possibly with new data

REsim(mixed_model)             # mean, median and sd of the random effect estimates

plotREsim(REsim(mixed_model))  # plot the interval estimates

```

It looks like even after incorporating Tent as a random effect, the Treatment effect is not significant.
This may be because I am looking at the 3-day average and I need a more focused approach.

```{r Bat T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Bat T`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)

bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Bat T`:`Postboost Bat T`, names_to = "Boost") |>
  mutate(Boost = factor(Boost, levels = c("Preboost Bat T", "Postboost Bat T")))

ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_pivot, aes(x = Boost, y = value)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5, position = position_nudge()) +
  geom_line(aes(group=BatID), colour="red", linetype="11") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Treatment) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


### Day 1 Post-boost

Reviewing some of the data from the sham group, I think it would be better to look only at post-boost day 1 or day 2.
This should avoid some of the impact of handling. 

```{r}
bat_delta_T_day1 <- bat_delta_full |>
  mutate(`Boost dT day1` = `BatT d1` - `Preboost Bat T`) |>
  mutate(`Boost dT day2` = `BatT d2` - `Preboost Bat T`)

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d1`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d2`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

mixed_model_d1 <- lmer(`Boost dT day1` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d1)
confint(mixed_model_d1)
ranef(mixed_model_d1)$Tent
coef(mixed_model_d1)$Tent
predictInterval(mixed_model_d1)   # for various model predictions, possibly with new data
REsim(mixed_model_d1)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d1))  # plot the interval estimates

mixed_model_d2 <- lmer(`Boost dT day2` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d2)
confint(mixed_model_d2)
ranef(mixed_model_d2)$Tent
coef(mixed_model_d2)$Tent
predictInterval(mixed_model_d2)   # for various model predictions, possibly with new data
REsim(mixed_model_d2)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d2))  # plot the interval estimates
```

The d2 model might show something interesting for the Poly(I:C) group, but it is a small effect.


## Hourly Temperature Differences

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults3 <- singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                    .by = "hour", 
                    hourly_delta = sum(`delta T`)/6)


ggplot(singlebatresults3, aes(x = date, y = hourly_delta, color = Boosted)) + 
  geom_line() +
  theme_cowplot()
```
A function to quantify this delta T by hour:

```{r}
tempDifference_fun2 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 `Preboost d-3` = daily_summary$daily_delta[1]/144,
                                 `Preboost d-2` = daily_summary$daily_delta[2]/144,
                                 `Preboost d-1` = daily_summary$daily_delta[3]/144,
                                 `Boost d0` = daily_summary$daily_delta[4]/144,
                                 `Boost d1` = daily_summary$daily_delta[5]/144,
                                 `Boost d2` = daily_summary$daily_delta[6]/144)
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
# tempDifference_fun2("1N8476", 3)
# bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
# bat_delta <- bind_rows(bat_delta_list)
# bat_delta_full <- bat_delta |>
#   mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
#   rowwise() |>
#   mutate(`Preboost Mean dT` = mean(c(`Preboost d-3`,`Preboost d-2`,`Preboost d-1`))) |>
#   mutate(`Postboost Mean dT` = mean(c(`Boost d0`,`Boost d1`,`Boost d2`)))
# bat_delta_full
```
