---
title: "MOCO 2025 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to look at each logger individually to verify that the data is clean.

Then I will plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read_excel("ibutton_assignments.xlsx", # read in metadata for ibuttons
    range = "A1:P59", na = "NA")

glimpse(Metadata) # check formatting

Metadata$Bat_ID = as.character(Metadata$Bat_ID) # change Bat_ID to a character

glimpse(Metadata) # check formatting again
```


```{r}
DATA_DIR <- "./MOCO Data/Edited Data" # load in each logger file

file_list <- as_tibble(list.files(DATA_DIR)) |>
  mutate(Bat_ID = gsub("MOCO_Band", "", value)) |> # removes everything after bat ID
  mutate(Bat_ID = gsub(".csv", "", Bat_ID)) |>  #removes .csv
  rename(Filename = value) 

Metadata <- left_join(Metadata, file_list, by = join_by(Bat_ID == Bat_ID)) # adds file name into metadata

batfilelist <- list.files(path = "./MOCO Data/Edited Data", 
                          pattern = "*.csv", full.names = T) # shows full file path name
batfilenames <- list.files(path = "./MOCO Data/Edited Data", 
                           pattern = "*.csv", full.names = F) # shows only file name
batfiles <- lapply(batfilelist, fread, 
                   col.names = c("Bat_ID", "DateTime","Unit", "Temperature"), sep = "," , 
                   skip = 20) # applies columns to each file, skips the first 20 rows

batnames <- gsub("MOCO_Band", "", batfilenames) # removes MOCO_Band from file name

batnames = gsub(".csv", "", batnames) # removes .csv from file name and leaves only Band 

names(batfiles) <- batnames # names each file by logger name
```


```{r}
allbatresults <- bind_rows(batfiles, .id = "Bat_ID") # creates a long dataset with each time point for every logger, 4 variables 

allbatresults <- allbatresults %>%
  mutate(DateTime = mdy_hm(DateTime)) # changes date to appropriate format
str(allbatresults) # similar to glimpse
summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
md.pattern(allbatresults) # checking for NA's
Metadata %>%
  filter(!Bat_ID%in% file_list$Bat_ID) %>% # only selects loggers with a matching file name
  drop_na(`Filename`) # removes all na's
Metadata %>%
  filter(is.na(`Filename`)) # shows which bats do not have a logger
```

Check files for bands 79, 80, 68.1, and 74. These loggers have 4,000 datapoints instead of 2,000. Could be because they were used during primary immunization for aim 3 bats. 

The bats with no files represent loggers that failed to download, 

The logger numbers with no bat ID have not been identified

## Missing Data

```{r}
no_hot_met <- Metadata |> 
  filter(!Treatment =="Hot") # remove data from group not used in this study

clean_met <- no_hot_met %>%
  drop_na(Filename) # removes any ibuttons with no file

head(clean_met) # look at the top of the dataset
md.pattern(clean_met, rotate.names = TRUE)
write_tsv(clean_met, file = "MOCO2025TempLoggerMetadata.tsv")
```

I am left with 34 ibuttons with files. There are some NAs in my database but its just from the notes and logger location which I will not be using for the analysis so they shouldn't interfere with the data right now. 

## Making a plot function

```{r}
plot_fun <- function(arg1) {
  allbatresults %>%
    dplyr::filter(Bat_ID == arg1) %>%
    ggplot(aes(x=DateTime, y=Temperature)) +
    geom_line() +
    ylab("Temperature (C)") + xlab("Date") + ylim(4,40) +
    theme_cowplot() 
}
```

```{r}
plotly_fun <- function(arg1) {
  plot_ly(data = allbatresults[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

## Looking at the data

```{r Bat 03}
ggplotly(plot_fun("3"))
```

```{r Plotly Bat 16}
plotly_fun("16")
```

Here we go!

This works fine, but is not interactive

```{r}
lapply(batnames, plotly_fun)
```

If you can't see the whole plot, use the "Show in New Window" icon above the output.

Aim 3 inoculation was on 26-May-2025
Aim 3 boost was on 16-Jun-2025
Group 1 boost was on 5-Jun-2025
Group 2 boost was on 7-Jun-2025

64.2 Data needs to be trimmed, some points dropping to -40. Rest of data looks good
Delete 73.3, logger only stayed on for a few minutes.
76.2 Data needs to be trimmed, some points dropping to -30. Rest of data looks good. 
I think as long as they were on for the 24 hours after the boost we should keep the data
Delete 78.2, logger only stayed on for a few hours. 

Additional things to consider:

39- Logger fell off and reattached 7-Jun
47- Logger fell off and reattached on 6-Jun
48 - Logger fell off and reattached on 7-Jun

## Calibration
Before I go any further, I need to calibrate the data. I am using a 2 point calibration from the low and high calibration temps. The formula I am using is CorrectedValue = (((RawValue – RawLow) * ReferenceRange) / RawRange) + ReferenceLow. Before I do that, I need to remove the incorrect values in 64.2 and 76.2 that are over 45 and under 0. Since I know these are incorrect readings and I would want them removed from ALL loggers, I am just going to apply it to the whole dataset.

```{r}
clean_results <- allbatresults |> 
  filter(Temperature > 0 & # removes temperature data under 0 degrees
        Temperature < 45) # removes temperature over 45 degrees

summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
summary(as.factor(clean_results$Bat_ID)) # comparing counts between the two
```
This looks good. Now I am going to add columns to the dataframe with the maximum and minimum values during calibration. There were two different calibrations, so first I have to select loggers only used in the second calibration. 

```{r}
Clean_results_met <- clean_results %>% 
  left_join(clean_met, by = "Bat_ID") # join metadata with results so I can extract loggers in second calibration

Cal_2_results <- Clean_results_met |>  # extracting second calibration results
  filter(Calibration == 2) |> 
  select(Bat_ID, DateTime, Unit, Temperature)

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[DateTime >= ymd_hms("2025-06-02 10:30:00") & 
                  DateTime <= ymd_hms("2025-06-02 14:20:00")],  na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[DateTime >= ymd_hms("2025-06-03 01:10:00") & 
                  DateTime <= ymd_hms("2025-06-3 11:10:00")],  na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range from average high temp and average low temp
  ungroup()  
```

Okay, now because Ken would want me to, I am going to load in the calibration data and get the range and average low. 

```{r}
Reference_cal_2 <- read_excel("MOCO_Calibration_Group1_2.xlsx", 
    range = "A1:D6362", na = "NA") #loading in the calibration 2 reference data

summary(Reference_cal_2) # looking at it to make sure it loaded in correctly

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(lowT = mean(T1[No.s >= 403 & 
                  No.s <= 1323],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. I needed to use the time point because the EA15 does not record date so it loaded in as 1899-12-31 for all times. The time points lines up with the same time used in averaging low and high logger temperatures.

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(highT = mean(T1[No.s >= 3923 & 
                  No.s <= 6323],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. 
```

The average low is 4.803, the average high is 41.16, and the range is 36.357. 
Okay, now I am going to create a column in my cal_2_results that has the formula for the corrected values. 

```{r}
Cal_2_results <- Cal_2_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 36.357)/ RangeT) + 4.803) # applying formula listed above to the calibration to get a corrected temperature value

summary(Cal_2_results)
```
I think this data looks good. I took a look at the dataset and saw that the range is different, the low and high looks like they were calculated correctly. 

Now I am going to do the same thing but for the other calibration day

```{r}
Cal_1_results <- Clean_results_met |> # extracting second calibration results
  filter(Calibration == 1) |> 
  select(Bat_ID, DateTime, Unit, Temperature)

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[DateTime >= ymd_hms("2025-05-19 00:40:00") & 
                  DateTime <= ymd_hms("2025-05-19 07:50:00")],  na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[DateTime >= ymd_hms("2025-05-19 14:40:00") & 
                  DateTime <= ymd_hms("2025-05-19 16:40:00")],  na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range and add column
  ungroup()  
```



```{r}
Reference_cal_1 <- read_excel("18_May_2025_Calibration.xlsx", 
    range = "A1:D5408", na = "NA") # loading in the calibration 2 reference data

summary(Reference_cal_1)

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(lowT = mean(T1[No.s >= 1498 & 
                  No.s <= 3220],  na.rm = TRUE)) #taking the average temperature during the low calibration for each logger

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(highT = mean(T1[No.s >= 4858 & 
                  No.s <= 5338],  na.rm = TRUE)) #taking the average temperature during the high calibration for each logger

summary(Reference_cal_1)
```

The average low is 3.672, the average high is 42.26, and the range is 38.588. 
Okay, now I am going to create a column in my cal_1_results that has the formula for the corrected values. 

```{r}
Cal_1_results <- Cal_1_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 38.588)/ RangeT) + 3.672)

summary(Cal_1_results)
```

Okay, now that I have my data calibrated, I am going to merge the two datasets.

```{r}
cal_data <- bind_rows(Cal_1_results, Cal_2_results) |> 
  select(Bat_ID, DateTime, corrected) |> 
  rename("Temperature" = "corrected")

```


## Left and Right Trimming the data

I will split the data, filter it, and then recombine them.

```{r}
# for analysis down below, I cannot use DateTime data so I need to convert my time into a numeric 
clean_met <- clean_met %>%
  mutate(Attachment_Date = as.numeric(`Date Attached`),
         Removal_Date = as.numeric(`Date Removed`)) 

cal_data <- cal_data %>%
  mutate(NumberDate = as.numeric(DateTime))

temps_with_meta <- cal_data %>%
  left_join(clean_met, by = "Bat_ID") #joining my calibrated data with my metadata

# Filter to keep only records after attachment date
left_trimmed <- temps_with_meta %>%
  filter(NumberDate >= Attachment_Date)

# Filter to keep only records before removal date
right_trimmed <- left_trimmed |> 
  filter(NumberDate <= Removal_Date)

# Making a new data table with trimmed data 
allbatresults2 <- right_trimmed |> 
  select(Bat_ID, DateTime, NumberDate, Temperature) |> 
  as.data.table()

#checking the data
summary(as.factor(allbatresults2$Bat_ID))

glimpse(allbatresults2)

```

```{r}
plotly_fun2 <- function(arg1) {
  plot_ly(data = allbatresults2[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(batnames, plotly_fun2)
```
## Left trim again

I left trimmed earlier to remove all data from before the day of attachment. I really want to include as much as possible so I am trimming again by the attachment time. This is going to take forever... :/

```{r}

# first trim
Trimming1 <- allbatresults2 |> 
  filter(Bat_ID %in% c("22", "16", "21", "3", "9"))
summary(as.factor(Trimming1$Bat_ID))

Trimming1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim1 <- Trimming1 |>
  filter(NumberDate >= "1748967000")

Trim1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

summary(as.factor(Trim1$Bat_ID))

#second trim

Trimming2 <- allbatresults2 |> 
  filter(Bat_ID %in% c("39", "41", "44"))
summary(as.factor(Trimming2$Bat_ID))

Trimming2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim2 <- Trimming2 |>
  filter(NumberDate >= "1749151200")

Trim2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#third trim

Trimming3 <- allbatresults2 |> 
  filter(Bat_ID %in% c("46"))
summary(as.factor(Trimming3$Bat_ID))

Trimming3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim3 <- Trimming3 |>
  filter(NumberDate >= "1749154800")

Trim3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#fourth trim

Trimming4 <- allbatresults2 |> 
  filter(Bat_ID %in% c("47", "48"))
summary(as.factor(Trimming4$Bat_ID))

Trimming4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim4 <- Trimming4 |>
  filter(NumberDate >= "1749229200")

Trim4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#fifth trim

Trimming5 <- allbatresults2 |> 
  filter(Bat_ID %in% c("49"))
summary(as.factor(Trimming5$Bat_ID))

Trimming5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim5 <- Trimming5 |>
  filter(NumberDate >= "1748978400")

Trim5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# sixth trim
Trimming6 <- allbatresults2 |> 
  filter(Bat_ID %in% c("50", "60", "62"))
summary(as.factor(Trimming6$Bat_ID))

Trimming6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim6 <- Trimming6 |>
  filter(NumberDate >= "1749220800")

Trim6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# seventh trim 

Trimming7 <- allbatresults2 |> 
  filter(Bat_ID %in% c("54", "55"))
summary(as.factor(Trimming7$Bat_ID))

Trimming7 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim7 <- Trimming7 |>
  filter(NumberDate >= "1749219000")

Trim7 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# eighth trim
Trimming8 <- allbatresults2 |> 
  filter(Bat_ID %in% c("64.1"))
summary(as.factor(Trimming8$Bat_ID))

Trimming8 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim8 <- Trimming8 |>
  filter(NumberDate >= "1747736400")

Trim8 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Ninth trim

Trimming9 <- allbatresults2 |> 
  filter(Bat_ID %in% c("64.2"))
summary(as.factor(Trimming9$Bat_ID))

Trimming9 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim9 <- Trimming9 |>
  filter(NumberDate >= "1749923400")

Trim9 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Tenth trim

Trimming10 <- allbatresults2 |> 
  filter(Bat_ID %in% c("68.1", "74.1"))
summary(as.factor(Trimming10$Bat_ID))

Trimming10 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim10 <- Trimming10 |>
  filter(NumberDate >= "1747741800")

Trim10 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 11
Trimming11 <- allbatresults2 |> 
  filter(Bat_ID %in% c("68.2", "74.2", "78.2"))
summary(as.factor(Trimming11$Bat_ID))

Trimming11 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim11 <- Trimming11 |>
  filter(NumberDate >= "1749919200")

Trim11 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Trim 12 

Trimming12 <- allbatresults2 |> 
  filter(Bat_ID %in% c("73.4"))
summary(as.factor(Trimming12$Bat_ID))

Trimming12 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim12 <- Trimming12 |>
  filter(NumberDate >= "1750091400")

Trim12 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 13

Trimming13 <- allbatresults2 |> 
  filter(Bat_ID %in% c("76.2"))
summary(as.factor(Trimming13$Bat_ID))

Trimming13 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim13 <- Trimming13 |>
  filter(NumberDate >= "1749987600")

Trim13 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 14

Trimming14 <- allbatresults2 |> 
  filter(Bat_ID %in% c("78.3"))
summary(as.factor(Trimming14$Bat_ID))

Trimming14 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim14 <- Trimming14 |>
  filter(NumberDate >= "1750006800")

Trim14 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Trim 15

Trimming15 <- allbatresults2 |> 
  filter(Bat_ID %in% c("83.2"))
summary(as.factor(Trimming15$Bat_ID))

Trimming15 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim15 <- Trimming15 |>
  filter(NumberDate >= "1749920400")

Trim15 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 16

Trimming16 <- allbatresults2 |> 
  filter(Bat_ID %in% c("26"))
summary(as.factor(Trimming16$Bat_ID))

Trimming16 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim16 <- Trimming16 |>
  filter(NumberDate >= "1748968200")

Trim16 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))


LTrimmed <- bind_rows(Trim1, Trim2, Trim3, Trim4, Trim5, Trim6, Trim7, Trim8, Trim9, Trim10, Trim11, Trim12, Trim13, Trim14, Trim15, Trim16)



summary(as.factor(LTrimmed$Bat_ID))
summary(as.factor(allbatresults2$Bat_ID))
```

## Right trim again 

```{r}
# Right trim 1
RTrimming1 <- LTrimmed |> 
  filter(Bat_ID %in% c("64.1"))
summary(as.factor(RTrimming1$Bat_ID))

RTrimming1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim1 <- RTrimming1 |>
  filter(NumberDate <= "1749673200")

RTrim1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right trim 2
RTrimming2 <- LTrimmed |> 
  filter(Bat_ID %in% c("68.1"))
summary(as.factor(RTrimming2$Bat_ID))

RTrimming2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim2 <- RTrimming2 |>
  filter(NumberDate <= "1748880600")

RTrim2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right trim 3

RTrimming3 <- LTrimmed |> 
  filter(Bat_ID %in% c("78.2"))
summary(as.factor(RTrimming3$Bat_ID))

RTrimming3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim3 <- RTrimming3 |>
  filter(NumberDate <= "1749931200")

RTrim3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right Trim 4 

RTrimming4 <- LTrimmed |> 
  filter(Bat_ID %in% c("22"))
summary(as.factor(RTrimming4$Bat_ID))

RTrimming4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim4 <- RTrimming4 |>
  filter(NumberDate <= "1749742800")

RTrim4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# RTrim 5

RTrimming5 <- LTrimmed |> 
  filter(Bat_ID %in% c("21"))
summary(as.factor(RTrimming5$Bat_ID))

RTrimming5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim5 <- RTrimming5 |>
  filter(NumberDate <= "1749439200")

RTrim5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# RTrim 6

RTrimming6 <- LTrimmed |> 
  filter(Bat_ID %in% c("3"))
summary(as.factor(RTrimming6$Bat_ID))

RTrimming6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim6 <- RTrimming6 |>
  filter(NumberDate <= "1749276000")

RTrim6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Keeping all results from LTrimmed besides the ones that were right trimmed again

RTrimmed_Temp <- LTrimmed |> 
  filter(!Bat_ID %in% c("64.1", "68.1", "78.2", "22", "21", "3"))

summary(as.factor(RTrimmed_Temp$Bat_ID))

RTrimmed <- bind_rows(RTrimmed_Temp, RTrim1, RTrim2, RTrim3, RTrim4, RTrim5, RTrim6)

summary(as.factor(RTrimmed$Bat_ID))

```

Now I want to look at the data and make sure it looks right and see what else needs to be trimmed

```{r}
plotly_fun3 <- function(arg1) {
  plot_ly(data = RTrimmed[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(batnames, plotly_fun3)
```



Dropping bad data and gaps

```{r}
# Dropping 1
Dropping1 <- RTrimmed |> 
  filter(Bat_ID %in% c("68.2"))

summary(as.factor(Dropping1$Bat_ID))

Drop1 <- Dropping1 |>
  filter(NumberDate <= "1750024800" | NumberDate >= "1750091400")

summary(as.factor(Drop1$Bat_ID))


#Dropping 2

Dropping2 <- RTrimmed |> 
  filter(Bat_ID %in% c("83.2"))

summary(as.factor(Dropping2$Bat_ID))

Drop2 <- Dropping2 |>
  filter(NumberDate <= "1750009200" | NumberDate >= "1750090800")

#Dropping 3

Dropping3 <- RTrimmed |> 
  filter(Bat_ID %in% c("22"))

summary(as.factor(Dropping3$Bat_ID))

Drop3 <- Dropping3 |>
  filter(NumberDate <= "1749084600" | NumberDate >= "1749099600")


# I also have mutliple loggers from the same bat, so I am going to merge these together so it shows up as one bat instead of two. 

Bat_78.2 <- RTrimmed |> 
  filter(Bat_ID == "78.2")
Bat_78.3 <- RTrimmed |> 
  filter(Bat_ID == "78.3")

Bat_78.9 <- bind_rows(Bat_78.2, Bat_78.3) |>  
  mutate(ID = "78.9" ) |> 
  select(DateTime, NumberDate, Temperature, ID) |> 
  rename(Bat_ID = ID)

# Removing the edited data in this chunk from RTrimmed so I can merge all the data back together


allbatresults3.temp <- RTrimmed |> 
  filter(!Bat_ID %in% c("68.2", "83.2", "78.2", "78.3", "22"))

clean_results <- bind_rows(allbatresults3.temp, Drop2, Drop1, Drop3, Bat_78.9)


summary(as.factor(clean_results$Bat_ID))
```

I need to remove some names from batnames. There are a handful of loggers from bats that are excluded from the study or have too few data points and were dropped during trimming

```{r}

KeptBatNames <- unique(clean_results$Bat_ID)

plotly_fun4 <- function(arg1) {
  plot_ly(data = clean_results[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(KeptBatNames, plotly_fun4)
```

```{r}

checking <- clean_results%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(checking)
``` 


```{r}
write_tsv(clean_results, file = "TrimmedTemperatureLoggerData.long.tsv")
```


## Ambient Temp
We had two cohorts of bats for non thermally challenged bats. They followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. Because of this, ambient temperature is different for group 1 D21 bats and group 2 D21 bats.

```{r}
 Union_Ambient <- read_excel("MOCO Data/Roost Temps/MOCO_Union_Ambient.xlsx") |> 
  rename(DateTime = "Date/Time")  |> # read in ambient temp
  mutate (DateTime = as.POSIXct(DateTime)) |> # correct format
  mutate(NumberDate = as.numeric(DateTime)) #add column changing date/time to numeric
  
Union_Ambient <- Union_Ambient |> 
filter(NumberDate >= "1747407000")

plot_ly(Union_Ambient,x = ~DateTime, y = ~Value,
          type = 'scatter', mode = 'lines')

Rivendell_Top_D0 <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Top_D0.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge 
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Bottom_D0 <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Bottom_D0.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge
  mutate(DateTime = as.POSIXct(DateTime)) |> 
   mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Primary <- Rivendell_Top_D0 |> 
  left_join(Rivendell_Bottom_D0, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

Rivendell_Primary <- Rivendell_Primary |> 
  select(DateTime.x, NumberDate, Value.x, Value.y, Unit.x)  |> # keeping certain columns 
  rename(Unit = "Unit.x", # renaming columns
         Top = "Value.x",
         Bottom = "Value.y",
         DateTime = "DateTime.x") |> 
  drop_na(Bottom) # dropping NA's because loggers were not started and stopped at the exact same time
glimpse(Rivendell_Primary)

# because I have temperature from the top and bottom of the incubator and there might be slight differences between them and I don't know which temp the bats are exposed to. I am going to get the average

Rivendell_Primary <- Rivendell_Primary |> 
  mutate(avg_pri = rowMeans(pick("Top", "Bottom"))) 
  
```

 Now I have to do the same thing but for the pre- thermal challenge
```{r}
Rivendell_Top_Pre <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Top_D-5.xlsx") |> # read in temp from top of incubator from D0 thermal challenge
  rename(DateTime = "Date/Time") |> 
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Bottom_Pre <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Bottom_D-5.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Pre <- Rivendell_Top_Pre |> 
  left_join(Rivendell_Bottom_Pre, join_by(NumberDate)) |> # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

Rivendell_Pre <- Rivendell_Pre |> 
  select(DateTime.x, NumberDate, Value.x, Value.y, Unit.x)  |> # keeping certain columns 
  rename(Unit = "Unit.x", # renaming columns
         Top = "Value.x",
         Bottom = "Value.y",
         DateTime = "DateTime.x") |> 
  drop_na(Bottom) # dropping NA's because loggers were not started and stopped at the exact same time
glimpse(Rivendell_Pre)

# because I have temperature from the top and bottom of the incubator and there might be slight differences between them and I don't know which temp the bats are exposed to. I am going to get the average

Rivendell_Pre <- Rivendell_Pre |> 
  mutate(avg_pre = rowMeans(pick("Top", "Bottom"))) 

```

Now I want to only keep the data from the actual thermal challenges
Start time for the pre-thermal challenge was 12:05 23-May-2025 to 01:17 24-May-2025
Start Time for primary thermal challenge was 16:36 26-May-2025 to 00:44 27-May-2025

```{r}
Rivendell_Primary_Trim <- Rivendell_Primary |> 
  filter(NumberDate > "1748277361", 
         NumberDate < "1748306641") |> 
  select(DateTime, NumberDate, avg_pri) |> 
  rename(Value = "avg_pri")

Rivendell_Pre_Trim <- Rivendell_Pre |> 
  filter(NumberDate > "1748001901", 
         NumberDate < "1748049421") |> 
  select(DateTime, NumberDate, avg_pre) |> 
  rename(Value = "avg_pre")
```

Okay, now that's done.. I need to take these times and add them to the corresponding rows. I think what I am going to do is separate the boost from the primary results, because my primary result bats were all in an incubator at that time. 
```{r}
Primary_results_working <- clean_results |> 
  filter(Bat_ID %in% c("64.1", "68.1", "74.1")) |> # results from primary immunization
  filter(NumberDate <= "1748736000")

Secondary_results_working <- clean_results |> 
  filter(!Bat_ID %in% c("64.1", "68.1", "74.1")) # results from secondary immunization

# trim down the ambient data to the dates and time loggers were attached to the bats
Ambient_primary_working <- Union_Ambient |> 
  filter(NumberDate <= "1748736000", 
         NumberDate >= "1747736400")

# now I need to filter out times when the bats were in the incubators instead of outside with ambient temperature
AP_Trim1 <- Ambient_primary_working |> 
  filter(NumberDate <= "1748001901")

AP_Trim2 <- Ambient_primary_working |> 
   filter(NumberDate >= "1748049421",
          NumberDate <= "1748277361")

Ap_Trim3 <- Ambient_primary_working |> 
  filter(NumberDate >= "1748306641")
          
# Now I am going to combine the incubator temps and ambient temps into one dataframe in one column. This willl make calculating my delta T much easier.. I think.... 

Primary_exposure_temp <- bind_rows(AP_Trim1, Rivendell_Pre_Trim, AP_Trim2, Rivendell_Primary_Trim, Ap_Trim3)

Primary_results_working <- Primary_results_working |>  
  left_join(Primary_exposure_temp, join_by("NumberDate"))

Primary_results_working <- Primary_results_working |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, Value, Unit) |> 
  rename(AmbientT = "Value",
         DateTime = "DateTime.x")

Primary_results <- Primary_results_working |> 
  mutate(deltaT = Temperature - AmbientT)

Secondary_results <- Secondary_results_working |> 
  left_join(Union_Ambient, join_by(NumberDate))

Secondary_results <- Secondary_results |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, Value, Unit) |> 
  rename(AmbientT = "Value",
         DateTime = "DateTime.x")

Secondary_results <- Secondary_results |> 
  mutate(deltaT = Temperature - AmbientT)

```



### Shifting Data

We had two cohorts of bats for non thermally challenged bats. They followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. Because of this, ambient temperature is different for group 1 D21 bats and group 2 D21 bats. 

```{r}
SecondaryBatNames <- unique(Secondary_results$Bat_ID)

Secondary_Met <- no_hot_met %>%
  filter(Bat_ID %in% SecondaryBatNames) |> 
  select(! "Logger location") |> 
  select(! "Notes")

summary(as.factor(Secondary_Met$Treatment))
# shows 6 TC bats but I thought we had 7.. look into this.

SecondaryResultsMet <- Secondary_results |> 
left_join(Secondary_Met, by = join_by(Bat_ID))

SecondaryResultsMet %>%
  group_by(Bat_ID) %>%
  summarize(Min_DateTime = min(DateTime), Max_DateTime = max(DateTime))
```

```{r}
SecondaryResultsMet$time_aligned <- SecondaryResultsMet$DateTime

md.pattern(SecondaryResultsMet, rotate.names = TRUE)

SecondaryResultsMet <- SecondaryResultsMet |> 
  select(! "Serial Number") |> 
  select(! "Logger Name") |> 
  select(! "Filename")

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] - lubridate::days(2)

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] - lubridate::days(11)
```

Okay, one of the last things I want to do is remove the decimal from the Bat ID. This was just to differentiate between loggers for one bat with all the different metadata. 
```{r}
Working_Secondary <- SecondaryResultsMet |> 
  filter(Treatment %in% c("SHAM", "eVLP+LPS", "TC"))

Bat_64 <- Working_Secondary |> 
  filter(Bat_ID == "64.2")
Bat_64$Bat_ID <- c("64")

Bat_68 <- Working_Secondary |> 
  filter(Bat_ID == "68.2")
Bat_68$Bat_ID <- c("68")

Bat_73 <- Working_Secondary |> 
  filter(Bat_ID == "73.4")
Bat_73$Bat_ID <- c("73")

Bat_74 <- Working_Secondary |> 
  filter(Bat_ID == "74.2")
Bat_74$Bat_ID <- c("74")

Bat_76 <- Working_Secondary |> 
  filter(Bat_ID == "76.2")
Bat_76$Bat_ID <- c("76")

Bat_78 <- Working_Secondary |> 
  filter(Bat_ID == "78.9")
Bat_78$Bat_ID <- c("78")

Bat_83 <- Working_Secondary |> 
  filter(Bat_ID == "83.2")
Bat_83$Bat_ID <- c("83")

Working_Secondary2 <- Working_Secondary |> 
  filter(!Bat_ID %in% c("64.2", "68.2", "73.4", "74.2", "76.2", "78.9", "83.2"))

Working_Secondary3 <- Working_Secondary2 |> 
  bind_rows(Bat_64, Bat_68, Bat_73, Bat_74, Bat_76, Bat_78, Bat_83)

Final_working_sec <- Working_Secondary3 |> 
  select(Bat_ID, time_aligned, Temperature, AmbientT, deltaT)

Final_Secondary <- Final_working_sec |> 
  rename(DateTime = "time_aligned") |> 
  mutate(NumberDate = as.numeric(DateTime)) 

SecondaryNames <- unique(Final_Secondary$Bat_ID)

Working_met_sec <- Working_Secondary3 |> 
  select(Bat_ID, `Bat Pittag`, Group, Treatment, `Innoculation Date`, `Boost Date`)

Secondary_Met2 <- Working_met_sec |> 
  unique()

SecondaryNames <- unique(Final_Secondary$Bat_ID)

temp_primary_met <- Primary_results |> 
  left_join(clean_met, join_by(Bat_ID))

Bat_64.1 <- temp_primary_met |> 
  filter(Bat_ID == "64.1")
Bat_64.1$Bat_ID <- c("64")

Bat_68.1 <- temp_primary_met |> 
  filter(Bat_ID == "68.1")
Bat_68.1$Bat_ID <- c("68")

Bat_74.1 <- temp_primary_met |> 
  filter(Bat_ID == "74.1")
Bat_74.1$Bat_ID <- c("74")

Primary_working <- bind_rows(Bat_64.1, Bat_68.1, Bat_74.1)

Final_Primary <- Primary_working |> 
  select(Bat_ID, DateTime, NumberDate, Temperature, AmbientT, deltaT)

PrimaryNames <- unique(Final_Primary$Bat_ID)

Primary_met_working <- Primary_working |> 
  select(Bat_ID, `Bat Pittag`, Group, Treatment, `Innoculation Date`, `Boost Date`)

Primary_Met <- Primary_met_working|> 
  unique()

Primary_Met <- clean_met |> 
  filter(Bat_ID %in% c("64.1", "68.1", "74.1")) 

Secondary_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1748967000" , 
          NumberDate < "1749772800")

Group_1_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1747406400", 
         NumberDate < "1750290600")

Group_2_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1747440000", 
         NumberDate < "1750447800")

Group_3_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1748217600", 
         NumberDate < "1750290600")
```

```{r}

plotly_fun6 <- function(arg1) {
  plot_ly(data = Final_Secondary[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}

lapply(SecondaryNames, plotly_fun6)

```
 
I shifted the data so now data all shows d21 on June 5th, the same day as D21 for Group 1. 
I think I have all the data wrangling down so I am going to export some tsv with my final primary data, secondary data, metadata for both immunizations, and ambient temperature.

```{r}
Secondary_Met2$Treatment[Secondary_Met2$Treatment == "TC"] <- "Cold"
Secondary_Met2$Treatment[Secondary_Met2$Treatment == "eVLP+LPS"] <- "NoCo"

Primary_Met$Treatment[Primary_Met$Treatment == "TC"] <- "Cold"
Primary_Met$Treatment[Primary_Met$Treatment == "eVLP+LPS"] <- "NoCo"

```



```{r}
write_tsv(Secondary_Met2, file = "MOCO2025Secondary_Met.tsv")
write_tsv(Final_Primary, file = "MOCO2025PrimaryTemperature.tsv")
write_tsv(Final_Secondary, file = "MOCO2025SecondaryTemperature.tsv")
write_tsv(Primary_Met, file = "MOCO2025Primary_Met.tsv")
write_tsv(Ambient_primary_working, file = "MOCO2025Primary_Ambient.tsv")
write_tsv(Secondary_Ambient, file = "MOCO2025Secondary_Ambient.tsv")
write_tsv(Group_1_Ambient, file = "MOCO2025_G1_Ambient.tsv")
write_tsv(Group_2_Ambient, file = "MOCO2025_G2_Ambient.tsv")
write_tsv(Group_3_Ambient, file = "MOCO2025_G3_Ambient.tsv")
write_tsv(Union_Ambient, file = "MOCO2025_Ambient.tsv")

```



