---
title: "MOCO 2025 Temperature Loggers Analysis"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read.delim("EPFU2025Metadata.tsv", # read in metadata for ibuttons
     na = "NA")

glimpse(Metadata) # check formatting

Metadata$Bat_ID = as.character(Metadata$Bat_ID) # change Bat_ID to a character

glimpse(Metadata) # check formatting again
```

I am going to load in my data and metadata for both immunizations

```{r}
Primary_Results <- read.delim("EPFU2025PrimaryTemperature.tsv", # read in primary results
     na = "NA")

Secondary_Results <- read.delim("EPFU2025SecondaryTemperature.tsv", # read in secondary results 
     na = "NA")

Bat_Results <- read.delim("EPFU2025BatResults.tsv", # read in secondary results 
     na = "NA")

Clean_Results <- read.delim("EPFU2025Cleaned_Results.tsv", # read in secondary results 
     na = "NA")

Basal_TC <- read.delim("EPFU2025Basal_TC.tsv", # read in secondary results 
     na = "NA")

Primary_TC <- read.delim("EPFU2025Primary_TC.tsv", # read in secondary results 
     na = "NA")

Secondary_TC <- read.delim("EPFU2025Secondary_TC.tsv", # read in secondary results 
     na = "NA")

Results <- read.delim("EPFU2025Results_Analysis.tsv",
                      na = "NA")
```

Make values for bat names
```{r}
BasalNames <- unique(Basal_TC$Bat_ID)
SecondaryBatNames <- unique(Secondary_Results$Bat_ID)
PrimaryBatNames <- unique(Primary_Results$Bat_ID)
ResultsNames <- unique(Results$Bat_ID)
```

## Average and running temp

```{r}
Results <- Results |> 
  group_by(Bat_ID) |> 
  mutate(Temp_norm = mean(Temperature[DateType == "None" &
                                      Temperature > 31], na.rm = TRUE)) |> 
  ungroup()

Results <- Results |> 
  # 1. Calculate the difference for every time point
  mutate(Temp_Diff = Temperature - Temp_norm) |> 
    # 2. Sort chronologically by bat and time to ensure cumsum adds in the right order
  # (Replace 'DateTime' with the actual name of your timestamp column)
  arrange(Bat_ID, NumberDate) |> 
    # 3. Group by the Bat and your existing Experiment Day column
  # (Replace 'Exp_Day' with the actual name of your D-5, D0, D1 column)
  group_by(Bat_ID, ExperimentDay) |> 
    # 4. Calculate the running sum. Because of the group_by, 
  # it will automatically reset to 0 at the start of each experiment day.
  mutate(running_AUC = cumsum(ifelse(is.na(Temp_Diff), 0, Temp_Diff)),
    Total_AUC = sum(Temp_Diff, na.rm = TRUE),
         valid_counts = sum(!is.na(Temp_Diff))) |> 
    # 5. Clean up grouping
  ungroup()

Analysis_results <- Results |> 
  filter(grepl("13:50", DateTime)) |> 
  select(Bat_ID, DateType, ExperimentDay, Sex, Treatment, Group, Temp_norm, Handled, Total_AUC, valid_counts, Weight, Cage, )
```

## First models

```{r}
handled_model <- glm(Total_AUC ~ Bat_ID + Handled, data = Analysis_results)
summary(handled_model)
```
Looks like handling day matters for some bats more than others. 

I am going to take AUC days bats were not handled and compare them to primary +3 and secondary +3. If there is no difference, I can group these days as non-handled 

```{r}
handling_check <- Analysis_results |> 
  filter(DateType %in% c("None", "Primary + 3", "Secondary + 3")) |> 
  mutate(DateType = factor(DateType),
    DateType = relevel(DateType, ref = "None"))

handle_check <- glm(Total_AUC ~ Bat_ID + DateType, 
                 data = handling_check, 
                 family = gaussian(link = "identity"))

summary(handle_check)
```
Hmm again, depends on the bat, but primary +3 is significantly different from non-handling days but secondary + 3 is not significantly different. I was considering grouping these days into "None" but I am going to keep it separate. 

```{r}
treatment_glm <- glm(Total_AUC ~ Treatment, data = Analysis_results)
summary(treatment_glm)
```
So AUC for NOCO, Primary, and Dual is not significant, but secondary and sham are significantly different. 

```{r}
treatment_days_glm <- glm(Total_AUC ~ Treatment + DateType , data = Analysis_results)
summary(treatment_days_glm)
```


I want to look at AUC for bats in basal TC and factor in cage.
```{r}
basal_analysis <- Analysis_results |> 
  filter(DateType == "Basal TC")

basal <- glm(Total_AUC ~ Cage, data = basal_analysis)
summary(basal)
```
So it looks like cage did not effect AUC


```{r}
BP_TC_analysis <- Analysis_results |> 
  filter(DateType %in% c("Basal TC", "Primary") & Treatment %in% c("Dual", "Primary"))

BP_Analysis <- glm(Total_AUC ~ DateType ,data = BP_TC_analysis)
summary(BP_Analysis)
```
Okay! Clearly a difference in AUC when comparing basal TC to primary TC!

I wanted to run a glm to see if the number of counts had an effect and it definitely does. I need to talk to Ken about how to deal with this...
```{r}
data_points_glm <- glm(Total_AUC ~ valid_counts , data = Analysis_results)
summary(data_points_glm)
```


Looking at weight
```{r}
weight_glm <- glm(Total_AUC ~ Weight , data = Analysis_results)
summary(weight_glm)
```
Significant but not surprising. I do need to check and see if the NAs are interfering with this. 

## Basal Temp

```{r}
Metadata <- Metadata |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id))

Basal_TC <- Basal_TC |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

Basal_met <- data.table::data.table(Metadata, key = "id")
Basal_TC2 <- data.table::data.table(Basal_TC, key = "id")

 basal_behavr <- behavr(Basal_TC2, Basal_met)
```

```{r}
basal_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment))+
  #geom_rect(data = shaded, 
          #  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
           #  alpha = 0.3, inherit.aes = FALSE) +
    stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(Treatment ~ .) +
#  scale_x_continuous(
#    name = "Date/time",
#    breaks = seq(1747767600, max(primary_behavr$t), by = 86400), 
#    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, #%H:%M"),
#    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(y = "Temp (C)")-> EPFU_Basal

ggsave("EPFU_Basal.jpeg", plot = EPFU_Basal, device = jpeg(), width = 12, height = 6.67)
```


##Primary Temp

```{r}

Primary_Results <- Primary_Results |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 Primary_met <- data.table::data.table(Metadata, key = "id")
 Primary_results2 <- data.table::data.table(Primary_Results, key = "id")

 primary_behavr <- behavr(Primary_results2, Primary_met)
 summary(primary_behavr)
 
 glimpse(primary_behavr)

```

### Primary Behavr 
#### NOTE FOR ASH 
   I need to update this code.. I want to change the x axis to show experiment day instead of date/time. I also want to reorder them to sham is on the bottom
```{r}
 
#shaded <- data.frame(xmin = 1748001900 , #adds shaded regions for both the basal and primary thermal #challenges
#                     xmax = 1748049000,
#                     label = c("Basal TC", "Primary TC"))
                              
  primary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment))+
  #geom_rect(data = shaded, 
          #  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
           #  alpha = 0.3, inherit.aes = FALSE) +
    stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Experiment Day",
    # Set breaks every 24 hours (86400 seconds)
    breaks = seq(1747767600, max(primary_behavr$t), by = 86400), 
    # Calculate the day number relative to your start timestamp
    labels = function(x) {
      day_num <- (as.numeric(x) - 1747767600)
      paste("Day", round(day_num))},
      expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(x = "Experiment Day", y = "Temp (°C)")-> EPFU_Primary 
#    name = "Date/time",
#    breaks = seq(1747767600, max(primary_behavr$t), by = 86400), 
#    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, #%H:%M"),
    #scale_x_date(labels = function(x), paste("Day", as.numeric(x - 1757289601))) +
   

ggsave("EPFU_Primary.jpeg", plot = EPFU_Primary, device = jpeg(), width = 12, height = 6.67)

```
It is plotting and looks fine, but I am getting some errors. Need Ken to look at it. I also want to add shaded regions that show when they are being handled. But I actually might not do this because these bats are handled every single night for feeding.. ugh stupid MOCO

```{r}
 primary_behavr |>
  subset(Treatment %in% c("NOCO", "Secondary")) |> 
  transform_meta
  ggetho(aes(x = t, y = Temperature, colour = Treatment))+
  #geom_rect(data = shaded, 
          #  aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
           #  alpha = 0.3, inherit.aes = FALSE) +
    stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1747767600, max(primary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
    labs(y = "Temp (C)")-> EPFU_Primary

ggsave("EPFU_Primary.jpeg", plot = EPFU_Primary, device = jpeg(), width = 12, height = 6.67)
```


## Behavr Analysis

```{r}

Secondary_Results <- Secondary_Results |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 Secondary_met <- data.table::data.table(Metadata, key = "id")
 Secondary_results2 <- data.table::data.table(Secondary_Results, key = "id")

 secondary_behavr <- behavr(Secondary_results2, Secondary_met)
 summary(secondary_behavr)
```

```{r}
secondary_etho  <- ggetho(secondary_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho() +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Temp (C)",
       y = "Bat ID")

ggsave("EPFU_Sec_Etho.jpeg", plot = secondary_etho, device = jpeg(), width = 12, height = 6.67)
```

```{r}
ggetho(secondary_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r, warning=FALSE}

secondary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(Treatment ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) -> EPFU_Secondary
 
ggsave("EPFU_Secondary.jpeg", plot = EPFU_Secondary, device = jpeg(), width = 12, height = 6.67)

```

```{r, warning=FALSE}
# i don't think this is working
ggetho(primary_behavr, aes(y= Temperature, colour= Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
     # facet_grid(id ~ .) +
      scale_y_continuous(name= "Temperature")
```


```{r}
ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(Treatment ~ .) +
      scale_y_continuous(name= "Temperature")
ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      scale_y_continuous(name= "Temperature")

Secondary_Trimmed <- Secondary_results2 |> 
  filter(NumberDate <= "1749254400") |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))


 Secondary_results3 <- data.table::data.table(Secondary_Trimmed, key = "id")

 secondary_behavr2 <- behavr(Secondary_results3, Secondary_met2)
 summary(secondary_behavr)

 
 ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      scale_y_continuous(name= "Temperature")
```

```{r Sham, warning=FALSE}

met_sham <- Secondary_met2 %>%
  filter(Treatment == "SHAM")
sham_results <- Secondary_results2 %>%
  filter(id %in% met_sham$id)
sham_behavr <- behavr(sham_results, met_sham)

sham_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r eVLP_LPS, warning=FALSE}
met_LPS <- Secondary_met2 %>%
  filter(Treatment == "eVLP+LPS")
LPS_results <- Secondary_results2 %>%
  filter(id %in% met_LPS$id)
LPS_behavr <- behavr(LPS_results, met_LPS)

LPS_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r TC, warning=FALSE}
met_TC <- Secondary_met2 %>%
  filter(Treatment == "TC")
TC_results <- Secondary_results2 %>%
  filter(id %in% met_TC$id)
TC_behavr <- behavr(TC_results, met_TC)

TC_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r, warning=FALSE}
ggetho(bat_behavr, aes(y=Temperature, colour=Treatment)) + 
  stat_pop_etho()
```

```{r, warning=FALSE}
bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(x=t, z=Temperature), multiplot = 2) +
      stat_tile_etho() +
      facet_wrap( ~ id)
```

## Subsetting the data to pre-boost and post-boost

Before I proceed with subsetting, I am going to make sure that I understand the data.

```{r}
allbatresults3 %>%
  group_by(batnames) %>%
  summarize(Min_timepoint = min(Timepoint), Min_DateTime = min(DateTime), Max_DateTime = max(DateTime),
            Min_Temp = min(Temperature), Max_Temp = max(Temperature))
```


```{r, warning=FALSE}
# summary(bat_behavr_preboost)
# bat_behavr_preboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_preboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

And now for the post-boost results. 
For starters, I will trim the 4 days after the injections.
That didn't work because ggetho kept saying that there was less than 3 days of data :/

Secondly just kept the above format but used more data.

```{r}


post_bat_results <- posttreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(post_bat_results)
  
bat_data_postboost <- clean_met3 %>%
  select(id) %>%
  left_join(post_bat_results, by = join_by(id))
str(bat_data_postboost)

bat_data_postboost2 <- data.table::data.table(bat_data_postboost, key = "id")

bat_behavr_postboost <- behavr(bat_data_postboost2, bat_met2)
summary(bat_behavr_postboost)
```
```{r Postboost_Summary, warning=FALSE}
# summary(bat_behavr_postboost)
# bat_behavr_postboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_postboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

So I think that the booster happened on day 4 on these graphs, but I need to verify.

## Booster Analysis

To be certain of using the correct timestamps, I am going to analyze with the openair package, as I did for the squirrels.

This is done one bat at a time...

https://bookdown.org/david_carslaw/openair/sections/trend-analysis/time-variation.html

First, comparing all pre-boost data to (boost day + 1 day following)

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

The results look very similar.


```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

To keep the comparison cleaner, I will continue with pre versus post. 
The only downside is that the confidence interval is a little wider on the "pre" data than the "other" data.

## Comparison of all bats

```{r}
bat_met2 %>%
  arrange(Treatment) %>%
  select(id, Treatment)
```

Let's make a function for the timeVariation plot

```{r}
timeVariation_fun <- function(arg1) {
  mybatID <- arg1

  
  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(2)

  singlebatresults <- allbatresults3 %>%
    filter(batnames == mybatID)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", 
                           group = "Boosted",  
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

```{r}
timeVariation_fun("1N8476")
```

Now let's look at the groups one at a time.

```{r SHAM}
bat_sham <- bat_met2 %>%
  filter(Treatment == "SHAM") %>%
  select(id) %>%
  pull
lapply(bat_sham, timeVariation_fun)
```

```{r eVLP LPS}
bat_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS") %>%
  select(id) %>%
  pull
lapply(bat_LPS, timeVariation_fun)
```

```{r TC}
bat_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP") %>%
  select(id) %>%
  pull
lapply(bat_eVLP, timeVariation_fun)
```

## Tent Logger Data

Two possible ways to incorporate this data. 
First, I will add it as a separate "pollutant".
Second, I will show it as a separate "feature".

```{r}
alltentresults <- fread("TemperatureLoggers/tentloggerdata_long.tsv")
alltentresults <- alltentresults %>%
  mutate(Tent = gsub("Tent ", "", Tent))
str(alltentresults)

tent_met <- Metadata %>%
  drop_na(Filename) %>%
  select(Filename, DMR_ID, Tag, Treatment, Age, Sex,
         Attachment_date = "Date Attached", 
         Treatment_date = Day_00, 
         RIP_date,
         Tent = Tent.x) %>%
  mutate(id = gsub(".DAT", "", Filename)) %>%
  relocate(id)

tentbatresults <- tent_met %>%
  select(id, Tent) %>%
  right_join(allbatresults3, by = join_by(id == batnames))

tentbatresults2 <- tentbatresults %>%
  left_join(alltentresults, by = join_by(Tent, DateTime)) %>%
  rename(`Bat T` = Temperature.x) %>%
  rename(`Tent T` = Temperature.y)

write_csv(tentbatresults2, file = "allbatandtentresults.csv")
```

A new plot function:

```{r}
timeVariation_fun2 <- function(arg1) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  # Immunization <- bat_met2 %>%
  #   filter(id == mybatID) %>%
  #   select(Treatment_date) %>%
  #   pull
  # Boost <- Immunization + days(21)
  # BoostWindow <- days(2)

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID)

  # singlebatresults2 <- singlebatresults %>%
  #   filter(DateTime <= Boost + BoostWindow) %>%
  #   mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  #   mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  #   rename(date = DateTime)
  
  singlebatresults2 <- singlebatresults %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant =c("Tent T", "Bat T"), 
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun2("1N8476")
```
That looks good. Let's plot each group at a time.

```{r Sham-Tent}
lapply(bat_sham, timeVariation_fun2)
```

```{r eVLP + LPS -Tent}
lapply(bat_eVLP, timeVariation_fun2)
```

```{r TC -Tent}
lapply(bat_LPS, timeVariation_fun2)
```

Okay, because the "difference" function worked fine, I don't think that I need to reorganize the data. 
Let's see if I can also use the boost "group" at the same time.

Dammit: "Can only have one pollutant with a grouping variable, or several pollutants and no grouping variable."

Okay, I think I will make a new variable called delta T and use that as the only pollutant.

I will also test different "postboost" windows.

A new plot function:

```{r}
timeVariation_fun3 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant = "delta T", 
                           group = "Boosted",
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun3("1N8476", 5)
```

```{r Sham-deltaT}
lapply(bat_sham, timeVariation_fun3, 3)
```

```{r eVLP + LPS -deltaT}
lapply(bat_eVLP, timeVariation_fun3, 3)
```

```{r TC -deltaT}
lapply(bat_LPS, timeVariation_fun3, 3)
```


## Comparing by Treatment Group

Using the "group" parameter in timeVariation and to make separate plots for pre-boost and then each day post-boost.


```{r Group Comparisons}
tentbatresults3 <- tentbatresults2 %>%
  mutate(`delta T` = `Bat T` - `Tent T`) %>%
  rename(date = DateTime)

bat_boost_date <- bat_met2 %>%
  select(id, Treatment, Treatment_date) %>%
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP",
                                                  "eVLP LPS", "eVLP POLY", "eVLP cGAMP")))

tentbatresults4 <- tentbatresults3 %>%
  left_join(bat_boost_date, by = join_by(id))

str(tentbatresults4)

preboost_tentbatresults <- tentbatresults4 %>%
  filter(date < (Treatment_date + days(21)))

boostday_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(21)) & date <= (Treatment_date + days(22)))

postboostday1_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(22)) & date <= (Treatment_date + days(23)))

postboostday2_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(23)) & date <= (Treatment_date + days(24)))

postboostday3_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday4_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday5_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(25)) & date <= (Treatment_date + days(26)))


preboost_plot <- timeVariation(preboost_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Preboost", "Preboost", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(preboost_plot, subset="hour")

boostday_plot <- timeVariation(boostday_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Boost Day", "Boost Day", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(boostday_plot, subset="hour")

postboostday1_plot <- timeVariation(postboostday1_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 1", "Postboost Day 1", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday1_plot, subset="hour")

postboostday2_plot <- timeVariation(postboostday2_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 2", "Postboost Day 2", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday2_plot, subset="hour")

postboostday3_plot <- timeVariation(postboostday3_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 3", "Postboost Day 3", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday3_plot, subset="hour")

postboostday4_plot <- timeVariation(postboostday4_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 4", "Postboost Day 4", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday4_plot, subset="hour")

postboostday5_plot <- timeVariation(postboostday5_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 5", "Postboost Day 5", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday5_plot, subset="hour")

```


## Quantifying Temperature Differences

Designing a function to measure the daily difference in Bat temperature compared to the tent temperature.

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
```


A function to quantify this delta T by day:

```{r}
tempDifference_fun <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  DMR_Id <- NewTempLoggerMetadata %>%
    filter(Id == mybatID) %>%
    select(DMR_ID) %>%
    pull
  
  Tent <- Metadata %>%
    filter(DMR_ID == DMR_Id) %>%
    select(Tent.x) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`)/144, 
                      daily_batT = mean(`Bat T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 Tent = Tent,
                                 `DeltaT d-3` = daily_summary$daily_delta[1],
                                 `DeltaT d-2` = daily_summary$daily_delta[2],
                                 `DeltaT d-1` = daily_summary$daily_delta[3],
                                 `DeltaT d0` = daily_summary$daily_delta[4],
                                 `DeltaT d1` = daily_summary$daily_delta[5],
                                 `DeltaT d2` = daily_summary$daily_delta[6],
                                 `BatT d-3` = daily_summary$daily_batT[1],
                                 `BatT d-2` = daily_summary$daily_batT[2],
                                 `BatT d-1` = daily_summary$daily_batT[3],
                                 `BatT d0` = daily_summary$daily_batT[4],
                                 `BatT d1` = daily_summary$daily_batT[5],
                                 `BatT d2` = daily_summary$daily_batT[6])
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
tempDifference_fun("1N8476", 3)
bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
bat_delta <- bind_rows(bat_delta_list)
bat_delta_full <- bat_delta |>
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
  rowwise() |>
  mutate(`Preboost Mean dT` = mean(c(`DeltaT d-3`,`DeltaT d-2`,`DeltaT d-1`))) |>
  mutate(`Postboost Mean dT` = mean(c(`DeltaT d0`,`DeltaT d1`,`DeltaT d2`))) |>
  mutate(`Preboost Bat T` = mean (c(`BatT d-3`,`BatT d-2`,`BatT d-1`))) |>
  mutate(`Postboost Bat T` = mean(c(`BatT d0`,`BatT d1`,`BatT d2`)))
bat_delta_full
```

And let's try to look at the various possible models.

```{r Delta T Models}
first_model <- glm(`Postboost Mean dT` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

```{r Delta T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Mean dT`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day Postboost ',Delta,'T'))) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)
bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Mean dT`:`Postboost Mean dT`, names_to = "Boost")
ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day ',Delta,'T')))

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


Well that is quite disappointing. 
Tent is clearly the biggest factor driving this.
Delta T is not the best way to analyze this phenomenon.

First, we need to look at Bat T instead of delta T.


```{r Bat T Models}
first_model <- glm(`Postboost Bat T` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

This is the ideal situation for a mixed effects model!

Let's see how it works.

Following guidance at https://m-clark.github.io/mixed-models-with-R/random_intercepts.html

```{r}
bat_delta_T <- bat_delta_full |>
  mutate(`Boost dT` = `Postboost Bat T` - `Preboost Bat T`)

mixed_model <- lmer(`Boost dT` ~ Treatment + (1|Tent), data = bat_delta_T)
summary(mixed_model)
confint(mixed_model)
ranef(mixed_model)$Tent
coef(mixed_model)$Tent

predictInterval(mixed_model)   # for various model predictions, possibly with new data

REsim(mixed_model)             # mean, median and sd of the random effect estimates

plotREsim(REsim(mixed_model))  # plot the interval estimates

```

It looks like even after incorporating Tent as a random effect, the Treatment effect is not significant.
This may be because I am looking at the 3-day average and I need a more focused approach.

```{r Bat T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Bat T`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)

bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Bat T`:`Postboost Bat T`, names_to = "Boost") |>
  mutate(Boost = factor(Boost, levels = c("Preboost Bat T", "Postboost Bat T")))

ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_pivot, aes(x = Boost, y = value)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5, position = position_nudge()) +
  geom_line(aes(group=BatID), colour="red", linetype="11") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Treatment) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


### Day 1 Post-boost

Reviewing some of the data from the sham group, I think it would be better to look only at post-boost day 1 or day 2.
This should avoid some of the impact of handling. 

```{r}
bat_delta_T_day1 <- bat_delta_full |>
  mutate(`Boost dT day1` = `BatT d1` - `Preboost Bat T`) |>
  mutate(`Boost dT day2` = `BatT d2` - `Preboost Bat T`)

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d1`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d2`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

mixed_model_d1 <- lmer(`Boost dT day1` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d1)
confint(mixed_model_d1)
ranef(mixed_model_d1)$Tent
coef(mixed_model_d1)$Tent
predictInterval(mixed_model_d1)   # for various model predictions, possibly with new data
REsim(mixed_model_d1)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d1))  # plot the interval estimates

mixed_model_d2 <- lmer(`Boost dT day2` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d2)
confint(mixed_model_d2)
ranef(mixed_model_d2)$Tent
coef(mixed_model_d2)$Tent
predictInterval(mixed_model_d2)   # for various model predictions, possibly with new data
REsim(mixed_model_d2)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d2))  # plot the interval estimates
```

The d2 model might show something interesting for the Poly(I:C) group, but it is a small effect.


## Hourly Temperature Differences

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults3 <- singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                    .by = "hour", 
                    hourly_delta = sum(`delta T`)/6)


ggplot(singlebatresults3, aes(x = date, y = hourly_delta, color = Boosted)) + 
  geom_line() +
  theme_cowplot()
```
A function to quantify this delta T by hour:

```{r}
tempDifference_fun2 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 `Preboost d-3` = daily_summary$daily_delta[1]/144,
                                 `Preboost d-2` = daily_summary$daily_delta[2]/144,
                                 `Preboost d-1` = daily_summary$daily_delta[3]/144,
                                 `Boost d0` = daily_summary$daily_delta[4]/144,
                                 `Boost d1` = daily_summary$daily_delta[5]/144,
                                 `Boost d2` = daily_summary$daily_delta[6]/144)
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
# tempDifference_fun2("1N8476", 3)
# bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
# bat_delta <- bind_rows(bat_delta_list)
# bat_delta_full <- bat_delta |>
#   mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
#   rowwise() |>
#   mutate(`Preboost Mean dT` = mean(c(`Preboost d-3`,`Preboost d-2`,`Preboost d-1`))) |>
#   mutate(`Postboost Mean dT` = mean(c(`Boost d0`,`Boost d1`,`Boost d2`)))
# bat_delta_full
```
