---
title: "MOCO 2025 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
library(zoo)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to look at each logger individually to verify that the data is clean.

Then I will plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read_excel("EPFU 2025 BUTT Assignments.xlsx", sheet = "Loggers",  # read in metadata for ibuttons
    range = "A1:M58", na = "NA")

glimpse(Metadata) # check formatting

Metadata$Bat_ID = as.character(Metadata$`Bat Pittag`) # change Bat_ID to a character

glimpse(Metadata) # check formatting again
```


```{r}
DATA_DIR <- "./Individual Bat" # load in each logger file

file_list <- as_tibble(list.files(DATA_DIR)) |>
  mutate(Bat_ID = gsub(".csv", "", value)) |>  #removes .csv
  rename(Filename = value) 

Metadata <- left_join(Metadata, file_list, by = join_by(`Bat Pittag` == Bat_ID)) # adds file name into metadata

batfilelist <- list.files(path = "./Individual Bat", 
                          pattern = "*.csv", full.names = T) # shows full file path name
batfilenames <- list.files(path = "./Individual Bat", 
                           pattern = "*.csv", full.names = F) # shows only file name
batfiles <- lapply(batfilelist, fread, 
                   col.names = c("DateTime","Unit", "Temperature"), sep = "," , 
                   skip = 20) # applies columns to each file, skips the first 20 rows

batnames <- gsub(".csv", "", batfilenames) # removes .csv from file name and leaves only Band 

names(batfiles) <- batnames # names each file by logger name
```


```{r}
allbatresults <- bind_rows(batfiles, .id = "Bat_ID") # creates a long dataset with each time point for every logger, 4 variables 

allbatresults <- allbatresults %>%
  mutate(DateTime = mdy_hms(DateTime)) # changes date to appropriate format

allbatresults <- allbatresults |> 
  mutate(NumberDate = as.numeric(DateTime))
str(allbatresults) # similar to glimpse
summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
md.pattern(allbatresults) # checking for NA's
Metadata %>%
  filter(!Bat_ID%in% file_list$Bat_ID) %>% # only selects loggers with a matching file name
  drop_na(`Filename`) # removes all na's
Metadata %>%
  filter(is.na(`Filename`)) # shows which bats do not have a logger
```

This looks good, I have 44 files and all the metadata with NAs for file name are either ibuttons that are are missing, unused, or lost data. 

Loggers have 2-8000 datapoints. All the ones used for boost have around 4,000 datapoints which makes sense.

## Missing Data

```{r}
clean_met <- Metadata %>%
  drop_na(Filename) |> # removes any ibuttons with no file
  filter(`Experiment D Died` == "D36")

head(clean_met) # look at the top of the dataset
md.pattern(clean_met, rotate.names = TRUE)
write_tsv(clean_met, file = "EPFU2025TempLoggerMetadata.tsv")
```

## Making a plot function

```{r}
plot_fun <- function(arg1) {
  allbatresults %>%
    dplyr::filter(Bat_ID == arg1) %>%
    ggplot(aes(x=DateTime, y=Temperature)) +
    geom_line() +
    ylab("Temperature (C)") + xlab("Date") + ylim(4,40) +
    theme_cowplot() 
}
```

```{r}
plotly_fun <- function(arg1) {
  plot_ly(data = allbatresults[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

## Looking at the data

```{r Bat 03}
ggplotly(plot_fun("APS36067"))
```

```{r Plotly Bat 16}
plotly_fun("APS60113")
```

Here we go!

This works fine, but is not interactive

```{r}
lapply(batnames, plotly_fun)
```

The data loads in well. There are a lot of values that are incorrect (logger never exposed to these temps). Will drop them later.
Will get in to the details about which ones need to be trimmed in a bit. 

```{r}
# here i am just checking the boost data all in one plot to see what it looks like
checking_boost_cal <- allbatresults |> 
  left_join(Metadata, join_by(Bat_ID)) |> # joining with metadata
  filter(`Date Attached` >= '2025-09-29') # filter by loggers attached for second immunization

checking <- allbatresults%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(checking)

```


## Calibration
Before I go any further, I need to calibrate the data. I am using a 2 point calibration from the low and high calibration temps. The formula I am using is CorrectedValue = (((RawValue – RawLow) * ReferenceRange) / RawRange) + ReferenceLow. Before I do that, I need to remove the incorrect values that are over 45 and under 0. Since I know these are incorrect readings and I would want them removed from ALL loggers, I am just going to apply it to the whole dataset.

Unfortunately, my calibration data for the second batch of loggers is really weird and does not make sense. Loggers were reading at 45C while in the hot calibration and 2C while in the cold calibration. The low temperature makes sense because the water bath was in the fridge but the reference thermometer was reading at 11C. The hot temperature does not make sense as the bath should not have been that hot and the reference thermometer read at 35C.

Thankfully, I have some loggers that went in for both calibrations, so I will calibrate all the loggers from the first calibration and then take their temperatures during the second calibration to calibrate the second batch of loggers.

```{r}
# here i am just checking the boost data all in one plot to see what it looks like
checking_boost_cal <- allbatresults |> 
  left_join(Metadata, join_by(Bat_ID)) |> # joining with metadata
  filter(`Date Attached` >= '2025-09-29') # filter by loggers attached for second immunization

p <- checking_boost_cal%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(p)

```

```{r}
clean_results <- allbatresults |> 
  mutate(Temperature = ifelse(Temperature < 0 | Temperature > 50, NA, Temperature)) # changes temps over 50 and under 0 to na

clean_results <- clean_results |> 
  # 2. Sort chronologically by bat and time
  arrange(Bat_ID, DateTime) |> 
  # 3. Group by Bat so you don't interpolate between different animals
  group_by(Bat_ID) |> 
  # 4. Fill the gaps (only if 1-3 consecutive points are missing)
  mutate(fixed_temp = na.approx(Temperature, maxgap = 3, na.rm = FALSE)) |> 
  # 5. Clean up grouping
  ungroup() |> 
  as.data.table()

summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
summary(as.factor(clean_results$Bat_ID)) # comparing counts between the two
```
This looks good. Now I am going to add columns to the dataframe with the maximum and minimum values during calibration. There were two different calibrations, so first I have to select loggers only used in the first calibration. 

### Calibration 1

```{r}
Clean_results_met <- clean_results %>% 
  left_join(clean_met, by = "Bat_ID") # join metadata with results so I can extract loggers in second calibration

Cal_1_results <- Clean_results_met |>  # extracting second calibration results
  filter(Calibration == 1) |> 
  select(Bat_ID, DateTime, Unit, fixed_temp, NumberDate) |> 
  rename(Temperature = "fixed_temp")

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[NumberDate >= 1756687201 & 
                  DateTime <= 1756692601 ], na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[NumberDate >= 1756738201 & 
                     NumberDate <= 1756801801], na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range from average high temp and average low temp
  ungroup()  
```

Okay, now because Ken would want me to, I am going to load in the calibration data and get the range and average low. 

```{r}
Reference_cal_1 <- read_excel("Calibration Data/31-Aug_Calibration.xlsx", 
    range = "A1:D8869", na = "NA") #loading in the calibration 2 reference data

summary(Reference_cal_1) # looking at it to make sure it loaded in correctly

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(lowT = mean(T1[No.s >= 1076 & 
                  No.s <= 1433],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. I needed to use the time point because the EA15 does not record date so it loaded in as 1899-12-31 for all times. The time points lines up with the same time used in averaging low and high logger temperatures.

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(highT = mean(T1[No.s >= 4473 & 
                  No.s <= 8713],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. 
```

The average low is 4.554749, the average high is 36.97654, and the range is 32.42. 
Okay, now I am going to create a column in my cal_1_results that has the formula for the corrected values. 

```{r}
Cal_1_results <- Cal_1_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 32.42)/ RangeT) + 4.55) # applying formula listed above to the calibration to get a corrected temperature value

summary(Cal_1_results)
```
I think this data looks good. I took a look at the dataset and saw that the range is different, the low and high looks like they were calculated correctly. 


### Calibration 2
Okay, so for this one, I am going to take the three loggers that went through both calibrations. Using the first calibration, the data from the second calibration should be true values. I am going to take these temperatures during the second calibration and use them to calibrate the second batch. 
```{r}
Cal_2_results <- Clean_results_met |> # extracting second calibration results
  filter(Calibration == 2) |> 
  select(Bat_ID, DateTime, Unit, fixed_temp, NumberDate) |> 
  rename(Temperature = "fixed_temp")

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[NumberDate >= 1759106401 & 
                  NumberDate <= 1759112401], na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[NumberDate >= 1759125601 & 
                  NumberDate <= 1759130401], na.rm = TRUE)) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range and add column
  ungroup()  
```

Okay, now I am going to take Cal_1_results and filter out the three loggers in both calibrations

```{r}
Reference_cal_2 <- Cal_1_results |> 
  filter(Bat_ID %in% c("APS36010_boost", "APS36057_boost", "APS60113_boost") ) |> 
  filter(NumberDate  > 1759104001, 
         NumberDate < 1759132801)

```

I need to make sure they calibrations are all reading about the same.
```{r}

ref_cal_2 <- Reference_cal_2%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = corrected, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(ref_cal_2)

```
Okay, I think this looks good

```{r}
summary(Reference_cal_2)

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(lowT = mean(corrected[NumberDate >= 1759106401 & 
                  NumberDate <= 1759112401]))  #taking the average temperature during the low calibration for each logger

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(highT = mean(corrected[NumberDate >= 1759125601 & 
                  NumberDate <= 1759130401])) 

summary(Reference_cal_2)
```

The average low is 4.733, the average high is 49.46, and the range is 44.73. 
Okay, now I am going to create a column in my cal_1_results that has the formula for the corrected values. 

```{r}
Cal_2_results <- Cal_2_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 44.73)/ RangeT) + 4.733)

summary(Cal_2_results)
```

Okay, now that I have my data calibrated, I am going to merge the two datasets.

```{r}
cal_data <- bind_rows(Cal_1_results, Cal_2_results) |> 
  select(Bat_ID, DateTime, NumberDate, corrected) |> 
  rename("Temperature" = "corrected")

```

## Left and Right Trimming the data

I will split the data, filter it, and then recombine them.

```{r}
# for analysis down below, I cannot use DateTime data so I need to convert my time into a numeric 
clean_met <- clean_met %>%
  mutate(Attachment_Date = as.numeric(`Date Attached`),
         Removal_Date = as.numeric(`Date Removed`)) 

temps_with_meta <- cal_data %>%
  left_join(clean_met, by = "Bat_ID") #joining my calibrated data with my metadata

# Filter to keep only records after attachment date
left_trimmed <- temps_with_meta %>%
  filter(NumberDate >= Attachment_Date)

# Filter to keep only records before removal date
right_trimmed <- left_trimmed |> 
  filter(NumberDate <= Removal_Date)

# Making a new data table with trimmed data 
allbatresults2 <- right_trimmed |> 
  as.data.table()

#checking the data
summary(as.factor(allbatresults2$Bat_ID))

glimpse(allbatresults2)

```

```{r}
plotly_fun2 <- function(arg1) {
  plot_ly(data = allbatresults2[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(batnames, plotly_fun2)
```
## Left trim again

I left trimmed earlier to remove all data from before the day of attachment. I really want to include as much as possible so I am trimming again by the attachment time.

```{r}

# first trim
Trimming1 <- allbatresults2 |> 
  filter(Attachment_Date == 1756771200) |> 
  filter(NumberDate >= 1756826401)
summary(as.factor(Trimming1$Bat_ID))

#second trim

Trimming2 <- allbatresults2 |> 
  filter(Attachment_Date == 1757116800) |> 
  filter(NumberDate >= 1757170201)
summary(as.factor(Trimming2$Bat_ID))

#third trim

Trimming3 <- allbatresults2 |> 
  filter(Attachment_Date == 1757376000) |> 
  filter(NumberDate >= 1757439601)
summary(as.factor(Trimming3$Bat_ID))

#fourth trim

Trimming4 <- allbatresults2 |> 
  filter(Attachment_Date == 1759104000) |> 
  filter(NumberDate >= 1759152601)
summary(as.factor(Trimming4$Bat_ID))

#fifth trim

Trimming5 <- allbatresults2 |> 
  filter(Attachment_Date == 1759363200) |> 
  filter(NumberDate >= 1759420201)
summary(as.factor(Trimming5$Bat_ID))

LTrimmed <- bind_rows(Trimming1, Trimming2, Trimming3, Trimming4, Trimming5)

```

## Right trim again 

```{r}
plotly_fun3 <- function(arg1) {
  plot_ly(data = LTrimmed[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }
```

```{r}
lapply(batnames, plotly_fun3)
```

Okay, this is looking good. I do have some more data to trim. So I am going to do that now.
```{r}
# Right trim 1
RTrimming1 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36008")) |> 
  filter(!NumberDate > 1760539201)
summary(as.factor(RTrimming1$Bat_ID))

# Right trim 2
RTrimming2 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36021")) |> 
  filter(!NumberDate > 1759998001)
summary(as.factor(RTrimming2$Bat_ID))

# Right trim 3

RTrimming3 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36026")) |> 
  filter(!NumberDate > 1760208601)
summary(as.factor(RTrimming3$Bat_ID))

# Right Trim 4 

RTrimming4 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36057_boost")) |> 
  filter(!NumberDate > 1760530201)
summary(as.factor(RTrimming4$Bat_ID))

# RTrim 5

RTrimming5 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36108")) |> 
  filter(!NumberDate > 1760455801)
summary(as.factor(RTrimming5$Bat_ID))

# RTrim 6

RTrimming6 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60081")) |> 
  filter(!NumberDate > 1758842401)
summary(as.factor(RTrimming6$Bat_ID))

# RTrim 7

RTrimming7 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60115")) |> 
  filter(!NumberDate > 1757681401)
summary(as.factor(RTrimming7$Bat_ID))

# RTrim 8

RTrimming8 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36007")) |> 
  filter(!NumberDate > 1759327201)
summary(as.factor(RTrimming8$Bat_ID))

# RTrim 9

RTrimming9 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36106")) |> 
  filter(!NumberDate > 1759015201)
summary(as.factor(RTrimming9$Bat_ID))

# RTrim 10

RTrimming10 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60113")) |> 
  filter(!NumberDate > 1759327201)
summary(as.factor(RTrimming10$Bat_ID))

#Keeping all results from LTrimmed besides the ones that were right trimmed again

RTrimmed_Temp <- LTrimmed |> 
  filter(!Bat_ID %in% c( "APS36008", "APS36021", "APS36026", "APS36057_boost", "APS36108", "APS60081", "APS60115", "APS36007", "APS36106", "APS60113"))

Trimmed <- bind_rows(RTrimmed_Temp, RTrimming1, RTrimming2, RTrimming3, RTrimming4, RTrimming5, RTrimming6, RTrimming7, RTrimming8, RTrimming9, RTrimming10)

summary(as.factor(Trimmed$Bat_ID))

#last I am going to drop APS36007 boost. The data is ALL over the place so we just can't trust these readings

Trimmed <- Trimmed |> 
  filter(!Bat_ID == 'APS36007_boost')

```

Now I want to look at the data and make sure it looks right and see what else needs to be trimmed

```{r}
cleaning_names <- unique(Trimmed$Bat_ID)


plotly_fun4 <- function(arg1) {
  plot_ly(data = Trimmed[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(cleaning_names, plotly_fun4)
```
This looks pretty good to me. The last thing I am going to do is drop data above 40 and below 13. When I calculate area under the curve later, I don't want to everything to be messed up though. So what I am going to do is for data that is dropped, instead of leaving it blank I am going to take the average between the points. For some bats I have huge chunks of data that I think will be dropped. For these chunks it would be hard to take the averages so I am going to have a cut off that if there are more than 3 na's in a row, the data gets dropped.

```{r}
final_results <- Trimmed |>
  mutate(Temperature = ifelse(Temperature < 13 | Temperature > 40, NA, Temperature)) 
```

```{r}

yyyy <- Trimmed %>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(yyyy)

```

```{r}
final_results <- final_results |> 
  # 2. Sort chronologically by bat and time
  arrange(Bat_ID, DateTime) |> 
  # 3. Group by Bat so you don't interpolate between different animals
  group_by(Bat_ID) |> 
  # 4. Fill the gaps (only if 1-3 consecutive points are missing)
  mutate(fixed_temp = na.approx(Temperature, maxgap = 3, na.rm = FALSE)) |> 
  # 5. Clean up grouping
  ungroup() |> 
  as.data.table()
```

```{r}
plotly_fun5 <- function(arg1) {
  plot_ly(data = final_results[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```

```{r}
lapply(cleaning_names, plotly_fun5)
```

```{r}
# write_tsv(clean_results, file = "TrimmedTemperatureLoggerData.long.tsv")
```

## Renaming 
Before I go any further, I want to rename some of my Bat Ids that have boost in the name 

```{r}
summary(as.factor(final_results$Bat_ID))

APS36010 <- final_results |> 
  filter(Bat_ID == "APS36010_boost")
APS36010$Bat_ID <- c("APS36010")

APS36057 <- final_results |> 
  filter(Bat_ID == "APS36057_boost")
APS36057$Bat_ID <- c("APS36057")

APS60079 <- final_results |> 
  filter(Bat_ID == "APS60079_boost")
APS60079$Bat_ID <- c("APS60079")

APS60081 <- final_results |> 
  filter(Bat_ID == "APS60081_boost")
APS60081$Bat_ID <- c("APS60081")

APS60088 <- final_results |> 
  filter(Bat_ID == "APS60088_boost")
APS60088$Bat_ID <- c("APS60088")

APS60113 <- final_results |> 
  filter(Bat_ID == "APS60113_boost")
APS60113$Bat_ID <- c("APS60113")

APS60115 <- final_results |> 
  filter(Bat_ID == "APS60115_boost")
APS60115$Bat_ID <- c("APS60115")

APS60123 <- final_results |> 
  filter(Bat_ID == "APS60123_boost")
APS60123$Bat_ID <- c("APS60123")

APS60156 <- final_results |> 
  filter(Bat_ID == "APS60156_boost")
APS60156$Bat_ID <- c("APS60156")

APS36106 <- final_results |> 
  filter(Bat_ID == "APS36106_boost")
APS36106$Bat_ID <- c("APS36106")

renaming_results <- final_results |> 
  filter(!Bat_ID %in% c ("APS60156_boost", "APS60123_boost", "APS60115_boost", "APS60113_boost", "APS60088_boost", "APS60081_boost", "APS60079_boost", "APS36057_boost", "APS36010_boost",  "APS36106_boost"))

Bat_results_met <- bind_rows(renaming_results, APS36010, APS36057, APS60079, APS60081, APS60088, APS60113, APS60115, APS60123, APS60156, APS36106)
```

## Ambient Temp
I do not have ambient data from the flight cage, but I do have ambient data and relative humidity from the thermal challenges. There were two sensor pushes for each thermal challenge. I am going to take the average of the two. 

### Basal TC
```{r}
Basal_Ambient1 <- read.csv("Ambient/09.04.2025_basaltempdata1.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Basal_Ambient2 <- read.csv("Ambient/09.04.2025_basaltempdata2.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Basal_working <- Basal_Ambient1 |> 
  left_join(Basal_Ambient2, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

Basal_working <- Basal_working |> 
  mutate(avg_temp = rowMeans(pick("Temperature.x", "Temperature.y"))) |> 
  mutate(avg_RH = rowMeans(pick("Humidity..x", "Humidity..y"))) 

Basal_Ambient <- Basal_working |>  
  select(Shift_date.x, NumberDate, avg_temp, avg_RH)  |> # keeping certain columns 
  rename(DateTime = "Shift_date.x") |>  # renaming columns 
  filter(NumberDate >= 1757001841 &
           NumberDate <= 1757081281)
```

### Primary TC Group 1

```{r}
G1D0_Ambient1 <- read.csv("Ambient/09.08.2025_G1D0_1.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G1D0_Ambient2 <- read.csv("Ambient/09.08.2025_G1D0_2.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G1D0_working <- G1D0_Ambient1|> 
  left_join(G1D0_Ambient2, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

G1D0_working <- G1D0_working |> 
  mutate(avg_temp = rowMeans(pick("Temperature.x", "Temperature.y"))) |> 
  mutate(avg_RH = rowMeans(pick("Humidity..x", "Humidity..y"))) 

G1D0_Ambient <- G1D0_working |>  
  select(Shift_date.x, NumberDate, avg_temp, avg_RH)  |> # keeping certain columns 
  rename(DateTime = "Shift_date.x") |>  # renaming columns 
  filter(NumberDate >= 1757344141 &
           NumberDate <= 1757423341)
```

### Primary TC Group 2

```{r}
G2D0_Ambient1 <- read.csv("Ambient/09.11.2025_G2D0_1.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G2D0_Ambient2 <- read.csv("Ambient/09.11.2025_G2D0_2.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G2D0_working <- G2D0_Ambient1|> 
  left_join(G2D0_Ambient2, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

G2D0_working <- G2D0_working |> 
  mutate(avg_temp = rowMeans(pick("Temperature.x", "Temperature.y"))) |> 
  mutate(avg_RH = rowMeans(pick("Humidity..x", "Humidity..y"))) 

G2D0_Ambient <- G2D0_working |>  
  select(Shift_date.x, NumberDate, avg_temp, avg_RH)  |> # keeping certain columns 
  rename(DateTime = "Shift_date.x") |>  # renaming columns 
  filter(NumberDate >= 1757602681 &
           NumberDate <= 1757681881)
```

I only have humidity for one sensor push for both secondary thermal challenges

### Secondary TC Group 1
```{r}
G1D21_Ambient1 <- read.csv("Ambient/09.29.2025_G1D21_1.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G1D21_Ambient2 <- read.csv("Ambient/09.29.2025_G1D21_2.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G1D21_working <- G1D21_Ambient1|> 
  left_join(G1D21_Ambient2, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

G1D21_working <- G1D21_working |> 
  mutate(avg_temp = rowMeans(pick("Temperature.x", "Temperature.y")))

G1D21_Ambient <- G1D21_working |>  
  select(Shift_date.x, NumberDate, avg_temp,Humidity.)  |> # keeping certain columns
  rename(DateTime = "Shift_date.x") |>
  rename(avg_RH = "Humidity.") |> # renaming columns 
  filter(NumberDate >= 1759155061 &
           NumberDate <= 1759234201)
```

### Secondary TC Group 2
```{r}
G2D21_Ambient1 <- read.csv("Ambient/10.02.2025_G2D21_1.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G2D21_Ambient2 <- read.csv("Ambient/10.02.2025_G2D21_2.csv") |>  # read in temp from top of incubator from D0 thermal challenge |> 
  mutate(DateTime = mdy_hm(DateTime)) |> 
  mutate(Shift_date = DateTime + seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

G2D21_working <- G2D21_Ambient1|> 
  left_join(G2D21_Ambient2, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

G2D21_working <- G2D21_working |> 
  mutate(avg_temp = rowMeans(pick("Temperature.x", "Temperature.y")))

G2D21_Ambient <- G2D21_working |>  
  select(Shift_date.x, NumberDate, avg_temp,Humidity.)  |> # keeping certain columns
  rename(DateTime = "Shift_date.x") |>
  rename(avg_RH = "Humidity.") |> # renaming columns 
  filter(NumberDate >= 1759417921 &
           NumberDate <= 1759497061)
```

Okay, now that I have all that sorted out. I am going to combine the temperature data with the metadata so I can get the match the ambient TC temp with the bats temp data.

```{r}
TC_results <- Bat_results_met |> 
  filter(Treatment %in% c("Primary", "Secondary", "Dual")) # keep only thermally challenged bats

working_Basal_results <- TC_results |> 
  left_join(Basal_Ambient, join_by(NumberDate)) |> # combining with basal ambient data
  drop_na(avg_temp) #needed to drop na's 
  
Basal_TC_results <- working_Basal_results |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, avg_temp, avg_RH) |> 
  rename(DateTime = "DateTime.x") |> 
  mutate(deltaT = Temperature - avg_temp )
```

**** Come back to this later ******
```{r}

#basal_tc  <- Basal_TC_results%>%
#  dplyr::filter(Bat_ID %in% batnames) %>%
#  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
#  geom_line() +
#  facet_grid()
#  ylim(0, 50) +
#  theme_minimal() +
#  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")
#
## Convert to interactive
#ggplotly(ref_cal_2)

```
### Shifting Data

We had two cohorts of bats that followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. The date is not very interesting to me, so I want to shift them so they all fall on the same project date.

```{r}
Bat_results_met$time_aligned <- Bat_results_met$DateTime

Bat_results_met$time_aligned[Bat_results_met$Group == "GP 2"] <- Bat_results_met$time_aligned[Bat_results_met$Group == "GP 2"] - lubridate::days(3)

all_results_met <- Bat_results_met |> 
  select(!c(`Logger location`, NumberDate, DateTime)) |> 
  rename(DateTime = "time_aligned") |> 
  mutate(NumberDate = as.numeric(DateTime))

```

```{r}
working_D0G1 <- all_results_met |> 
  left_join(G1D0_Ambient, join_by(NumberDate)) |> 
  filter(Treatment %in% c("Primary", "Dual")) |> 
  drop_na(avg_temp)

G1D0_results <- working_D0G1 |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, avg_temp, avg_RH) |>  rename(DateTime = "DateTime.x") |> 
  mutate(deltaT = Temperature - avg_temp )
  
working_D0G2 <- all_results_met |> 
  left_join(G2D0_Ambient, join_by(NumberDate)) |> 
  filter(Treatment %in% c("Primary", "Dual")) |> 
  drop_na(avg_temp)

G2D0_results <- working_D0G2 |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, avg_temp, avg_RH) |>  rename(DateTime = "DateTime.x") |> 
  mutate(deltaT = Temperature - avg_temp )

Primary_TC_results <- bind_rows(G1D0_results, G2D0_results)
```

```{r}
working_D21G1 <- all_results_met |> 
  left_join(G1D21_Ambient, join_by(NumberDate)) |> 
  filter(Treatment %in% c("Secondary", "Dual")) |> 
  drop_na(avg_temp)

G1D21_results <- working_D21G1 |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, avg_temp, avg_RH) |>  rename(DateTime = "DateTime.x") |> 
  mutate(deltaT = Temperature - avg_temp )
  
working_D21G2 <- all_results_met |> 
  left_join(G2D21_Ambient, join_by(NumberDate)) |> 
  filter(Treatment %in% c("Secondary", "Dual")) |> 
  drop_na(avg_temp)

G2D21_results <- working_D21G2 |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, avg_temp, avg_RH) |>  rename(DateTime = "DateTime.x") |> 
  mutate(deltaT = Temperature - avg_temp )

Secondary_TC_results <- bind_rows(G1D21_results, G2D21_results)
```

```{r}
Bat_results <- all_results_met |> 
  select(Bat_ID, DateTime, NumberDate, Temperature)

Bat_Metadata_temp <- all_results_met |> 
  select(Bat_ID, `Bat Band`, Treatment, Group, Sex)

Bat_Metadata <- Bat_Metadata_temp[!duplicated(Bat_Metadata_temp$Bat_ID)]
```

```{r}
Bat_Primary <- Bat_results |> 
  filter(NumberDate >= 1757181601 & 
          NumberDate < 1757808001)

Bat_Secondary <- Bat_results |> 
  filter(NumberDate >= 1758844801 &
          NumberDate < 1759536001)
```
 
I shifted the data so now data all shows d21 on June 5th, the same day as D21 for Group 1. 
I think I have all the data wrangling down so I am going to export some tsv with my final primary data, secondary data, metadata for both immunizations, and ambient temperature.


Okay, now I am going to remove the days when we handled the bats so I can have some clean circadian rhythm data
```{r}

GP_1_cleaned <- all_results_met |> 
  filter(Group == "GP 1") |> 
  filter(!NumberDate <= 1757181601 | !NumberDate >= 1756987201) |> 
  filter(!NumberDate <= 1757694601 | !NumberDate >= 1757329201) |> 
  filter(!NumberDate <= 1759509001 | !NumberDate >= 1759147201) |> 
  filter(!NumberDate <= 1759838401| !NumberDate >= 1759752001)

GP_2_cleaned <- all_results_met |> 
  filter(Group == "GP 2") |> 
  filter(!NumberDate <= 1756923001 | !NumberDate >= 1756728001) |> 
  filter(!NumberDate <= 1757694601 | !NumberDate >= 1757329201) |> 
  filter(!NumberDate <= 1759509001 | !NumberDate >= 1759147201) |> 
  filter(!NumberDate <= 1759838401| !NumberDate >= 1759752001)

Results_Cleaned <- bind_rows(GP_1_cleaned, GP_2_cleaned) |> 
  select(Bat_ID, DateTime, NumberDate, Temperature)
```

Okay, I just want to make sure this worked correctly so I am going to plot it all. 

```{r}
final_names <- unique(Bat_results$Bat_ID)

check <- Results_Cleaned%>%
  dplyr::filter(Bat_ID %in% final_names) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(10, 40) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(check)

```

## Experiment Day Labeling

```{r}
Bat_results_met$time_aligned[Bat_results_met$Group == "GP 2"] <- Bat_results_met$time_aligned[Bat_results_met$Group == "GP 2"] - lubridate::days(3)

Results_Analysis <- all_results_met |> 
  mutate(DateType = "")

day_0_anchor <- ymd_hms("2025-09-08 14:00:01") 

Results_Analysis <- Results_Analysis |> 
  mutate(Days_From_Anchor = as.numeric(difftime(DateTime, day_0_anchor, units = "days")),
    # 4. Use floor() so that Day 0 spans from Sept 9 2pm to Sept 10 1:59pm
    Day_Number = floor(Days_From_Anchor),
      # 5. Format it to look like "D0", "D1", "D-1", etc.
    ExperimentDay = case_when(
      Day_Number == 0  ~ "D0",
      Day_Number > 0   ~ paste0("D", Day_Number),
      Day_Number < 0   ~ paste0("D", Day_Number) # Automatically handles the negative sign (e.g., "D-1")
    ))
```


Looks incredible ugly but we get the point. I want to add a column that says what each day of the experiment is (Primary immunization, basal TC, etc.). I will also add another column that is handled vs not handled, and another that is experiment day (D21, D22)

```{r}

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment == "NoCo" & NumberDate %in% c(1757253601:1757339401) ~ "Primary - 1",
      Treatment == "NoCo" & NumberDate %in% c(1757340001:1757425801) ~ "Primary",
      Treatment == "NoCo" & NumberDate %in% c(1757426401: 1757512201) ~ "Primary + 1",
      Treatment == "NoCo" & NumberDate %in% c(1757512801: 1757598601) ~ "Primary + 2",
      Treatment == "NoCo" & NumberDate %in% c(1757599201: 1757685001) ~ "Primary + 3",
      Treatment == "NoCo" & NumberDate %in% c(1759068001: 1759153801) ~ "Secondary - 1",
      Treatment == "NoCo" & NumberDate %in% c(1759154401: 1759240201) ~ "Secondary",
      Treatment == "NoCo" & NumberDate %in% c(1759240801: 1759326601) ~ "Secondary + 1",
      Treatment == "NoCo" & NumberDate %in% c(1759327201: 1759413001) ~ "Secondary + 2",
      Treatment == "NoCo" & NumberDate %in% c(1759413601: 1759499401) ~ "Secondary + 3",
      Treatment == "NoCo" & NumberDate %in% c(1759759201: 1759845001) ~ "D28",
      Treatment == "NoCo" & Day_Number < 0 ~ "Pre",
      Treatment == "NoCo" ~ "None", 
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment == "SHAM" & NumberDate %in% c(1757253601:1757339401) ~ "Primary - 1",
      Treatment == "SHAM" & NumberDate %in% c(1757340001:1757425801) ~ "Primary",
      Treatment == "SHAM" & NumberDate %in% c(1757426401: 1757512201) ~ "Primary + 1",
      Treatment == "SHAM" & NumberDate %in% c(1757512801: 1757598601) ~ "Primary + 2",
      Treatment == "SHAM" & NumberDate %in% c(1757599201: 1757685001) ~ "Primary + 3",
      Treatment == "SHAM" & NumberDate %in% c(1759068001: 1759153801) ~ "Secondary - 1",
      Treatment == "SHAM" & NumberDate %in% c(1759154401: 1759240201) ~ "Secondary",
      Treatment == "SHAM" & NumberDate %in% c(1759240801: 1759326601) ~ "Secondary + 1",
      Treatment == "SHAM" & NumberDate %in% c(1759327201: 1759413001) ~ "Secondary + 2",
      Treatment == "SHAM" & NumberDate %in% c(1759413601: 1759499401) ~ "Secondary + 3",
      Treatment == "SHAM" & NumberDate %in% c(1759759201: 1759845001) ~ "D28",
      Treatment == "SHAM" & Day_Number < 0 ~ "Pre",
      Treatment == "SHAM" ~ "None", 
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment == "Dual" & NumberDate %in% c(1757253601:1757339401) ~ "Primary - 1",
      Treatment == "Dual" & NumberDate %in% c(1757340001:1757425801) ~ "Primary",
      Treatment == "Dual" & NumberDate %in% c(1757426401: 1757512201) ~ "Primary + 1",
      Treatment == "Dual" & NumberDate %in% c(1757512801: 1757598601) ~ "Primary + 2",
      Treatment == "Dual" & NumberDate %in% c(1757599201: 1757685001) ~ "Primary + 3",
      Treatment == "Dual" & NumberDate %in% c(1759068001: 1759153801) ~ "Secondary - 1",
      Treatment == "Dual" & NumberDate %in% c(1759154401: 1759240201) ~ "Secondary",
      Treatment == "Dual" & NumberDate %in% c(1759240801: 1759326601) ~ "Secondary + 1",
      Treatment == "Dual" & NumberDate %in% c(1759327201: 1759413001) ~ "Secondary + 2",
      Treatment == "Dual" & NumberDate %in% c(1759413601: 1759499401) ~ "Secondary + 3",
      Treatment == "Dual" & NumberDate %in% c(1759759201: 1759845001) ~ "D28",
      Treatment == "Dual" & Day_Number < 0 ~ "Pre",
      Treatment == "Dual" ~ "None", 
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment == "Primary" & NumberDate %in% c(1757253601:1757339401) ~ "Primary - 1",
      Treatment == "Primary" & NumberDate %in% c(1757340001:1757425801) ~ "Primary",
      Treatment == "Primary" & NumberDate %in% c(1757426401: 1757512201) ~ "Primary + 1",
      Treatment == "Primary" & NumberDate %in% c(1757512801: 1757598601) ~ "Primary + 2",
      Treatment == "Primary" & NumberDate %in% c(1757599201: 1757685001) ~ "Primary + 3",
      Treatment == "Primary" & NumberDate %in% c(1759068001: 1759153801) ~ "Secondary - 1",
      Treatment == "Primary" & NumberDate %in% c(1759154401: 1759240201) ~ "Secondary",
      Treatment == "Primary" & NumberDate %in% c(1759240801: 1759326601) ~ "Secondary + 1",
      Treatment == "Primary" & NumberDate %in% c(1759327201: 1759413001) ~ "Secondary + 2",
      Treatment == "Primary" & NumberDate %in% c(1759413601: 1759499401) ~ "Secondary + 3",
      Treatment == "Primary" & NumberDate %in% c(1759759201: 1759845001) ~ "D28",
      Treatment == "Primary" & Day_Number < 0 ~ "Pre",
      Treatment == "Primary" ~ "None", 
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment == "Secondary" & NumberDate %in% c(1757253601:1757339401) ~ "Primary - 1",
      Treatment == "Secondary" & NumberDate %in% c(1757340001:1757425801) ~ "Primary",
      Treatment == "Secondary" & NumberDate %in% c(1757426401: 1757512201) ~ "Primary + 1",
      Treatment == "Secondary" & NumberDate %in% c(1757512801: 1757598601) ~ "Primary + 2",
      Treatment == "Secondary" & NumberDate %in% c(1757599201: 1757685001) ~ "Primary + 3",
      Treatment == "Secondary" & NumberDate %in% c(1759068001: 1759153801) ~ "Secondary - 1",
      Treatment == "Secondary" & NumberDate %in% c(1759154401: 1759240201) ~ "Secondary",
      Treatment == "Secondary" & NumberDate %in% c(1759240801: 1759326601) ~ "Secondary + 1",
      Treatment == "Secondary" & NumberDate %in% c(1759327201: 1759413001) ~ "Secondary + 2",
      Treatment == "Secondary" & NumberDate %in% c(1759413601: 1759499401) ~ "Secondary + 3",
      Treatment == "Secondary" & NumberDate %in% c(1759759201: 1759845001) ~ "D28",
      Treatment == "Secondary" & Day_Number < 0 ~ "Pre",
      Treatment == "Secondary" ~ "None", 
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 1" & NumberDate %in% c(1756908001:1756993801) ~ "Basal TC-1",
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 1" & NumberDate %in% c(1756994401:1757080201) ~ "Basal TC",
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 1" & NumberDate %in% c(1757080801:1757166601) ~ "Basal TC + 1",
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis|> 
  mutate(DateType = case_when(
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 2" & NumberDate %in% c(1756648801:1756734601) ~ "Basal TC-1",
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 2" & NumberDate %in% c(1756735201:1756821001) ~ "Basal TC",
      Treatment %in% c("Secondary", "Dual", "Primary") & Group == "GP 2" & NumberDate %in% c(1756821601:1756907401) ~ "Basal TC + 1",
      TRUE ~ DateType))

Results_Analysis <- Results_Analysis |> 
  mutate(Handled = if_else(DateType == "None", "No", "Yes"))
```
Okay got that done! Now I want to add another column that is experiment date. 

```{r}

Results_Analysis <- Results_Analysis |> 
  mutate(Cage = case_when(
    Treatment %in% c("SHAM", "NoCo") ~ "None",
    Treatment %in% c("Dual", "Primary") & DateType == "Primary" & Group == "GP 1" ~ "D+P1",
    Treatment %in% c("Dual", "Primary") & DateType == "Primary" & Group == "GP 2" ~ "D+P2",
    Treatment %in% c("Dual", "Secondary") & DateType == "Secondary" & Group == "GP 1" ~ "D+S1",
    Treatment %in% c("Dual", "Secondary") & DateType == "Secondary" & Group == "GP 2" ~ "D+S2",
    Treatment %in% c("Dual") & DateType == "Basal TC" ~ "Dual",
    Treatment %in% c("Primary") & DateType == "Basal TC" ~ "Primary",
    Treatment %in% c("Secondary") & DateType == "Basal TC" ~ "Secondary",
    TRUE ~ "None"))

Results <- Results_Analysis |> 
  select(Bat_ID, DateTime, NumberDate, Temperature, DateType,Treatment, Group, Sex, Handled, ExperimentDay, Cage)
```

## Weight Data

```{r}
Weight_Data <- read_excel("EPFU_Interpolated_Weights.xlsx", # read in weight data
    range = "A1:C1744", na = "NA")

glimpse(Weight_Data)

Weight_Data <- Weight_Data |> 
  # 1. Create a temporary numeric column for correct chronological sorting
  mutate(Day_Num = as.numeric(str_remove(ExperimentDay, "D"))) |> 
  
  # 2. Sort chronologically by bat and the actual numeric day
  arrange(Bat_ID, Day_Num) |> 
  
  # 3. Group by Bat so you don't interpolate between different animals
  group_by(Bat_ID) |> 
  
  # 4. Fill the gaps (linear interpolation)
  mutate(Weight = na.approx(Weight, na.rm = FALSE)) |> 
  
  # 5. Clean up grouping and remove the temporary column
  ungroup() |>

  # 6. Convert to data.table
  as.data.table()


Results <- Results |> 
  left_join(Weight_Data |> select(Bat_ID, ExperimentDay, Weight), join_by(Bat_ID,ExperimentDay))

```

```{r}

final_check <- Results%>%
  dplyr::filter(Bat_ID %in% final_names) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(10, 45) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(final_check)

```

Okay, one last thing this time... We changed my treatment names from Secondary to CP, Primary to CS, and Dual to CP+S. I need to change these in the dataset, but I did not want to have to go through my whole code to change it so I'll just do it here. 

```{r}
Bat_Metadata$Treatment[Bat_Metadata$Treatment == "Secondary"] <- "CS"
Bat_Metadata$Treatment[Bat_Metadata$Treatment == "Primary"] <- "CP"
Bat_Metadata$Treatment[Bat_Metadata$Treatment == "Dual"] <- "CP+S"

Results$Treatment[Results$Treatment == "Secondary"] <- "CS"
Results$Treatment[Results$Treatment == "Primary"] <- "CP"
Results$Treatment[Results$Treatment == "Dual"] <- "CP+S"
```

```{r}
write_tsv(Bat_Secondary, file = "EPFU2025SecondaryTemperature.tsv")
write_tsv(Bat_Primary, file = "EPFU2025PrimaryTemperature.tsv")
write_tsv(Bat_Metadata, file = "EPFU2025Metadata.tsv")
write_tsv(Bat_results, file = "EPFU2025BatResults.tsv")
write_tsv(Results_Cleaned, file = "EPFU2025Cleaned_Results.tsv")
write_tsv(Primary_TC_results, file = "EPFU2025Primary_TC.tsv")
write_tsv(Basal_TC_results, file = "EPFU2025Basal_TC.tsv")
write_tsv(Secondary_TC_results, file = "EPFU2025Secondary_TC.tsv")
write_tsv(Results, file = "EPFU2025Results_Analysis.tsv")

```


