---
title: "MOCO 2025 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to look at each logger individually to verify that the data is clean.

Then I will plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read_excel("EPFU 2025 BUTT Assignments.xlsx", sheet = "Loggers",  # read in metadata for ibuttons
    range = "A1:M58", na = "NA")

glimpse(Metadata) # check formatting

Metadata$Bat_ID = as.character(Metadata$`Bat Pittag`) # change Bat_ID to a character

glimpse(Metadata) # check formatting again
```


```{r}
DATA_DIR <- "./Individual Bat" # load in each logger file

file_list <- as_tibble(list.files(DATA_DIR)) |>
  mutate(Bat_ID = gsub(".csv", "", value)) |>  #removes .csv
  rename(Filename = value) 

Metadata <- left_join(Metadata, file_list, by = join_by(`Bat Pittag` == Bat_ID)) # adds file name into metadata

batfilelist <- list.files(path = "./Individual Bat", 
                          pattern = "*.csv", full.names = T) # shows full file path name
batfilenames <- list.files(path = "./Individual Bat", 
                           pattern = "*.csv", full.names = F) # shows only file name
batfiles <- lapply(batfilelist, fread, 
                   col.names = c("DateTime","Unit", "Temperature"), sep = "," , 
                   skip = 20) # applies columns to each file, skips the first 20 rows

batnames <- gsub(".csv", "", batfilenames) # removes .csv from file name and leaves only Band 

names(batfiles) <- batnames # names each file by logger name
```


```{r}
allbatresults <- bind_rows(batfiles, .id = "Bat_ID") # creates a long dataset with each time point for every logger, 4 variables 

allbatresults <- allbatresults %>%
  mutate(DateTime = mdy_hms(DateTime)) # changes date to appropriate format

allbatresults <- allbatresults |> 
  mutate(NumberDate = as.numeric(DateTime))
str(allbatresults) # similar to glimpse
summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
md.pattern(allbatresults) # checking for NA's
Metadata %>%
  filter(!Bat_ID%in% file_list$Bat_ID) %>% # only selects loggers with a matching file name
  drop_na(`Filename`) # removes all na's
Metadata %>%
  filter(is.na(`Filename`)) # shows which bats do not have a logger
```

This looks good, I have 44 files and all the metadata with NAs for file name are either ibuttons that are are missing, unused, or lost data. 

Loggers have 2-8000 datapoints. All the ones used for boost have around 4,000 datapoints which makes sense.

## Missing Data

```{r}
clean_met <- Metadata %>%
  drop_na(Filename) |> # removes any ibuttons with no file
  filter(`Experiment D Died` == "D36")

head(clean_met) # look at the top of the dataset
md.pattern(clean_met, rotate.names = TRUE)
write_tsv(clean_met, file = "EPFU2025TempLoggerMetadata.tsv")
```

## Making a plot function

```{r}
plot_fun <- function(arg1) {
  allbatresults %>%
    dplyr::filter(Bat_ID == arg1) %>%
    ggplot(aes(x=DateTime, y=Temperature)) +
    geom_line() +
    ylab("Temperature (C)") + xlab("Date") + ylim(4,40) +
    theme_cowplot() 
}
```

```{r}
plotly_fun <- function(arg1) {
  plot_ly(data = allbatresults[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

## Looking at the data

```{r Bat 03}
ggplotly(plot_fun("APS36067"))
```

```{r Plotly Bat 16}
plotly_fun("APS60113")
```

Here we go!

This works fine, but is not interactive

```{r}
lapply(batnames, plotly_fun)
```

The data loads in well. There are a lot of values that are incorrect (logger never exposed to these temps). Will drop them later.
Will get in to the details about which ones need to be trimmed in a bit. 

## Calibration
Before I go any further, I need to calibrate the data. I am using a 2 point calibration from the low and high calibration temps. The formula I am using is CorrectedValue = (((RawValue – RawLow) * ReferenceRange) / RawRange) + ReferenceLow. Before I do that, I need to remove the incorrect values that are over 45 and under 0. Since I know these are incorrect readings and I would want them removed from ALL loggers, I am just going to apply it to the whole dataset.

Unfortunately, my calibration data for the second batch of loggers is really weird and does not make sense. Loggers were reading at 45C while in the hot calibration and 2C while in the cold calibration. The low temperature makes sense because the water bath was in the fridge but the reference thermometer was reading at 11C. The hot temperature does not make sense as the bath should not have been that hot and the reference thermometer read at 35C.

Thankfully, I have some loggers that went in for both calibrations, so I will calibrate all the loggers from the first calibration and then take their temperatures during the second calibration to calibrate the second batch of loggers.

```{r}
# here i am just checking the boost data all in one plot to see what it looks like
checking_boost_cal <- allbatresults |> 
  left_join(Metadata, join_by(Bat_ID)) |> # joining with metadata
  filter(`Date Attached` >= '2025-09-29') # filter by loggers attached for second immunization

p <- checking_boost_cal%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(p)

```

```{r}
clean_results <- allbatresults |> 
  filter(Temperature > 0 & # removes temperature data under 0 degrees
        Temperature < 50) # removes temperature over 50 degrees

summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
summary(as.factor(clean_results$Bat_ID)) # comparing counts between the two
```
This looks good. Now I am going to add columns to the dataframe with the maximum and minimum values during calibration. There were two different calibrations, so first I have to select loggers only used in the first calibration. 

### Calibration 1

```{r}
Clean_results_met <- clean_results %>% 
  left_join(clean_met, by = "Bat_ID") # join metadata with results so I can extract loggers in second calibration

Cal_1_results <- Clean_results_met |>  # extracting second calibration results
  filter(Calibration == 1) |> 
  select(Bat_ID, DateTime, Unit, Temperature, NumberDate)

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[NumberDate >= 1756687201 & 
                  DateTime <= 1756692601 ])) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[NumberDate >= 1756738201 & 
                                    NumberDate <= 1756801801])) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_1_results <- Cal_1_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range from average high temp and average low temp
  ungroup()  
```

Okay, now because Ken would want me to, I am going to load in the calibration data and get the range and average low. 

```{r}
Reference_cal_1 <- read_excel("Calibration Data/31-Aug_Calibration.xlsx", 
    range = "A1:D8869", na = "NA") #loading in the calibration 2 reference data

summary(Reference_cal_1) # looking at it to make sure it loaded in correctly

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(lowT = mean(T1[No.s >= 1076 & 
                  No.s <= 1433],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. I needed to use the time point because the EA15 does not record date so it loaded in as 1899-12-31 for all times. The time points lines up with the same time used in averaging low and high logger temperatures.

Reference_cal_1 <- Reference_cal_1 |> 
  mutate(highT = mean(T1[No.s >= 4473 & 
                  No.s <= 8713],  na.rm = TRUE)) #taking the average temperature during the low calibration for reference data. 
```

The average low is 4.554749, the average high is 36.97654, and the range is 32.42. 
Okay, now I am going to create a column in my cal_1_results that has the formula for the corrected values. 

```{r}
Cal_1_results <- Cal_1_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 32.42)/ RangeT) + 4.55) # applying formula listed above to the calibration to get a corrected temperature value

summary(Cal_1_results)
```
I think this data looks good. I took a look at the dataset and saw that the range is different, the low and high looks like they were calculated correctly. 


### Calibration 2
Okay, so for this one, I am going to take the three loggers that went through both calibrations. Using the first calibration, the data from the second calibration should be true values. I am going to take these temperatures during the second calibration and use them to calibrate the second batch. 
```{r}
Cal_2_results <- Clean_results_met |> # extracting second calibration results
  filter(Calibration == 2) |> 
  select(Bat_ID, DateTime, Unit, Temperature, NumberDate)

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(lowT = mean(Temperature[NumberDate >= 1759106401 & 
                  NumberDate <= 1759112401])) |> 
  ungroup() #taking the average temperature during the low calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |> 
  mutate(highT = mean(Temperature[NumberDate >= 1759125601 & 
                  NumberDate <= 1759130401])) |> 
  ungroup() #taking the average temperature during the high calibration for each logger

Cal_2_results <- Cal_2_results |> 
  group_by(Bat_ID) |>                # Group by individual bat
  mutate(RangeT = highT-lowT) |> # Calculate range and add column
  ungroup()  
```

Okay, now I am going to take Cal_1_results and filter out the three loggers in both calibrations

```{r}
Reference_cal_2 <- Cal_1_results |> 
  filter(Bat_ID %in% c("APS36010_boost", "APS36057_boost", "APS60113_boost") ) |> 
  filter(NumberDate  > 1759104001, 
         NumberDate < 1759132801)

```

I need to make sure they calibrations are all reading about the same.
```{r}

ref_cal_2 <- Reference_cal_2%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = corrected, color = Bat_ID)) +
  geom_line() +
  ylim(0, 50) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(ref_cal_2)

```
Okay, I think this looks good

```{r}
summary(Reference_cal_2)

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(lowT = mean(corrected[NumberDate >= 1759106401 & 
                  NumberDate <= 1759112401]))  #taking the average temperature during the low calibration for each logger

Reference_cal_2 <- Reference_cal_2 |> 
  mutate(highT = mean(corrected[NumberDate >= 1759125601 & 
                  NumberDate <= 1759130401])) 

summary(Reference_cal_2)
```

The average low is 4.733, the average high is 49.46, and the range is 44.73. 
Okay, now I am going to create a column in my cal_1_results that has the formula for the corrected values. 

```{r}
Cal_2_results <- Cal_2_results |> 
  mutate(corrected = ((( Temperature - lowT ) * 44.73)/ RangeT) + 4.733)

summary(Cal_2_results)
```

Okay, now that I have my data calibrated, I am going to merge the two datasets.

```{r}
cal_data <- bind_rows(Cal_1_results, Cal_2_results) |> 
  select(Bat_ID, DateTime, NumberDate, corrected) |> 
  rename("Temperature" = "corrected")

```


## Left and Right Trimming the data

I will split the data, filter it, and then recombine them.

```{r}
# for analysis down below, I cannot use DateTime data so I need to convert my time into a numeric 
clean_met <- clean_met %>%
  mutate(Attachment_Date = as.numeric(`Date Attached`),
         Removal_Date = as.numeric(`Date Removed`)) 

temps_with_meta <- cal_data %>%
  left_join(clean_met, by = "Bat_ID") #joining my calibrated data with my metadata

# Filter to keep only records after attachment date
left_trimmed <- temps_with_meta %>%
  filter(NumberDate >= Attachment_Date)

# Filter to keep only records before removal date
right_trimmed <- left_trimmed |> 
  filter(NumberDate <= Removal_Date)

# Making a new data table with trimmed data 
allbatresults2 <- right_trimmed |> 
  as.data.table()

#checking the data
summary(as.factor(allbatresults2$Bat_ID))

glimpse(allbatresults2)

```

```{r}
plotly_fun2 <- function(arg1) {
  plot_ly(data = allbatresults2[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(batnames, plotly_fun2)
```
## Left trim again

I left trimmed earlier to remove all data from before the day of attachment. I really want to include as much as possible so I am trimming again by the attachment time. This is going to take forever... :/

```{r}

# first trim
Trimming1 <- allbatresults2 |> 
  filter(Attachment_Date == 1756771200) |> 
  filter(NumberDate >= 1756826401)
summary(as.factor(Trimming1$Bat_ID))

#second trim

Trimming2 <- allbatresults2 |> 
  filter(Attachment_Date == 1757116800) |> 
  filter(NumberDate >= 1757170201)
summary(as.factor(Trimming2$Bat_ID))

#third trim

Trimming3 <- allbatresults2 |> 
  filter(Attachment_Date == 1757376000) |> 
  filter(NumberDate >= 1757439601)
summary(as.factor(Trimming3$Bat_ID))

#fourth trim

Trimming4 <- allbatresults2 |> 
  filter(Attachment_Date == 1759104000) |> 
  filter(NumberDate >= 1759152601)
summary(as.factor(Trimming4$Bat_ID))

#fifth trim

Trimming5 <- allbatresults2 |> 
  filter(Attachment_Date == 1759363200) |> 
  filter(NumberDate >= 1759419601)
summary(as.factor(Trimming5$Bat_ID))

LTrimmed <- bind_rows(Trimming1, Trimming2, Trimming3, Trimming4, Trimming5)

```

## Right trim again 

```{r}
plotly_fun3 <- function(arg1) {
  plot_ly(data = LTrimmed[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }
```

```{r}
lapply(batnames, plotly_fun3)
```

Okay, this is looking good. I do have some more data to trim. So I am going to do that now.
```{r}
# Right trim 1
RTrimming1 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36008")) |> 
  filter(!NumberDate > 1760539201)
summary(as.factor(RTrimming1$Bat_ID))

# Right trim 2
RTrimming2 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36021")) |> 
  filter(!NumberDate > 1759998001)
summary(as.factor(RTrimming2$Bat_ID))

# Right trim 3

RTrimming3 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36026")) |> 
  filter(!NumberDate > 1760208601)
summary(as.factor(RTrimming3$Bat_ID))

# Right Trim 4 

RTrimming4 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36057_boost")) |> 
  filter(!NumberDate > 1760530201)
summary(as.factor(RTrimming4$Bat_ID))

# RTrim 5

RTrimming5 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36108")) |> 
  filter(!NumberDate > 1760455801)
summary(as.factor(RTrimming5$Bat_ID))

# RTrim 6

RTrimming6 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60081")) |> 
  filter(!NumberDate > 1758842401)
summary(as.factor(RTrimming6$Bat_ID))

# RTrim 7

RTrimming7 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60115")) |> 
  filter(!NumberDate > 1757681401)
summary(as.factor(RTrimming7$Bat_ID))

# RTrim 8

RTrimming8 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36007")) |> 
  filter(!NumberDate > 1759327201)
summary(as.factor(RTrimming8$Bat_ID))

# RTrim 9

RTrimming9 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS36106")) |> 
  filter(!NumberDate > 1759015201)
summary(as.factor(RTrimming9$Bat_ID))

# RTrim 10

RTrimming10 <- LTrimmed |> 
  filter(Bat_ID %in% c("APS60113")) |> 
  filter(!NumberDate > 1759327201)
summary(as.factor(RTrimming10$Bat_ID))

#Keeping all results from LTrimmed besides the ones that were right trimmed again

RTrimmed_Temp <- LTrimmed |> 
  filter(!Bat_ID %in% c( "APS36008", "APS36021", "APS36026", "APS36057_boost", "APS36108", "APS60081", "APS60115", "APS36007", "APS36106", "APS60113"))

Trimmed <- bind_rows(RTrimmed_Temp, RTrimming1, RTrimming2, RTrimming3, RTrimming4, RTrimming5, RTrimming6, RTrimming7, RTrimming8, RTrimming9, RTrimming10)

summary(as.factor(Trimmed$Bat_ID))

```

Now I want to look at the data and make sure it looks right and see what else needs to be trimmed

```{r}
cleaning_names <- unique(Trimmed$Bat_ID)


plotly_fun4 <- function(arg1) {
  plot_ly(data = Trimmed[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(cleaning_names, plotly_fun4)
```
This looks pretty good to me. The last thing I am going to do is drop data above 40 and below 12.

```{r}
final_results <- Trimmed |> 
  filter(Temperature > 12 & # removes temperature data under 0 degrees
        Temperature < 40) 
```

```{r}
results_clean <- final_results%>%
  dplyr::filter(Bat_ID %in% cleaning_names) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  ylim(10, 40) +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(results_clean)
```


I need to remove some names from batnames. There are a handful of loggers from bats that are excluded from the study or have too few data points and were dropped during trimming

```{r}

KeptBatNames <- unique(clean_results$Bat_ID)

plotly_fun4 <- function(arg1) {
  plot_ly(data = clean_results[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
  }

```


```{r}
lapply(KeptBatNames, plotly_fun4)
```

```{r}

checking <- clean_results%>%
  dplyr::filter(Bat_ID %in% batnames) %>%
  ggplot(aes(x = DateTime, y = Temperature, color = Bat_ID)) +
  geom_line() +
  theme_minimal() +
  labs(title = "All Bats Comparison", x = "Date", y = "Temperature (C)")

# Convert to interactive
ggplotly(checking)
``` 


```{r}
write_tsv(clean_results, file = "TrimmedTemperatureLoggerData.long.tsv")
```


## Ambient Temp
We had two cohorts of bats for non thermally challenged bats. They followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. Because of this, ambient temperature is different for group 1 D21 bats and group 2 D21 bats.

```{r}
 Union_Ambient <- read_excel("MOCO Data/Roost Temps/MOCO_Union_Ambient.xlsx") |> 
  rename(DateTime = "Date/Time")  |> # read in ambient temp
  mutate (DateTime = as.POSIXct(DateTime)) |> # correct format
  mutate(NumberDate = as.numeric(DateTime)) #add column changing date/time to numeric
  
Union_Ambient <- Union_Ambient |> 
filter(NumberDate >= "1747407000")

plot_ly(Union_Ambient,x = ~DateTime, y = ~Value,
          type = 'scatter', mode = 'lines')

Rivendell_Top_D0 <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Top_D0.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge 
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Bottom_D0 <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Bottom_D0.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge
  mutate(DateTime = as.POSIXct(DateTime)) |> 
   mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Primary <- Rivendell_Top_D0 |> 
  left_join(Rivendell_Bottom_D0, join_by(NumberDate)) |>  # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

Rivendell_Primary <- Rivendell_Primary |> 
  select(DateTime.x, NumberDate, Value.x, Value.y, Unit.x)  |> # keeping certain columns 
  rename(Unit = "Unit.x", # renaming columns
         Top = "Value.x",
         Bottom = "Value.y",
         DateTime = "DateTime.x") |> 
  drop_na(Bottom) # dropping NA's because loggers were not started and stopped at the exact same time
glimpse(Rivendell_Primary)

# because I have temperature from the top and bottom of the incubator and there might be slight differences between them and I don't know which temp the bats are exposed to. I am going to get the average

Rivendell_Primary <- Rivendell_Primary |> 
  mutate(avg_pri = rowMeans(pick("Top", "Bottom"))) 
  
```

 Now I have to do the same thing but for the pre- thermal challenge
```{r}
Rivendell_Top_Pre <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Top_D-5.xlsx") |> # read in temp from top of incubator from D0 thermal challenge
  rename(DateTime = "Date/Time") |> 
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Bottom_Pre <- read_excel("MOCO Data/Roost Temps/Rivendell-Cold_Bottom_D-5.xlsx") |> rename(DateTime = "Date/Time") |> # read in temp from top of incubator from D0 thermal challenge
  mutate(DateTime = as.POSIXct(DateTime)) |> 
  mutate(Shift_date = DateTime - seconds(1)) |> # data was recorded on the 00:00:01 need to shift it
  mutate(NumberDate = as.numeric(Shift_date))

Rivendell_Pre <- Rivendell_Top_Pre |> 
  left_join(Rivendell_Bottom_Pre, join_by(NumberDate)) |> # joining the top and bottom temperature data from the primary thermal challenge
  as.data.table()

Rivendell_Pre <- Rivendell_Pre |> 
  select(DateTime.x, NumberDate, Value.x, Value.y, Unit.x)  |> # keeping certain columns 
  rename(Unit = "Unit.x", # renaming columns
         Top = "Value.x",
         Bottom = "Value.y",
         DateTime = "DateTime.x") |> 
  drop_na(Bottom) # dropping NA's because loggers were not started and stopped at the exact same time
glimpse(Rivendell_Pre)

# because I have temperature from the top and bottom of the incubator and there might be slight differences between them and I don't know which temp the bats are exposed to. I am going to get the average

Rivendell_Pre <- Rivendell_Pre |> 
  mutate(avg_pre = rowMeans(pick("Top", "Bottom"))) 

```

Now I want to only keep the data from the actual thermal challenges
Start time for the pre-thermal challenge was 12:05 23-May-2025 to 01:17 24-May-2025
Start Time for primary thermal challenge was 16:36 26-May-2025 to 00:44 27-May-2025

```{r}
Rivendell_Primary_Trim <- Rivendell_Primary |> 
  filter(NumberDate > "1748277361", 
         NumberDate < "1748306641") |> 
  select(DateTime, NumberDate, avg_pri) |> 
  rename(Value = "avg_pri")

Rivendell_Pre_Trim <- Rivendell_Pre |> 
  filter(NumberDate > "1748001901", 
         NumberDate < "1748049421") |> 
  select(DateTime, NumberDate, avg_pre) |> 
  rename(Value = "avg_pre")
```

Okay, now that's done.. I need to take these times and add them to the corresponding rows. I think what I am going to do is separate the boost from the primary results, because my primary result bats were all in an incubator at that time. 
```{r}
Primary_results_working <- clean_results |> 
  filter(Bat_ID %in% c("64.1", "68.1", "74.1")) |> # results from primary immunization
  filter(NumberDate <= "1748736000")

Secondary_results_working <- clean_results |> 
  filter(!Bat_ID %in% c("64.1", "68.1", "74.1")) # results from secondary immunization

# trim down the ambient data to the dates and time loggers were attached to the bats
Ambient_primary_working <- Union_Ambient |> 
  filter(NumberDate <= "1748736000", 
         NumberDate >= "1747736400")

# now I need to filter out times when the bats were in the incubators instead of outside with ambient temperature
AP_Trim1 <- Ambient_primary_working |> 
  filter(NumberDate <= "1748001901")

AP_Trim2 <- Ambient_primary_working |> 
   filter(NumberDate >= "1748049421",
          NumberDate <= "1748277361")

Ap_Trim3 <- Ambient_primary_working |> 
  filter(NumberDate >= "1748306641")
          
# Now I am going to combine the incubator temps and ambient temps into one dataframe in one column. This willl make calculating my delta T much easier.. I think.... 

Primary_exposure_temp <- bind_rows(AP_Trim1, Rivendell_Pre_Trim, AP_Trim2, Rivendell_Primary_Trim, Ap_Trim3)

Primary_results_working <- Primary_results_working |>  
  left_join(Primary_exposure_temp, join_by("NumberDate"))

Primary_results_working <- Primary_results_working |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, Value, Unit) |> 
  rename(AmbientT = "Value",
         DateTime = "DateTime.x")

Primary_results <- Primary_results_working |> 
  mutate(deltaT = Temperature - AmbientT)

Secondary_results <- Secondary_results_working |> 
  left_join(Union_Ambient, join_by(NumberDate))

Secondary_results <- Secondary_results |> 
  select(Bat_ID, DateTime.x, NumberDate, Temperature, Value, Unit) |> 
  rename(AmbientT = "Value",
         DateTime = "DateTime.x")

Secondary_results <- Secondary_results |> 
  mutate(deltaT = Temperature - AmbientT)

```



### Shifting Data

We had two cohorts of bats for non thermally challenged bats. They followed the exact same timeline but were treated on different days of the week. This was done to limit handling time of bats at one time. Because of this, ambient temperature is different for group 1 D21 bats and group 2 D21 bats. 

```{r}
SecondaryBatNames <- unique(Secondary_results$Bat_ID)

Secondary_Met <- no_hot_met %>%
  filter(Bat_ID %in% SecondaryBatNames) |> 
  select(! "Logger location") |> 
  select(! "Notes")

summary(as.factor(Secondary_Met$Treatment))
# shows 6 TC bats but I thought we had 7.. look into this.

SecondaryResultsMet <- Secondary_results |> 
left_join(Secondary_Met, by = join_by(Bat_ID))

SecondaryResultsMet %>%
  group_by(Bat_ID) %>%
  summarize(Min_DateTime = min(DateTime), Max_DateTime = max(DateTime))
```

```{r}
SecondaryResultsMet$time_aligned <- SecondaryResultsMet$DateTime

md.pattern(SecondaryResultsMet, rotate.names = TRUE)

SecondaryResultsMet <- SecondaryResultsMet |> 
  select(! "Serial Number") |> 
  select(! "Logger Name") |> 
  select(! "Filename")

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "2"] - lubridate::days(2)

SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] <- SecondaryResultsMet$time_aligned[SecondaryResultsMet$Group == "3"] - lubridate::days(11)
```

Okay, one of the last things I want to do is remove the decimal from the Bat ID. This was just to differentiate between loggers for one bat with all the different metadata. 
```{r}
Working_Secondary <- SecondaryResultsMet |> 
  filter(Treatment %in% c("SHAM", "eVLP+LPS", "TC"))

Bat_64 <- Working_Secondary |> 
  filter(Bat_ID == "64.2")
Bat_64$Bat_ID <- c("64")

Bat_68 <- Working_Secondary |> 
  filter(Bat_ID == "68.2")
Bat_68$Bat_ID <- c("68")

Bat_73 <- Working_Secondary |> 
  filter(Bat_ID == "73.4")
Bat_73$Bat_ID <- c("73")

Bat_74 <- Working_Secondary |> 
  filter(Bat_ID == "74.2")
Bat_74$Bat_ID <- c("74")

Bat_76 <- Working_Secondary |> 
  filter(Bat_ID == "76.2")
Bat_76$Bat_ID <- c("76")

Bat_78 <- Working_Secondary |> 
  filter(Bat_ID == "78.9")
Bat_78$Bat_ID <- c("78")

Bat_83 <- Working_Secondary |> 
  filter(Bat_ID == "83.2")
Bat_83$Bat_ID <- c("83")

Working_Secondary2 <- Working_Secondary |> 
  filter(!Bat_ID %in% c("64.2", "68.2", "73.4", "74.2", "76.2", "78.9", "83.2"))

Working_Secondary3 <- Working_Secondary2 |> 
  bind_rows(Bat_64, Bat_68, Bat_73, Bat_74, Bat_76, Bat_78, Bat_83)

Final_working_sec <- Working_Secondary3 |> 
  select(Bat_ID, time_aligned, Temperature, AmbientT, deltaT)

Final_Secondary <- Final_working_sec |> 
  rename(DateTime = "time_aligned") |> 
  mutate(NumberDate = as.numeric(DateTime)) 

SecondaryNames <- unique(Final_Secondary$Bat_ID)

Working_met_sec <- Working_Secondary3 |> 
  select(Bat_ID, `Bat Pittag`, Group, Treatment, `Innoculation Date`, `Boost Date`)

Secondary_Met2 <- Working_met_sec |> 
  unique()

SecondaryNames <- unique(Final_Secondary$Bat_ID)

temp_primary_met <- Primary_results |> 
  left_join(clean_met, join_by(Bat_ID))

Bat_64.1 <- temp_primary_met |> 
  filter(Bat_ID == "64.1")
Bat_64.1$Bat_ID <- c("64")

Bat_68.1 <- temp_primary_met |> 
  filter(Bat_ID == "68.1")
Bat_68.1$Bat_ID <- c("68")

Bat_74.1 <- temp_primary_met |> 
  filter(Bat_ID == "74.1")
Bat_74.1$Bat_ID <- c("74")

Primary_working <- bind_rows(Bat_64.1, Bat_68.1, Bat_74.1)

Final_Primary <- Primary_working |> 
  select(Bat_ID, DateTime, NumberDate, Temperature, AmbientT, deltaT)

PrimaryNames <- unique(Final_Primary$Bat_ID)

Primary_met_working <- Primary_working |> 
  select(Bat_ID, `Bat Pittag`, Group, Treatment, `Innoculation Date`, `Boost Date`)

Primary_Met <- Primary_met_working|> 
  unique()

Primary_Met <- clean_met |> 
  filter(Bat_ID %in% c("64.1", "68.1", "74.1")) 

Secondary_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1748967000" , 
          NumberDate < "1749772800")

Group_1_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1747406400", 
         NumberDate < "1750290600")

Group_2_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1747440000", 
         NumberDate < "1750447800")

Group_3_Ambient <- Union_Ambient |> 
  filter(NumberDate > "1748217600", 
         NumberDate < "1750290600")
```

```{r}

plotly_fun6 <- function(arg1) {
  plot_ly(data = Final_Secondary[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}

lapply(SecondaryNames, plotly_fun6)

```
 
I shifted the data so now data all shows d21 on June 5th, the same day as D21 for Group 1. 
I think I have all the data wrangling down so I am going to export some tsv with my final primary data, secondary data, metadata for both immunizations, and ambient temperature.



```{r}
write_tsv(Secondary_Met2, file = "MOCO2025Secondary_Met.tsv")
write_tsv(Final_Primary, file = "MOCO2025PrimaryTemperature.tsv")
write_tsv(Final_Secondary, file = "MOCO2025SecondaryTemperature.tsv")
write_tsv(Primary_Met, file = "MOCO2025Primary_Met.tsv")
write_tsv(Ambient_primary_working, file = "MOCO2025Primary_Ambient.tsv")
write_tsv(Secondary_Ambient, file = "MOCO2025Secondary_Ambient.tsv")
write_tsv(Group_1_Ambient, file = "MOCO2025_G1_Ambient.tsv")
write_tsv(Group_2_Ambient, file = "MOCO2025_G2_Ambient.tsv")
write_tsv(Group_3_Ambient, file = "MOCO2025_G3_Ambient.tsv")
write_tsv(Union_Ambient, file = "MOCO2025_Ambient.tsv")

```



