---
title: "EPFU 2025 Temperature Loggers Analysis"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
#if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("zoo")) install.packages("zoo"); library(zoo)
if (!require("emmeans")) install.packages("emmeans"); library(emmeans)
if (!require("multcomp")) install.packages("multcomp"); library(multcomp)
if (!require("interactions")) install.packages("interactions"); library(interactions)
if (!require("colorspace")) install.packages("colorspace"); library(colorspace)
if (!require("broom")) install.packages("broom"); library(broom)
if (!require("performance")) install.packages("performance"); library(performance)
if (!require("MASS")) install.packages("MASS"); library(MASS)
if (!require("lmerTest")) install.packages("lmerTest"); library(lmerTest)
if (!require("patchwork")) install.packages("patchwork"); library(patchwork)



conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

This code is going to be AUC models. In EPFU2025TempWrangling, I calibrated and trimmed the data as well as dropping any bad values. In EPFU2025TempGeneralModels, I ran models to see what types variables affected AUC so I know what I need to include in my models. Both weight and group had a significant affect on AUC so I want to be sure to correct for these differences.

## Metadata

```{r}
Analysis_results <- read.delim("EPFU2025Analysis_Results.tsv", # read in metadata for ibuttons
     na = "NA") 

glimpse(Analysis_results) # check formatting

Analysis_results$Bat_ID = as.character(Analysis_results$Bat_ID) # change Bat_ID to a character

glimpse(Analysis_results) # check formatting again
```

I am going to load in my data and metadata for both immunizations

Make values for bat names and assign colors for groups
```{r}
ResultsNames <- unique(Analysis_results$Bat_ID)

treatmentcolors <- c("CP" = "#9EB0FFFF",
            "CP+S" = "#2D7597FF",
            "CS" = "#FFACACFF", 
            "NoCo" =  "#8A3F2A",
            "SHAM" = "#11181DFF",
            "Tnorm" = "black")

primary_colors <- c("Cold" = "#9EB0FFFF",
                    "NoCo" = "#8A3F2A",
                    "SHAM" = "#11181DFF" )
```

Okay, first thing I am going to do is group all CS and NoCo and CP and CP+S into two groups before secondary. I am then only going to pull out days before Secondary.
```{r}
Analysis_results <- Analysis_results |> 
  mutate(Primary_Treatment = case_when(Treatment %in% c("CP+S" , "CP") ~  "Primary_Cold",
         Treatment %in% c("CS" , "NoCo") ~ "Primary_NoCo",
         TRUE ~ "SHAM"))
```
I am also going to make two additional columns. One that is regular treatment+datetype and another that is 'primary treatment'+ datetype. 

```{r}
Analysis_results$Group_Combo <- factor(paste(Analysis_results$Treatment, Analysis_results$DateType, sep = "_"))

Analysis_results$Primary_Combo <- factor(paste(Analysis_results$Primary_Treatment, Analysis_results$DateType, sep = "_"))

Analysis_results <- Analysis_results |> 
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "NoCo", "CP", "CS", "CP+S")))

Analysis_results <- Analysis_results |> 
  mutate(DateType = factor(DateType, levels = c("None", "Basal TC-1", "Basal TC", "Basal TC + 1", "Pre", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3", "D27", "D28")))

```

```{r}
Primary_Results <- Analysis_results |> 
  filter(!ExperimentDay %in% c("D20", "D21", "D22", "D23", "D24", "D25", "D26", "D27", "D28", "D29", "D30", "D31", "D32", "D33", "D34", "D35", "D36"),
         !DateType %in% c("Basal TC-1", "Basal TC", "Basal TC + 1", "Pre"))

Primary_Results <- droplevels(Primary_Results)
levels(Primary_Results$Primary_Combo)

```

## Primary Models
Okay, I am going to run some primary models now.

```{r}
primary_glm <- glm(Total_AUC ~ Bat_ID + Group + Weight + Primary_Combo, data = Primary_Results)

primary_emmeans <- emmeans(primary_glm, ~ Primary_Combo)

primary_subset <- list(
  "SvN_P" = c(0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0),
  "SvN_1" = c(0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0),
  "SvN_2" = c(0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0),
  "SvN_3" = c(0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1),
  "NvC_P" = c(0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
  "NvC_1" = c(0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
  "NvC_2" = c(0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0),
  "NvC_3" = c(0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0),
  "SvN_None" = c(0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0),
  "NvC_None" = c(1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0),
  "SvN_-1" = c(0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0),
  "NvC_-1" = c(0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0))

# 3. Run the contrast. 
# R will automatically apply these exact vectors to Males, and then to Females.
contrast(primary_emmeans, method = primary_subset, adjust = "bonferroni")
```
Okay, I am getting NAs and I think it is because of Bat_ID
I am going to rerun the model without Bat_ID
```{r}
primary_glm2 <- glm(Total_AUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

primary_emmeans2 <- emmeans(primary_glm2, ~ Primary_Combo)

contrast(primary_emmeans2, method = primary_subset, adjust = "bonferroni")
```
Okay this ran and the significant values make sense; however, after asking some questions with gemini I don't think I can ignore Bat_ID because the glm assumes that every row is independent of other rows. Since bat_id is found multiple times across rows they are not independent of each other. It wants me to run a mixed effects model. I know Ken and I talked about this and he doesnt want me to run one, but I am going to anyways.

```{r}
primary_mixed <- lmerTest::lmer(Total_AUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(primary_mixed)

primarymixed_emmeans <- emmeans(primary_mixed, ~ Primary_Combo)
primary_emmeans_df <- as.data.frame(primarymixed_emmeans)

# 2. Run your custom contrasts
contrast(primarymixed_emmeans, method = primary_subset, adjust = "bonferroni")
```
Okay, this shows that there is no statistically significant difference between sham and noco for any primary days including non-handled days. Bats that were cold at primary were significantly different than noco bats on primary which makes sense, but also on Primary +2 which is interesting.

#### Figure for primary model

```{r}
# Plot Estimated Marginal Means with Capped Error Bars (95% CI)
ggplot(primary_emmeans_df, aes(x = Primary_Combo, y = emmean)) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.15,      # Creates the "caps" on error bars
    linewidth = 0.8
  ) +
  labs(
    x = "Primary Combination Treatment",
    y = "Estimated Marginal Mean (Total AUC)",
    title = "Model-Adjusted Total AUC by Primary Combo"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Okay I don't like this...maybe it is just the labels but this feel really hard to follow

```{r}
em_df_split <- primary_emmeans_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

em_df_split <- em_df_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))

em_df_split <- em_df_split |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))

# Interaction plot with connecting lines and capped error bars
ggplot(em_df_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (NTD)",
    color = "Primary Treatment",
    title = "Interaction Effect on NTD"
  ) +
  theme_bw()
```

Redoing my figure

```{r}
p_top <- ggplot(
  subset(em_df_split, Treatment == "NoCo"), 
  aes(x = DateType, y = emmean, color = Treatment, group = 1)
) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  labs(
    x = "DateType",
    y = NULL,) +
  theme_bw() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none",
    panel.grid.major = element_blank(), # Removes major grid lines
    panel.grid.minor = element_blank()
  )

print(p_top)

# 2. Bottom 6 Panels: Treatments separated by DateType
p_bottom <- ggplot(em_df_split, 
  aes(x = Treatment, y = emmean, color = Treatment, group = Treatment)
) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + # Splits into 6 panels (2 rows x 3 columns)
  labs(
    x = "Primary Treatment",
    y = NULL,
    color = "Primary Treatment" ) +
  theme_bw() +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right", # Set to "right" if you'd prefer a legend on the side
    panel.grid.major = element_blank(), # Removes major grid lines
    panel.grid.minor = element_blank()
  )
print(p_bottom)
# 3. Combine Top & Bottom (Height ratio 1:2)
combined_plot <- (p_top / p_bottom) + 
  plot_layout(guides = "collect") & 
  labs(y = "EMMean (NTD)")

p_divider <- ggplot() + 
  theme_void() + 
  geom_hline(yintercept = 0.5, color = "gray70", linewidth = 0.8)

# 2. Stack them: Top plot -> Divider line -> Bottom plot (giving the divider a very small height)
combined_plot <- (p_top / p_divider / p_bottom) + 
  plot_layout(heights = c(1, 0.05, 2.2), guides = "collect", axis_titles = "collect") & 
  labs(y = "EMMean (NTD)")

print(combined_plot)

ggsave(
  filename = "primary_NTDplot.png", 
  plot = combined_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

Trying to include the raw data. Ken also wants to look at 75% CI instead of 95%
```{r}
p_emmean_75 <- as.data.frame(primarymixed_emmeans, level = 0.75)

split_75 <- p_emmean_75 %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

split_75 <- split_75 |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))

split_75 <- split_75 |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))

p_emmean_50 <- as.data.frame(primarymixed_emmeans, level = 0.50)

split_50 <- p_emmean_50 %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = case_when(
      str_starts(Primary_Combo, "Primary_Cold_") ~ str_remove(Primary_Combo, "Primary_Cold_"),
      str_starts(Primary_Combo, "Primary_NoCo_") ~ str_remove(Primary_Combo, "Primary_NoCo_"),
      str_starts(Primary_Combo, "SHAM_")         ~ str_remove(Primary_Combo, "SHAM_"),
      TRUE                                       ~ NA_character_))

split_50 <- split_50 |> 
  mutate(DateType = factor(DateType, levels = c("None", "Primary - 1", "Primary", "Primary + 1", "Primary + 2", "Primary + 3")))

split_50 <- split_50 |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "Cold", "SHAM")))
```

Renaming some things in raw data so it matches my dataframe
```{r}
Primary_raw <- Primary_Results |> 
  rename("Assigned_treatment" = "Treatment") |> 
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Treatment, "Primary_Cold") ~ "Cold",
      str_starts(Primary_Treatment, "Primary_NoCo") ~ "NoCo",
      str_starts(Primary_Treatment, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ))
```

This is my 95%
```{r}
p_top95 <- ggplot() +
  geom_jitter(
    data = subset(Primary_raw, Treatment == "NoCo"),
    aes(x = DateType, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(em_df_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(em_df_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(em_df_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_top95)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
p_bottom95 <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = em_df_split,
    aes(x = Treatment, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = em_df_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3
  ) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = NULL, color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_bottom95)

combined_plot95 <- (p_top95 / p_divider / p_bottom95) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "NTD (Degree Hours)") 

print(combined_plot95)

ggsave(
  filename = "primary_NTDplot_95.png", 
  plot = combined_plot95, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

This is my 75% CI
```{r}
p_top75 <- ggplot() +
  geom_jitter(
    data = subset(Primary_raw, Treatment == "NoCo"),
    aes(x = DateType, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(split_75, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(split_75, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(split_75, Treatment == "NoCo"),
    aes(x = DateType, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_top75)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
p_bottom75 <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = split_75,
    aes(x = Treatment, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = split_75,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3
  ) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = NULL, color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_bottom75)

combined_plot75 <- (p_top75 / p_divider / p_bottom75) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "NTD (Degree Hours)") 

print(combined_plot75)

ggsave(
  filename = "primary_NTDplot_75.png", 
  plot = combined_plot75, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```
50% CI
```{r}
p_top50 <- ggplot() +
  geom_jitter(
    data = subset(Primary_raw, Treatment == "NoCo"),
    aes(x = DateType, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(split_50, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(split_50, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(split_50, Treatment == "NoCo"),
    aes(x = DateType, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  scale_color_manual(values = primary_colors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_top50)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
p_bottom50 <- ggplot() +
  geom_jitter(
    data = Primary_raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = split_50,
    aes(x = Treatment, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = split_50,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3
  ) +
  scale_color_manual(values = primary_colors) +
  scale_fill_manual(values = primary_colors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Primary Treatment", y = NULL, color = "Primary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p_bottom50)

combined_plot50 <- (p_top50 / p_divider / p_bottom50) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "NTD (Degree Hours)") 

print(combined_plot50)

ggsave(
  filename = "primary_NTDplot_50.png", 
  plot = combined_plot50, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```
Okay this is really really small so we are going to stick with 95% CI. 75% is hard to justify, but 50% is what is normally in box plots so thats why we tried it. 


### Hypo NTD
I am going to run the same models with Hypo and Hyper AUC. I am doing this because I want to see if immunized bats had a febrile response and if they did, if the thermal challenge affected that response. The way AUC is calculated currently 'cancels' the extreme values out.

```{r}
primary_hypo <- glm(HypoAUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

p_hypo_emmeans <- emmeans(primary_hypo, ~ Primary_Combo)

contrast(p_hypo_emmeans, method = primary_subset, adjust = "bonferroni")
```

```{r}
hypo_mixed <- lmerTest::lmer(HypoAUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(hypo_mixed)

hypomixed_emmeans <- emmeans(hypo_mixed, ~ Primary_Combo)

levels(hypomixed_emmeans)

# 2. Run your custom contrasts
contrast(hypomixed_emmeans, method = primary_subset, adjust = "bonferroni")
```

#### Hypo Figure
```{r}
hypo_df_p <- as.data.frame(emmeans(hypo_mixed, ~ Primary_Combo))

hypo_split <- hypo_df_p %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Primary_Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "Primary_NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = str_remove(Primary_Combo, paste0("^", Treatment, "_?"))
  )

# Interaction plot with connecting lines and capped error bars
ggplot(hypo_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hypo NTD)",
    color = "Primary Treatment",
    title = "Interaction Effect on Hypo NTD"
  ) +
  theme_bw()
```

### Hyper NTD

```{r}
primary_hyper <- glm(HyperAUC ~ Group + Weight + Primary_Combo, data = Primary_Results)

p_hyper_emmeans <- emmeans(primary_hyper, ~ Primary_Combo)

contrast(p_hyper_emmeans, method = primary_subset, adjust = "bonferroni")
```
Okay, this is very interesting. I am going to run this as a mixed effects model just so I am including Bat_ID still

```{r}
hyper_mixed <- lmerTest::lmer(HyperAUC ~ Primary_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Primary_Results)

# Check the summary
summary(hyper_mixed)

hypermixed_emmeans <- emmeans(hyper_mixed, ~ Primary_Combo)

levels(hypermixed_emmeans)



# 2. Run your custom contrasts
contrast(hypermixed_emmeans, method = primary_subset, adjust = "bonferroni")
```
#### Hyper Figure 

```{r}
hyper_df_p <- as.data.frame(emmeans(hyper_mixed, ~ Primary_Combo))

hyper_split <- hyper_df_p %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Primary_Combo, "Primary_Cold") ~ "Primary_Cold",
      str_starts(Primary_Combo, "Primary_NoCo") ~ "Primary_NoCo",
      str_starts(Primary_Combo, "SHAM")         ~ "SHAM",
      TRUE                                      ~ NA_character_
    ),
    
    # Remove Factor A and the connecting underscore to leave Factor B
    DateType = str_remove(Primary_Combo, paste0("^", Treatment, "_?"))
  )

# Interaction plot with connecting lines and capped error bars
ggplot(hyper_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hyper NTD)",
    color = "Primary Treatment",
    title = "Interaction Effect on Hyper NTD"
  ) +
  theme_bw()
```

## Secondary Models

Okay, now I am going to run the same things but for secondary models. I need to decide if I should run this model with their original treatment groups or their 'this is what they actually did groups' since I have some bats that did not go cold in the primary thermal challenge.

I am going to look at a histogram and see if I can tell which bats used torpor or not. I am going to look at both total AUC and hypoAUC. Just in case there are some bats that used torpor, but if they remained euthermic or even had a fever at some point during the day, their total AUC might not be as low.

```{r}
# 1. Visualize the distribution to find the "cut point"
Primary_only_results <- Analysis_results |> 
  filter(DateType == "Primary")

hist(Primary_only_results$Total_AUC, breaks = 20, main = "Distribution of Primary Torpor")
hist(Primary_only_results$HypoAUC, breaks = 20, main = "Hypo AUC")

```
I talked to Ken about this and the bats did not go cold during the primary thermal challenge still experienced different energetic requirements that bats that were not cold at primary. For that reason, we should not group them together for the secondary analysis even if they were not cold at the primary.

```{r}
Secondary_Results <- Analysis_results |> 
  filter(!ExperimentDay %in% c("D0", "D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14", "D15", "D16", "D17", "D18", "D19", "D-1", "D27", "D28"),
         !DateType %in% c("Basal TC-1", "Basal TC", "Basal TC + 1", "Pre"))

secondary_bats <- Secondary_Results %>%
  filter(ExperimentDay == "D21") %>%
  pull(Bat_ID) %>%
  unique()

# 2. Keep all rows for those bats across all days
Secondary_Results_filtered <- Secondary_Results %>%
  filter(Bat_ID %in% secondary_bats)

Secondary_Results <- droplevels(Secondary_Results)
levels(Secondary_Results$Group_Combo)
```

I am going to run the same models for seconary as I did for primary
```{r}
secondary_glm <- glm(Total_AUC ~ Bat_ID + Group + Weight + Group_Combo, data = Secondary_Results)

secondary_emmeans <- emmeans(secondary_glm, ~ Group_Combo)

secondary_subset <- list(
  "SvN_S" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0),
"SvN_1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0),
"SvN_2"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0),
"SvN_3"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1),
"SvN_-1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0),
"SvN_None"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0),
"NvCP_S" = c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
"NvCP_1"= c(0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
"NvCP_2"= c(0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0),
"NvCP_3"= c(0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0),
"NvCP_-1"= c(0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0),
"NvCP_None"= c(1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_S"= c(0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_1"= c(0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_2"= c(0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_3"= c(0,0,0,0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_-1"= c(0,0,0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPSvCS_None"= c(0,0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_S"= c(0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_1"= c(0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_2"= c(0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_3" = c(0,0,0,0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_-1"= c(0,0,-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"CPvCPS_None" = c(-1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
"NvCS_S"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0),
"NvCS_1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0),
"NvCS_2" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0),
"NvCS_3" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0),
"NvCS_-1"= c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0),
"NvCS_None" = c(0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,-1,0,0,0,0,0,0,0,0,0,0,0))

# 3. Run the contrast. 
# R will automatically apply these exact vectors to Males, and then to Females.
contrast(secondary_emmeans, method = secondary_subset, adjust = "bonferroni")
```
Okay, I am getting NAs and I think it is because of Bat_ID
I am going to rerun the model without Bat_ID
```{r}
secondary_glm2 <- glm(Total_AUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

secondary_emmeans2 <- emmeans(secondary_glm2, ~ Group_Combo)

contrast(secondary_emmeans2, method = secondary_subset, adjust = "bonferroni")
```

Running mixed effects model.. 
```{r}
secondary_mixed <- lmerTest::lmer(Total_AUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(secondary_mixed)

secondarymixed_emmeans <- emmeans(secondary_mixed, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(secondarymixed_emmeans, method = secondary_subset, adjust = "bonferroni")
```
So there is no statistically difference in AUC between SHAM and NoCo, CP+S and CS, or NoCo and CP on any day. CP+S had a significantly lower AUC than CP on D21 and CS had a significantly lower AUC than Noco on D21. 

#### Secondary Figure

```{r}
Sec_df <- as.data.frame(secondarymixed_emmeans, level = 0.95)

sec_df_split <- Sec_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

# Interaction plot with connecting lines and capped error bars
ggplot(sec_df_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (NTD)",
    color = "Secondary Treatment",
    title = "Interaction Effect on NTD"
  ) +
  theme_bw()
```

Okay rerunning with the same as primary.

```{r}
sec_df_split <- sec_df_split |> 
  mutate(DateType = factor(DateType, levels = c("None", "Secondary - 1", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")))

sec_df_split <- sec_df_split |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")))

Secondary_Raw <- Secondary_Results_filtered |> 
  mutate(Treatment = factor(Treatment, levels = c("NoCo", "CP", "CP+S", "CS", "SHAM")))
```

```{r}
s_top <- ggplot() +
  geom_jitter(
    data = subset(Secondary_Raw, Treatment == "NoCo"),
    aes(x = DateType, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_line(
    data = subset(sec_df_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment, group = 1),
    linewidth = 1
  ) +
  geom_point(
    data = subset(sec_df_split, Treatment == "NoCo"),
    aes(x = DateType, y = emmean, color = Treatment),
    size = 3
  ) +
  geom_errorbar(
    data = subset(sec_df_split, Treatment == "NoCo"),
    aes(x = DateType, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  scale_color_manual(values = treatmentcolors) +
  labs(x = "DateType", y = NULL) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(s_top)

# 2. Bottom Plot (Faceted treatments with raw data points in the background)
s_bottom <- ggplot() +
  geom_jitter(
    data = Secondary_Raw,
    aes(x = Treatment, y = Total_AUC, color = Treatment),
    width = 0.15, height = 0, alpha = 0.3, size = 1.5
  ) +
  geom_errorbar(
    data = sec_df_split,
    aes(x = Treatment, ymin = lower.CL, ymax = upper.CL, color = Treatment),
    width = 0.1, linewidth = 0.7
  ) +
  geom_point(
    data = sec_df_split,
    aes(x = Treatment, y = emmean, fill = Treatment),
    shape = 22, color = "black", stroke = 0.5, size = 3
  ) +
  scale_color_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  facet_wrap(~ DateType, ncol = 3) + 
  labs(x = "Secondary Treatment", y = NULL, color = "Secondary Treatment") +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(s_bottom)

secondary_plot <- (s_top / p_divider / s_bottom) + 
  plot_layout(heights = c(1, 0.05, 1.9), guides = "collect", axis_titles = "collect") & 
  labs(y = "NTD (Degree Hours)") 

print(secondary_plot)

ggsave(
  filename = "secondary_NTDplot.png", 
  plot = secondary_plot, 
  width = 8,  # Narrower width (inches)
  height = 10,  # Much taller height (inches)
  dpi = 300)
```

### Hypo Models
I am going to run the same models with Hypo and Hyper AUC. I am doing this because I want to see if immunized bats had a febrile response and if they did, if the thermal challenge affected that response. The way AUC is calculated currently 'cancels' the extreme values out.

```{r}
secondary_hypo <- glm(HypoAUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

s_hypo_emmeans <- emmeans(secondary_hypo, ~ Group_Combo)

contrast(s_hypo_emmeans, method = secondary_subset, adjust = "bonferroni")
```
**I tried running this on hyper and i was getting an error. I think the error was because variance for hyper was almost 0, so they model was not able to factor in group and bat_id as a random effect.**
```{r}
s_hypo_mixed <- lmerTest::lmer(HypoAUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hypo_mixed)

s_hypo_emmeans <- emmeans(s_hypo_mixed, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(s_hypo_emmeans, method = secondary_subset, adjust = "bonferroni")
```

#### Hypo Figure

```{r}
S_hypo_df <- as.data.frame(s_hypo_emmeans)

s_hypo_split <- S_hypo_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

# Interaction plot with connecting lines and capped error bars
ggplot(s_hypo_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hypo NTD)",
    color = "Secondary Treatment",
    title = "Interaction Effect on Hypo NTD"
  ) +
  theme_bw()
```

### Hyper Models
```{r}
secondary_hyper <- glm(HyperAUC ~ Group + Weight + Group_Combo, data = Secondary_Results_filtered)

s_hyper_emmeans <- emmeans(secondary_hyper, ~ Group_Combo)

contrast(s_hyper_emmeans, method = secondary_subset, adjust = "bonferroni")
```

```{r}
s_hyper_mixed <- lmerTest::lmer(HyperAUC ~ Group_Combo + Weight + (1 | Group / Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hyper_mixed)

s_hypermixed_emmeans <- emmeans(s_hyper_mixed, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(s_hypermixed_emmeans, method = secondary_subset, adjust = "bonferroni")
```

```{r}
s_hyper_mixed2 <- lmerTest::lmer(HyperAUC ~ Group_Combo + Weight + (1 | Bat_ID), 
                      data = Secondary_Results_filtered)

# Check the summary
summary(s_hyper_mixed2)

s_hypermixed_emmeans2 <- emmeans(s_hyper_mixed2, ~ Group_Combo)

# 2. Run your custom contrasts
contrast(s_hypermixed_emmeans2, method = secondary_subset, adjust = "bonferroni")
```

#### Hyper Figure

```{r}
S_hyper_df <- as.data.frame(s_hypermixed_emmeans2)

s_hyper_split <- S_hyper_df %>%
  mutate(
    # Extract Factor A using exact prefix matching
    Treatment = case_when(
      str_starts(Group_Combo, "CP\\+S") ~ "CP+S",
      str_starts(Group_Combo, "CP") ~ "CP",
      str_starts(Group_Combo, "CS")  ~ "CS",
      str_starts(Group_Combo, "NoCo")  ~ "NoCo",
      str_starts(Group_Combo, "SHAM")  ~ "SHAM",
      TRUE                                      ~ NA_character_
    )) |> 
  mutate(
    DateType = Group_Combo %>% 
      # Remove Factor_A literally
      str_remove(fixed(Treatment)) %>% 
      # Remove any leading underscore that was left behind
      str_remove("^_"))

# Interaction plot with connecting lines and capped error bars
ggplot(s_hyper_split, aes(x = DateType, y = emmean, color = Treatment, group = Treatment)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.1,
    linewidth = 0.7
  ) +
  labs(
    x = "DateType",
    y = "Estimated Marginal Mean (Hyper NTD)",
    color = "Secondary Treatment",
    title = "Interaction Effect on Hyper NTD"
  ) +
  theme_bw()
```

## Longitudinal Models

Okay, for these models I need to keep bats that have data from both D0 and D21
```{r}
D0 <- Analysis_results |> 
  filter(ExperimentDay == "D0") |> 
  rename(AUC_D0 = "Total_AUC",
         HyperD0 = "HyperAUC",
         HypoD0 = "HypoAUC") |> 
  select(Bat_ID, AUC_D0, Treatment, Group, HyperD0, HypoD0)
D1 <- Analysis_results |> 
  filter(ExperimentDay == "D1") |> 
  rename(AUC_D1 = "Total_AUC",
         HyperD1 = "HyperAUC",
         HypoD1 = "HypoAUC") |> 
  select(Bat_ID, AUC_D1, Treatment, Group, HyperD1, HypoD1)
D2 <- Analysis_results |> 
  filter(ExperimentDay == "D2") |> 
  rename(AUC_D2 = "Total_AUC",
         HyperD2 = "HyperAUC",
         HypoD2 = "HypoAUC") |> 
  select(Bat_ID, AUC_D2, Treatment, Group, HyperD2, HypoD2)
D3 <- Analysis_results |> 
  filter(ExperimentDay == "D3") |> 
  rename(AUC_D3 = "Total_AUC",
         HyperD3 = "HyperAUC",
         HypoD3 = "HypoAUC") |> 
  select(Bat_ID, AUC_D3, Treatment, Group, HyperD3, HypoD3)
D21 <- Analysis_results |> 
  filter(ExperimentDay == "D21") |> 
  rename(AUC_D21 = "Total_AUC",
         HyperD21 = "HyperAUC",
         HypoD21 = "HypoAUC") |> 
  select(Bat_ID, AUC_D21, Treatment, Group, HyperD21, HypoD21)
D22 <- Analysis_results |> 
  filter(ExperimentDay == "D22") |> 
  rename(AUC_D22 = "Total_AUC",
         HyperD22 = "HyperAUC",
         HypoD22 = "HypoAUC") |> 
  select(Bat_ID, AUC_D22, Treatment, Group, HyperD22, HypoD22)
D23 <- Analysis_results |> 
  filter(ExperimentDay == "D23") |> 
  rename(AUC_D23 = "Total_AUC",
         HyperD23 = "HyperAUC",
         HypoD23 = "HypoAUC") |> 
  select(Bat_ID, AUC_D23, Treatment, Group, HyperD23, HypoD23)
D24 <- Analysis_results |> 
  filter(ExperimentDay == "D24") |> 
  rename(AUC_D24 = "Total_AUC" ,
         HyperD24 = "HyperAUC",
         HypoD24 = "HypoAUC") |> 
  select(Bat_ID, AUC_D24, Treatment, Group, HyperD24,HypoD24)
Basal_TC <- Analysis_results |> 
  filter(DateType == "Basal TC") |> 
  rename(AUC_BTC = "Total_AUC",
         Hyper_BTC = "HyperAUC",
         Hypo_BTC = "HypoAUC") |> 
  select(Bat_ID, AUC_BTC, Treatment, Group, Hyper_BTC, Hypo_BTC)
Basal_TC_1 <- Analysis_results |> 
  filter(DateType == "Basal TC + 1") |> 
  rename(AUC_BTC_1 = "Total_AUC",
         Hyper_BTC_1 = "HyperAUC",
         Hypo_BTC_1 = "HypoAUC") |> 
  select(Bat_ID, AUC_BTC_1, Treatment, Group, Hyper_BTC_1, Hypo_BTC_1)
```

** yeah I'm tired of this.... So since my current dataset is wide and a bat only has one row of data, it is not working as a random effect. 
```{r}
Full_1 <- inner_join(D0, D21, join_by(Bat_ID)) |> 
  inner_join(Basal_TC, join_by(Bat_ID))

# TC_bats_full <- lmer(AUC_D21 ~ Treatment + AUC_BTC + AUC_D0 + (1 | Bat_ID), 
#                      data = Full_1)
```
Okay, I talked to Ken and because I can't use Bat_ID as random effect I will try using treatment. However, I don't really want to 'ignore' treatment. But I'll run it anyways. 

```{r}
TC_bats_full_2 <- lmerTest::lmer(AUC_D21 ~ AUC_BTC + AUC_D0 + (1 | Treatment), 
                      data = Full_1)

summary(TC_bats_full_2)
```
Okay so this is similar to my secondary results that showed groups that received the same treatment at the secondary did not have a different thermal response at the secondary despite having different treatments at the primary. I am going to run this as just a glm and see what i get

```{r}
TC_bats_glm <- glm(AUC_D21 ~ Treatment + AUC_BTC + AUC_D0, 
                      data = Full_1)
summary(TC_bats_glm)
```
Again, this makes sense. CP+S and CS at the secondary is significanlty different than CP. But there is no correlation found between NTD on day 21 and D0 or BTC. The last thing I am going to do is just look at all my bats and only D0 vs D21. 

```{r}
Full_2 <- inner_join(D0, D21, join_by(Bat_ID)) |> 
  select(! Treatment.y) |> 
  select(! Group.y) |> 
  rename(Treatment = "Treatment.x",
         Group = "Group.x")

All_bats_glm <- glm(AUC_D21 ~ Treatment + AUC_D0, 
                      data = Full_2)

summary(All_bats_glm)
```
As suspected no correlation even when including bats that never underwent a thermal challenge. 



### Primary vs 1, 2, 3

Okay, now I am going to run models to see if primary effects days 1, 2, and 3. 

```{r}
Primary_days <- inner_join(D0, D1, join_by(Bat_ID)) |> 
  inner_join(D2, join_by(Bat_ID)) |> 
  inner_join(D3, join_by(Bat_ID)) |> 
  select(Bat_ID, AUC_D0, AUC_D1, AUC_D2, AUC_D3, Treatment.x, Group.x, HyperD0, HyperD1, HyperD2, HyperD3, HypoD0, HypoD1, HypoD2, HypoD3) |> 
  rename(Treatment = "Treatment.x",
         Group = "Group.x")

Primary_long_glm_1 <- glm(AUC_D1 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(Primary_long_glm_1)

Primary_long_glm_2 <- glm(AUC_D2 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(Primary_long_glm_2)

Primary_long_glm_3 <- glm(AUC_D3 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(Primary_long_glm_3)
```
Okay so these results shows that AUC on day 0 is correlated with auc on d1 but not d2 or d3. Since i saw significance in hyper auc on d1, 2 and 3, i want to run this same model looking at hyper and hypo. 

```{r}
interact_plot(Primary_long_glm_1, pred = AUC_D0, modx = Treatment, interval = TRUE, legend.main = "Treatment Group", plot.points = TRUE) +
  labs(x = "Primary AUC", y = "Primary+1 AUC") +
  theme_bw()
```

#### Hyper and hypo 

```{r}
HyperD1 <- glm(HyperD1 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(HyperD1)

HyperD2 <- glm(HyperD2 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(HyperD2)

HyperD3 <- glm(HyperD3 ~ Treatment + AUC_D0, 
                      data = Primary_days)
summary(HyperD3)
```




Similar to AUC, only hyper AUC on D1 was correlated with hyper AUC on D0. Not sure this is the best way to run this though....


### Secondary +1, 2, 3
```{r}
Secondary_Days <- inner_join(D21, D22, join_by(Bat_ID)) |> 
  inner_join(D23, join_by(Bat_ID)) |> 
  inner_join(D24, join_by(Bat_ID)) 

Secondary_Days <- Secondary_Days |> 
  select(Bat_ID, AUC_D21, AUC_D22, AUC_D23, AUC_D24, Treatment.x, Group.x) |> 
 rename(Group = "Group.x", 
        Treatment = "Treatment.x")
```

```{r}
Secondary_long_glm_1 <- glm(AUC_D22 ~ Treatment + AUC_D21, 
                      data = Secondary_Days)
summary(Secondary_long_glm_1)

Secondary_long_glm_2 <- glm(AUC_D23 ~ Treatment + AUC_D21, 
                      data = Secondary_Days)
summary(Secondary_long_glm_2)

Secondary_long_glm_3 <- glm(AUC_D23 ~ Treatment + AUC_D21, 
                      data = Secondary_Days)
summary(Secondary_long_glm_3)
```
## Weight Models

I am interested in if there was a difference in weight change between groups. So I am going to make a column that is delta Weight. 

```{r}
Analysis_results <- Analysis_results |> 
  arrange(Bat_ID, Day_Num) |> 
  group_by(Bat_ID) |> 
  mutate(Day_gap = lead(Day_Num) - Day_Num,
         weight_diff = if_else(Day_gap == 1,
                               lead(Weight) - Weight,
                               NA_real_)) |> 
  ungroup()

Weight_analysis <- Analysis_results |> 
  filter(ExperimentDay %in% c("D0", "D1", "D2", "D3", "D21", "D22", "D23", "D24")) |> mutate(Phase = case_when(ExperimentDay %in% c("D0", "D1", "D2", "D3") ~ "Primary",
                        ExperimentDay %in% c("D21", "D22", "D23", "D24") ~ "Secondary",
                        TRUE ~ NA_character_),
      Post_inject = case_when(
      ExperimentDay %in% c("D0", "D21") ~ 0,
      ExperimentDay %in% c("D1", "D22") ~ 1,
      ExperimentDay %in% c("D2", "D23") ~ 2,
      ExperimentDay %in% c("D3", "D24") ~ 3,
      TRUE ~ NA_real_ ))
```

```{r}
weight_model1 <- lmerTest::lmer(weight_diff ~ Treatment * HypoAUC + Treatment * HyperAUC + (1| Bat_ID), data = Weight_analysis)

summary(weight_model1)
```
Variance from bat_id is 0, so once I have accounted for treatment and hyper and hypo auc there is no difference left for bat_id to account for. Going to run as a regular lm first 

```{r}
weight_model_lm <- lm(
  weight_diff ~ Treatment * HyperAUC + Treatment * HypoAUC + ExperimentDay,
  data = Weight_analysis)

summary(weight_model_lm)
```
I ran these results through gemini and because my experiment days are different from each other, it recommended I add a days post injection and phase (primary or secondary). I'll run it again with this.. not sure if that is really allowed. 

```{r}
weight_model_lm2 <- lm(
  weight_diff ~ Treatment * HyperAUC + Treatment * HypoAUC + Phase + Post_inject,
  data = Weight_analysis)

summary(weight_model_lm2)
```
So there is no significant effect on weight including treatment. I originally asked gemini to include treatment but I'm not sure this was the best thing to do since I had bats that went hot even in cold chambers. However, dropping treatment groups sham bats with all other immunized bats so... I'll talk to Ken about this. 

```{r}
weight_model_lm3 <- lm(
  weight_diff ~ HyperAUC + HypoAUC + Phase + Post_inject,
  data = Weight_analysis)

summary(weight_model_lm3)
```



# Older models that I ran.... 
Okay, now running some models to look at AUC with treatment and date type.
```{r}
treatment_glm <- glm(Total_AUC ~ DateType * Treatment, data = Analysis_results)
summary(treatment_glm)
em_interaction <- emmeans(treatment_glm, ~ Treatment | DateType)
comps_treatment_glm <- pairs(em_interaction, adjust = "tukey")
 summary(comps_treatment_glm) 
 Treatxdatetype <- summary(comps_treatment_glm) 

glm_auc_txd <- as.data.frame(Treatxdatetype)

write.csv(glm_auc_txd, "glm_auc_txd.csv", row.names = FALSE)
 
#  adding weight into this model since we already established that it effects AUC
treatment_glm2 <- glm(Total_AUC ~ Weight + DateType * Treatment, data = Analysis_results)
summary(treatment_glm2)

em_interaction2 <- emmeans(treatment_glm2, ~ Treatment | DateType)

comps_treatment_glm2 <- pairs(em_interaction2, adjust = "tukey")
 summary(comps_treatment_glm2) 

Treatxdatetype2 <- summary(comps_treatment_glm2) 

glm_auc_txd2 <- as.data.frame(Treatxdatetype2)

write.csv(glm_auc_txd2, "glm_auc_txd2.csv", row.names = FALSE)
```


Rerunning the model above but including group. Because I had more bats in group 2 and group 1 bats were handled on D 3, 4 and 5 of the experiment but group 2 bats weren't there is likely a difference.
```{r}
treatment_glm3 <- glm(Total_AUC ~ Group + Weight + DateType * Treatment, data = Analysis_results)
summary(treatment_glm3)

em_interaction3 <- emmeans(treatment_glm3, ~ Treatment | DateType)

comps_treatment_glm3 <- pairs(em_interaction3, adjust = "tukey")
 summary(comps_treatment_glm3) 

Treatxdatetype3 <- summary(comps_treatment_glm3) 

glm_auc_txd3 <- as.data.frame(Treatxdatetype3)

```
Okay, this is great, but its running a lot of interactions that I do not care about, so I want to only run interesting interactions.

```{r}
# 1. Create a combined grouping variable
Analysis_results$Group_Combo <- factor(paste(Analysis_results$Treatment, Analysis_results$DateType, sep = "_"))

# 2. Run the model with the combined variable instead of the separate ones
treatment_glm4 <- glm(Total_AUC ~ Group + Weight + Group_Combo, data = Analysis_results)

levels(Analysis_results$Group_Combo)
```

```{r}
# Calculate the adjusted means for all 73 combinations
all_model_means <- emmeans(treatment_glm4, ~ Group_Combo)

all_pairs_tukey <- pairs(all_model_means, adjust = "tukey")

# ==============================================================================
# 4. SURGICAL STRIKES (SUBSET THEN PAIRS)
# ==============================================================================

### Hypothesis 1: Comparisons on Day 1 ###
# First, slice out just the Day 1 groups you care about
Primary1_model <- pairs(subset(all_model_means, 
  Group_Combo %in% c("CS_Primary", "NoCo_Primary", "SHAM_Primary")))

# Second, pair them up (R only calculates 3 comparisons here)
#pairs(Primary1_model, adjust = "tukey")


### Hypothesis 2: Comparisons on Day 2 ###
# Slice out just the Day 2 groups you care about
Primary2_model <- pairs(subset(
  all_model_means, 
  Group_Combo %in% c("CP_Primary", "CP+S_Primary", "NoCo_Primary", "CS_Primary")))

# Pair them up
#pairs(Primary2_model, adjust = "tukey")
```

```{r}
Secondary1_model <- pairs(subset(all_model_means, 
  Group_Combo %in% c("SHAM_Secondary", "NoCo_Secondary")))

# Pair them up
#pairs(Secondary1_model, adjust = "tukey")

Secondary2_model <- pairs(subset(all_model_means, 
  Group_Combo %in% c("NoCo_Secondary", "CP_Secondary")))

# Pair them up
#pairs(Secondary2_model, adjust = "tukey")

Secondary3_model <- pairs(subset(all_model_means, 
  Group_Combo %in% c("CP+S_Secondary", "CS_Secondary")))

# Pair them up
#pairs(Secondary3_model, adjust = "tukey")
```

```{r}
None_model <- pairs(subset(all_model_means, 
  Group_Combo %in% c("SHAM_None", "NoCo_None", "CP_None", "CS_None", "CP+S_None")))

#pairs(None_model, adjust = "tukey")
```

```{r D1 Models} 
D1_1 <- pairs(subset(all_model_means, 
  Group_Combo %in% c("CS_Primary + 1", "NoCo_Primary + 1", "SHAM_Primary + 1")))

# Second, pair them up (R only calculates 3 comparisons here)
#pairs(D1_1, adjust = "tukey")


### Hypothesis 2: Comparisons on Day 2 ###
# Slice out just the Day 2 groups you care about
D1_2 <- pairs(subset(
  all_model_means, 
  Group_Combo %in% c("CP_Primary + 1", "CP+S_Primary + 1", "NoCo_Primary + 1", "CS_Primary + 1")))

# Pair them up
#pairs(D1_2, adjust = "tukey")
```

```{r D2 Models} 
D2_1 <- pairs(subset(all_model_means, 
  Group_Combo %in% c("CS_Primary + 2", "NoCo_Primary + 2", "SHAM_Primary + 2")))

# Second, pair them up (R only calculates 3 comparisons here)
#pairs(D2_1, adjust = "tukey")


### Hypothesis 2: Comparisons on Day 2 ###
# Slice out just the Day 2 groups you care about
D2_2 <- pairs(subset(
  all_model_means, 
  Group_Combo %in% c("CP_Primary + 2", "CP+S_Primary + 2", "NoCo_Primary + 2", "CS_Primary + 2")))

# Pair them up
#pairs(D2_2, adjust = "tukey")
```

```{r D3 Models} 
D3_1 <- pairs(subset(all_model_means, 
  Group_Combo %in% c("CS_Primary + 3", "NoCo_Primary + 3", "SHAM_Primary + 3")))

# Second, pair them up (R only calculates 3 comparisons here)
#pairs(D3_1, adjust = "tukey")


### Hypothesis 2: Comparisons on Day 2 ###
# Slice out just the Day 2 groups you care about
D3_2 <- pairs(subset(
  all_model_means, 
  Group_Combo %in% c("CP_Primary + 3", "CP+S_Primary + 3", "NoCo_Primary + 3", "CS_Primary + 3")))

# Pair them up
#pairs(D3_2, adjust = "tukey")
```

```{r}
my_selected_comparisons <- rbind(None_model, Primary1_model, Primary2_model, D1_1, D1_2, D2_1, D2_2, D3_1, D3_2, Secondary1_model, Secondary2_model, Secondary3_model)

# 3. Run ONE unified Tukey test across just this curated pool
summary(my_selected_comparisons, adjust = "tukey")
```


I don't really want to go too far down running these models because I know some bats did not use torpor while in their thermal challenge. So p-values could be misleading since they are still in their assigned groups and not grouped by what they did.

## Max and Min models

```{r}
cor(Analysis_results[, c('Total_AUC', 'Day_max', 'Day_min')])
```
so minimum temp and AUC are probably correlated. 

```{r}
Temp_model1 <- glm(Total_AUC ~ Bat_ID + Day_max + Day_min, data = Analysis_results)
summary(Temp_model1)

stepAIC(Temp_model1, direction = "backward")

```

## Hyper and Hypo AUC

```{r}
cor(Analysis_results[, c('Total_AUC', 'HyperAUC', 'HypoAUC')])
```

I want to look at AUC for bats in basal TC and factor in cage.
```{r}
basal_analysis <- Analysis_results |> 
  filter(DateType == "Basal TC")

basal <- glm(Total_AUC ~ Cage, data = basal_analysis)
summary(basal)
#basal_comps <- glht(basal, linfct = mcp(Cage = "Tukey"))
#summary(basal_comps)
```
So it looks like cage did not effect AUC.

Okay, now I am going to look at tc for each treatment group.
```{r}
BP_TC_analysis <- Analysis_results |> 
  filter(DateType %in% c("Basal TC", "Primary") & Treatment %in% c("CP+S", "CP"))
CP_TC <- Analysis_results |> 
  filter(DateType %in% c("Basal TC", "Primary") & Treatment == "CP")
CPS_TC <- Analysis_results |> 
  filter(DateType %in% c("Basal TC", "Primary", "Secondary") & Treatment == "CP+S") |> 
  group_by(Bat_ID) |> 
  filter(n() >= 3)
CS_TC <- Analysis_results |> 
  filter(DateType %in% c("Basal TC", "Secondary") & Treatment == "CS") 


BP_Analysis <- glm(Total_AUC ~ DateType, data = BP_TC_analysis)
summary(BP_Analysis)
#comps <- glht(BP_Analysis, linfct = mcp(DateType = "Tukey"))
#summary(comps)

CP_TCglm <- glm(Total_AUC ~ DateType, data = CP_TC)
summary(CP_TCglm)
CP_TCglm_comp <- pairs(emmeans(CP_TCglm, ~ DateType), adjust = "tukey")
summary(CP_TCglm_comp)

CPS_TCglm <- glm(Total_AUC ~ DateType, data = CPS_TC)
summary(CPS_TCglm)
CPS_TCglm_comp <- pairs(emmeans(CPS_TCglm, ~ DateType), adjust = "tukey")
summary(CPS_TCglm_comp)

CS_TCglm <- glm(Total_AUC ~ DateType, data = CS_TC)
summary(CS_TCglm)
CS_TCglm_comp <- pairs(emmeans(CS_TCglm, ~ DateType), adjust = "tukey")
summary(CS_TCglm_comp)
```
Okay! Clearly a difference in AUC when comparing basal TC to primary TC!



```{r}
SHAM <- Analysis_results |> 
  filter(Treatment == "SHAM")

NoCo <- Analysis_results |> 
  filter(Treatment == "NoCo")

CPS <- Analysis_results |> 
  filter(Treatment == "CP+S")

CP <- Analysis_results |> 
  filter(Treatment == "CP")

CS <- Analysis_results |> 
  filter(Treatment == "CS")
```

```{r}
SHAM_glm <- glm(Total_AUC ~ Bat_ID , data = SHAM)
summary(SHAM_glm)
```

```{r}
NoCO_glm <- glm(Total_AUC ~ Bat_ID , data = NoCo)
summary(NoCO_glm)
```

```{r}
CPS_glm <- glm(Total_AUC ~ Bat_ID , data = CPS)
summary(CPS_glm)
```

```{r}
CP_glm <- glm(Total_AUC ~ Bat_ID , data = CP)
summary(CP_glm)
```

```{r}
CS_glm <- glm(Total_AUC ~ Bat_ID , data = CS)
summary(CS_glm)
```

```{r}
Primary_comp <- Analysis_results |> 
  filter(DateType == "Primary")

Sec_comp <- Analysis_results |> 
  filter(DateType == "Secondary")

PS_Compare <- inner_join(Primary_comp, Sec_comp, join_by(Bat_ID))
PS_glm <- glm(Total_AUC.y ~ Total_AUC.x, data = PS_Compare)
summary(PS_glm)
emtrends(PS_glm, ~ Total_AUC.x, var = "Total_AUC.x")
```
Not significant, but i want to look at this with only TC bats. Also I want to factor treatment into it. 

```{r}
PS_glm2 <- glm(Total_AUC.y ~ Total_AUC.x *Treatment.x, data = PS_Compare)
summary(PS_glm2)
emtrends(PS_glm2, ~ Treatment.x, var = "Total_AUC.x")
```
Again no significance.

```{r}
TC_primary <- Primary_comp |> 
  filter(Treatment %in% c("CP+S", "CP", "CS"))

TC_sec <- Sec_comp |> 
  filter(Treatment %in% c("CP+S", "CP", "CS"))

TC_PS_Compare <- inner_join(TC_primary, TC_sec, join_by(Bat_ID))
TC_PS_glm <- glm(Total_AUC.y ~ Total_AUC.x, data = TC_PS_Compare)
summary(TC_PS_glm)
emmeans(TC_PS_glm, ~ Total_AUC.x, var = "Total_AUC.x")
```
Okay, this shows significance! 

I am going to now take average AUC for primary to primary +3 and secondary to secondary +3

```{r}
Full_primary <- Analysis_results |> 
  filter(DateType %in% c("Primary", "Primary + 1", "Primary + 2", "Primary + 3")) |> 
  group_by(Bat_ID) |> 
  mutate(Primary_AUC = mean(Total_AUC)) |> 
  filter(DateType == "Primary") |> 
  ungroup()

Full_secondary <- Analysis_results |> 
  filter(DateType %in% c("Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")) |> 
  group_by(Bat_ID) |> 
  mutate(Secondary_AUC = mean(Total_AUC)) |> 
  filter(DateType == "Secondary") |> 
  ungroup()

PS_full <- inner_join(Full_primary, Full_secondary |> select(Bat_ID, Secondary_AUC), join_by("Bat_ID"))

PS_full_glm <- glm(Secondary_AUC ~ Primary_AUC * Treatment, data = PS_full)
summary(PS_full_glm)
emmeans(PS_full_glm, ~ Treatment | Primary_AUC)

pairs(emmeans(PS_full_glm, ~ Treatment | Primary_AUC, at = list(Primary_AUC = -140)))
```

```{r}
# 1. Visualize the distribution to find the "cut point"
hist(PS_Compare$Total_AUC.x, breaks = 20, main = "Distribution of Primary Torpor")

# 2. Filter for 'Responders'
CPandNOCO_compare <- PS_Compare |> 
    filter(Treatment.x == "NoCo" | Treatment.x == "CP" & Total_AUC.x < -200) |> 
  mutate(Primary_ratio = Total_AUC.x/Weight.x,
         Secondary_ratio = Total_AUC.y/Weight.y)

```


```{r}
CPandNOCO_compare_glm2 <- glm(Total_AUC.y ~ Weight.y + Total_AUC.x * Treatment.x, data = CPandNOCO_compare)
summary(CPandNOCO_compare_glm2)
CPandNOCO <-pairs(emmeans(CPandNOCO_compare_glm2, ~ Treatment.x | Total_AUC.x)) 
summary(CPandNOCO)

#Fix this...
#interact_plot(CPandNOCO_compare_glm2, pred = Total_AUC.x, modx = Treatment.x, interval = #TRUE, legend.main = "Treatment Group", plot.points = TRUE) +
#   xlim(-1800, 400) + 
#  ylim(-2000, 200) +
#  labs(x = "Primary AUC", y = "Secondary AUC") +
#  # Make it look professional
#  theme_bw()

CPandNOCO_compare_glm <- glm(Total_AUC.y ~ Total_AUC.x * Treatment.x, data = CPandNOCO_compare)
emmeans(CPandNOCO_compare_glm, ~ Treatment.x | Total_AUC.x)
pairs(emmeans(CPandNOCO_compare_glm, ~ Treatment.x | Total_AUC.x))
```

```{r}
Primary_Secondary_hist <- Analysis_results |> 
  filter(DateType %in% c("Primary", "Primary + 1", "Primary + 2", "Primary + 3", "Secondary", "Secondary + 1", "Secondary + 2", "Secondary + 3")) |> 
  mutate(DateType = fct_relevel(DateType, "Primary", "Secondary", "Primary + 1", "Secondary + 1", "Primary + 2", "Secondary + 2", "Primary + 3",  "Secondary + 3")) |> 
  mutate(Treatment = fct_relevel(Treatment, "CP", "CP+S", "CS", "NoCo", "SHAM"))

Hist <- ggplot(Primary_Secondary_hist, aes(x= Total_AUC, fill = Treatment)) +
  geom_histogram(binwidth = 200) +
  facet_wrap(~DateType, ncol = 2) +
  labs(x ="Daily AUC", y = "Count") +
  scale_fill_manual(values = treatmentcolors) +
  theme(text = element_text(size = 18), panel.spacing.x = unit(2, "lines"))


ggsave("EPFU_Histogram.jpeg", plot = Hist, device = jpeg(), width = 8, height = 6.67)

```


## Willis equation
I am interested to see if there is any difference in 

## Basal Temp

```{r}
Metadata <- Metadata |> 
  mutate(id = Bat_ID) |> 
  mutate(id = as.character(id)) |> 
    mutate(Treatment = fct_relevel(Treatment, "CP", "CP+S", "CS", "NoCo", "SHAM"))

Metadata2 <- Metadata |> 
  left_join(
    Analysis_results |> select(Bat_ID, Tnorm) |> distinct(), 
    by = "Bat_ID")

None_data <- Results |> 
  filter(DateType == "None") |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

none_met <- data.table::data.table(Metadata2, key = "id")
none_data2 <- data.table::data.table(None_data, key = "id")

 none_behavr <- behavr(none_data2, none_met)

none_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment, fill = Treatment)) +
  geom_hline(aes(yintercept = Tnorm), 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.8) +
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")
  ) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Date/time",
    # Start at the minimum time in your data, end at max, move by 2 hours (7200s)
    breaks = seq(min(none_behavr$t), max(none_behavr$t), by = 7200), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%H:%M"),
    expand = c(0,0)
  ) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(text = element_text(size = 18), panel.grid = element_blank())+
  labs(y = "Temp (C)") -> EPFU_none

print(EPFU_none)


ggsave("EPFU_none.jpeg", plot = EPFU_none, device = jpeg(), width = 12, height = 6.67)
```


```{r}
Basal_TC <- Basal_TC |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

Basal_met <- data.table::data.table(Metadata2, key = "id")
Basal_TC2 <- data.table::data.table(Basal_TC, key = "id")

 basal_behavr <- behavr(Basal_TC2, Basal_met)
```

```{r}
basal_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment, fill = Treatment)) +
  geom_hline(aes(yintercept = Tnorm), 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.8) +
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")
  ) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Date/time",
    # Start at the minimum time in your data, end at max, move by 2 hours (7200s)
    breaks = seq(min(basal_behavr$t), max(basal_behavr$t), by = 7200), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%H:%M"),
    expand = c(0,0)
  ) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(text = element_text(size = 18), panel.grid = element_blank())+
  labs(y = "Temp (C)") -> EPFU_Basal

print(EPFU_Basal)


ggsave("EPFU_Basal.jpeg", plot = EPFU_Basal, device = jpeg(), width = 12, height = 6.67)
```


##Primary Temp

```{r}

Primary_Results <- Primary_Results |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 Primary_met <- data.table::data.table(Metadata2, key = "id")
 Primary_results2 <- data.table::data.table(Primary_Results, key = "id")

 primary_behavr <- behavr(Primary_results2, Primary_met) 
 summary(primary_behavr)
 
 glimpse(primary_behavr)


```

### Primary Behavr 

```{r}
shaded <- data.frame(Treatment = c("CP", "CP+S"), xmin = 1757344201 , 
  xmax = 1757422801,#adds shaded regions for both the basal and primary thermal #challenges
                     label = c("Primary TC"))

primary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment, fill = Treatment)) +
  
  # 1. Shaded region
  geom_rect(data = shaded, 
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
            
  # 2. Baseline line (Moved colour OUTSIDE aes)
  geom_hline(aes(yintercept = Tnorm), 
             colour = "black",        # Static color, won't show in legend
             linetype = "dashed", 
             linewidth = 0.8) +
             
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Date/Time",
    breaks = seq(1757253601, to = max(primary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%H:%M"),
    expand = c(0,0) ) +
  # 3. Use your original treatmentcolors without the "Tnorm" entry
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    text = element_text(size = 16),
    panel.grid = element_blank()) +
  labs(y = "Temp (C)", colour = "Treatment", fill = "Treatment") -> EPFU_Primary

print(EPFU_Primary)
 
ggsave("EPFU_Primary.jpeg", plot = EPFU_Primary, device = jpeg(), width = 10, height = 6.67)

```

## Behavr Analysis

```{r}

Secondary_Results <- Secondary_Results |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 Secondary_met <- data.table::data.table(Metadata2, key = "id")
 Secondary_results2 <- data.table::data.table(Secondary_Results, key = "id")

 secondary_behavr <- behavr(Secondary_results2, Secondary_met)
 summary(secondary_behavr)
```

```{r}
secondary_etho  <- ggetho(secondary_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho() +
  scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(fill = "Temp (C)",
       y = "Bat ID")

ggsave("EPFU_Sec_Etho.jpeg", plot = secondary_etho, device = jpeg(), width = 12, height = 6.67)
```

```{r}
ggetho(secondary_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r, warning=FALSE}

shaded2 <- data.frame(Treatment = c("CS", "CP+S"), xmin = 1759155601 , 
  xmax = 1759234201,#adds shaded regions for both the basal and primary thermal #challenges
  label = c("Secondary TC"))

secondary_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
   geom_hline(aes(yintercept = Tnorm), 
             colour = "black",        # Static color, won't show in legend
             linetype = "dashed", 
             linewidth = 0.8) +
  geom_rect(data = shaded2, 
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
            fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 39600,
    l_duration = 43200, 
    ld_colours = c("black", "white")) +
  facet_grid(Treatment ~ .) +
  scale_x_continuous(
    name = "Date/Time",
    # Start at the minimum time in your data, end at max, move by 2 hours (7200s)
    breaks = seq(1758895201, to = max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%H:%M"),
    expand = c(0,0)) +
  scale_colour_manual(values = treatmentcolors) +
  scale_fill_manual(values = treatmentcolors) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(text = element_text(size = 16), panel.grid = element_blank())+
  labs(y = "Temp (C)") -> EPFU_Secondary

print(EPFU_Secondary)
 
ggsave("EPFU_Secondary.jpeg", plot = EPFU_Secondary, device = jpeg(), width = 10, height = 6.67)
```

```{r, warning=FALSE}
None_test <- Results |> 
  filter(DateType == "None")

None_test <- None_test |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))

 none_met <- data.table::data.table(Metadata2, key = "id")
 None_test2 <- data.table::data.table(None_test, key = "id")

 none_behavr <- behavr(None_test2, none_met)
 summary(none_behavr)

ggetho(none_behavr, aes(y= Temperature, colour= Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
     # facet_grid(id ~ .) +
      scale_y_continuous(name= "Temperature")
```


```{r}
ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(Treatment ~ .) +
      scale_y_continuous(name= "Temperature")
ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      scale_y_continuous(name= "Temperature")

Secondary_Trimmed <- Secondary_results2 |> 
  filter(NumberDate <= "1749254400") |> 
  mutate(id = Bat_ID) |> 
  mutate(t = NumberDate) |> 
  mutate(id = as.character(id))


 Secondary_results3 <- data.table::data.table(Secondary_Trimmed, key = "id")

 secondary_behavr2 <- behavr(Secondary_results3, Secondary_met2)
 summary(secondary_behavr)

 
 ggetho(secondary_behavr, aes(y= Temperature, colour=Treatment), time_wrap = 86400) +
      stat_pop_etho() +
      stat_ld_annotations() +
      scale_y_continuous(name= "Temperature")
```

```{r Sham, warning=FALSE}

met_sham <- Secondary_met2 %>%
  filter(Treatment == "SHAM")
sham_results <- Secondary_results2 %>%
  filter(id %in% met_sham$id)
sham_behavr <- behavr(sham_results, met_sham)

sham_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r eVLP_LPS, warning=FALSE}
met_LPS <- Secondary_met2 %>%
  filter(Treatment == "eVLP+LPS")
LPS_results <- Secondary_results2 %>%
  filter(id %in% met_LPS$id)
LPS_behavr <- behavr(LPS_results, met_LPS)

LPS_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r TC, warning=FALSE}
met_TC <- Secondary_met2 %>%
  filter(Treatment == "TC")
TC_results <- Secondary_results2 %>%
  filter(id %in% met_TC$id)
TC_behavr <- behavr(TC_results, met_TC)

TC_behavr %>%
 ggetho(aes(x = t, y = Temperature, colour = id))+
  stat_pop_etho() +
  stat_ld_annotations(
    period = 86400,
    phase = 25200,
    l_duration = 43200, 
    ld_colours = c( "white", "black")) +
  facet_grid(id ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = seq(1748977200, max(secondary_behavr$t), by = 86400), 
    labels = function(x) format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%b %d, %H:%M"),
    expand = c(0,0)) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r, warning=FALSE}
ggetho(bat_behavr, aes(y=Temperature, colour=Treatment)) + 
  stat_pop_etho()
```
