---
title: "MOCO 2025 Temperature Loggers Individual"
author: "Ashlyn Sak"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("plotly")) install.packages("plotly"); library(plotly)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("mice")) install.packages("mice"); library(mice)
if (!require("lubridate")) install.packages("lubridate"); library(lubridate)
if (!require("openair")) install.packages("openair"); library(openair)
if (!require("latticeExtra")) install.packages("latticeExtra"); library(latticeExtra)
if (!require("behavr")) install.packages("behavr"); library(behavr)
if (!require("ggetho")) install.packages("ggetho"); library(ggetho)
if (!require("damr")) install.packages("damr"); library(damr)
if (!require("sleepr")) install.packages("sleepr"); library(sleepr)
if (!require("timetk")) install.packages("timetk"); library(timetk)
if (!require("lme4")) install.packages("lme4"); library(lme4)
if (!require("merTools")) install.packages("merTools"); library(merTools)
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
conflict_prefer_all("dplyr", quiet = TRUE)
conflicts_prefer(plotly::layout)
conflicts_prefer(lubridate::hours)
conflicts_prefer(lubridate::days)
```

Check to make sure that all of those packages are needed, once I am done.

The goal here is to look at each logger individually to verify that the data is clean.

Then I will plot the pre- and post- boost data for each individual.

## Metadata

```{r}
Metadata <- read_excel("EPFU 2025 BUTT Assignments.xlsx", sheet = "Loggers", 
    range = "A1:L58", na = "NA")
glimpse(Metadata)

Metadata$Bat_ID = as.character(Metadata$BatID)

glimpse(Metadata)
```


```{r}
DATA_DIR <- "./MOCO Data/Edited Data" # load in each logger file

file_list <- as_tibble(list.files(DATA_DIR)) |>
  mutate(Bat_ID = gsub("MOCO_Band", "", value)) |> # removes everything after bat ID
  mutate(Bat_ID = gsub(".csv", "", Bat_ID)) |>  #removes .csv
  rename(Filename = value) 

# file_list$Bat_ID = as.double(file_list$Bat_ID)

Metadata <- left_join(Metadata, file_list, by = join_by(Bat_ID == Bat_ID)) # adds file name into metadata

batfilelist <- list.files(path = "./MOCO Data/Edited Data", 
                          pattern = "*.csv", full.names = T) # shows full file path name
batfilenames <- list.files(path = "./MOCO Data/Edited Data", 
                           pattern = "*.csv", full.names = F) # shows only file name
batfiles <- lapply(batfilelist, fread, 
                   col.names = c("Bat_ID", "DateTime","Unit", "Temperature"), sep = "," , 
                   skip = 20) # applies columns to each file, skips the first 20 rows

batnames <- gsub("MOCO_Band", "", batfilenames) # removes MOCO_Band from file name

batnames = gsub(".csv", "", batnames) # removes .csv from file name and leaves only Band 

names(batfiles) <- batnames # names each file by logger name
allbatresults <- bind_rows(batfiles, .id = "Bat_ID") # creates a long dataset with each time point for every logger, 4 variables 

allbatresults <- allbatresults %>%
  mutate(DateTime = mdy_hm(DateTime)) # changes date to appropriate format
str(allbatresults) # similar to glimpse
summary(as.factor(allbatresults$Bat_ID)) # display all bat names and their counts use to verify they ran correctly
md.pattern(allbatresults) # checking for NA's
Metadata %>%
  filter(!Bat_ID%in% file_list$Bat_ID) %>% # only selects loggers with a matching file name
  drop_na(`Filename`) # removes all na's
Metadata %>%
  filter(is.na(`Filename`)) # shows which bats do not have a logger
```

Check files for bands 79, 80, 68, and 74. These loggers have 4,000 datapoints instead of 2,000. Could be because they were used during initial innoculation for aim 3 bats. 

The bats with no files no files represent loggers that failed to download, 

The logger numbers with no bat ID have not been identified

## Missing Data

```{r}
no_hot_met <- Metadata |> 
  filter(!Treatment =="Hot")

clean_met <- no_hot_met %>%
  drop_na(Filename) # removes any ibuttons with no file

head(clean_met) # look at the top of the dataset
md.pattern(clean_met, rotate.names = TRUE)
write_tsv(clean_met, file = "MOCO2025TempLoggerMetadata.tsv")
```

I am left with 34 ibuttons with files. There are some NAs in my database but they shouldn't interfere with the data right now. NAs are in the notes column, boost column, date removed column 

## Making a plot function

```{r}
plot_fun <- function(arg1) {
  allbatresults %>%
    dplyr::filter(Bat_ID == arg1) %>%
    ggplot(aes(x=DateTime, y=Temperature)) +
    geom_line() +
    ylab("Temperature (C)") + xlab("Date") + ylim(4,40) +
    theme_cowplot() 
}
```

```{r}
plotly_fun <- function(arg1) {
  plot_ly(data = allbatresults[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))
}
```

## Looking at the data

```{r Bat 03}
ggplotly(plot_fun("3"))
```

```{r Plotly Bat 16}
plotly_fun("16")
```

Here we go!

This works fine, but is not interactive: lapply(batnames2023, plot_fun)

```{r}
lapply(batnames, plotly_fun)
```

If you can't see the whole plot, use the "Show in New Window" icon above the output.

Aim 3 inoculation was on 26-May-2025
Aim 3 boost was on 16-Jun-2025
Group 1 boost was on 5-Jun-2025
Group 2 boost was on 7-Jun-2025

64.2 Data needs to be trimmed, some points dropping to -40. Rest of data looks good
Delete 73.3, logger only stayed on for a few minutes.
76.2 Data needs to be trimmed, some points dropping to -30. Rest of data looks good. 
I think as long as they were on for the 24 hours after the boost we should keep the data
Delete 78.2, logger only stayed on for a few hours. 

```{r Plotly 3}
plotly_fun("3")
```


```{r Plotly 54}
plotly_fun("54")
```

```{r Plotly MST60101}
plotly_fun("78.2")
```

Additional things to consider:

39- Logger fell off and reattached 7-Jun
47- Logger fell off and reattached on 6-Jun
48 - Logger fell off and reattached on 7-Jun

## Left and Right Trimming the data

I will split the data, filter it, and then recombine them.

```{r}

clean_met <- clean_met %>%
  mutate(Attachment_Date = as.numeric(`Date Attached`),
         Removal_Date = as.numeric(`Date Removed`)) 

allbatresults <- allbatresults %>%
  mutate(NumberDate = as.numeric(DateTime))

temps_with_meta <- allbatresults %>%
  left_join(clean_met, by = "Bat_ID")

# Filter to keep only records after attachment date
left_trimmed <- temps_with_meta %>%
  filter(NumberDate >= Attachment_Date)

right_trimmed <- left_trimmed |> 
  filter(NumberDate <= Removal_Date)

allbatresults2 <- right_trimmed |> 
  select(Bat_ID, DateTime, NumberDate, Temperature)

summary(as.factor(allbatresults2$Bat_ID))

```

```{r}
plotly_fun2 <- function(arg1) {
  plot_ly(data = allbatresults2[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))}
```


```{r}
lapply(batnames, plotly_fun2)
```
## Left trim again

I left trimmed earlier to remove all data from before the day of attachment. I really want to include as much as possible so I am trimming again by the attachment time. This is going to take forever,,, :/

```{r}

# first trim
Trimming1 <- allbatresults2 |> 
  filter(Bat_ID %in% c("22", "16", "21", "3", "9"))
summary(as.factor(Trimming1$Bat_ID))

Trimming1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim1 <- Trimming1 |>
  filter(NumberDate >= "1748967000")

Trim1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

summary(as.factor(Trim1$Bat_ID))


#second trim

Trimming2 <- allbatresults2 |> 
  filter(Bat_ID %in% c("39", "41", "44"))
summary(as.factor(Trimming2$Bat_ID))

Trimming2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim2 <- Trimming2 |>
  filter(NumberDate >= "1749151200")

Trim2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#third trim

Trimming3 <- allbatresults2 |> 
  filter(Bat_ID %in% c("46"))
summary(as.factor(Trimming3$Bat_ID))

Trimming3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim3 <- Trimming3 |>
  filter(NumberDate >= "1749154800")

Trim3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#fourth trim

Trimming4 <- allbatresults2 |> 
  filter(Bat_ID %in% c("47", "48"))
summary(as.factor(Trimming4$Bat_ID))

Trimming4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim4 <- Trimming4 |>
  filter(NumberDate >= "1749229200")

Trim4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#fifth trim

Trimming5 <- allbatresults2 |> 
  filter(Bat_ID %in% c("49"))
summary(as.factor(Trimming5$Bat_ID))

Trimming5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim5 <- Trimming5 |>
  filter(NumberDate >= "1748978400")

Trim5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# sixth trim
Trimming6 <- allbatresults2 |> 
  filter(Bat_ID %in% c("50", "60", "62"))
summary(as.factor(Trimming6$Bat_ID))

Trimming6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim6 <- Trimming6 |>
  filter(NumberDate >= "1749220800")

Trim6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# seventh trim 

Trimming7 <- allbatresults2 |> 
  filter(Bat_ID %in% c("54", "55"))
summary(as.factor(Trimming7$Bat_ID))

Trimming7 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim7 <- Trimming7 |>
  filter(NumberDate >= "1749219000")

Trim7 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# eighth trim
Trimming8 <- allbatresults2 |> 
  filter(Bat_ID %in% c("64.1"))
summary(as.factor(Trimming8$Bat_ID))

Trimming8 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim8 <- Trimming8 |>
  filter(NumberDate >= "1747736400")

Trim8 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Ninth trim

Trimming9 <- allbatresults2 |> 
  filter(Bat_ID %in% c("64.2"))
summary(as.factor(Trimming9$Bat_ID))

Trimming9 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim9 <- Trimming9 |>
  filter(NumberDate >= "1749923400")

Trim9 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Tenth trim

Trimming10 <- allbatresults2 |> 
  filter(Bat_ID %in% c("68.1", "74.1"))
summary(as.factor(Trimming10$Bat_ID))

Trimming10 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim10 <- Trimming10 |>
  filter(NumberDate >= "1747741800")

Trim10 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 11
Trimming11 <- allbatresults2 |> 
  filter(Bat_ID %in% c("68.2", "74.2", "78.2"))
summary(as.factor(Trimming11$Bat_ID))

Trimming11 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim11 <- Trimming11 |>
  filter(NumberDate >= "1749919200")

Trim11 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Trim 12 

Trimming12 <- allbatresults2 |> 
  filter(Bat_ID %in% c("73.4"))
summary(as.factor(Trimming12$Bat_ID))

Trimming12 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim12 <- Trimming12 |>
  filter(NumberDate >= "1750091400")

Trim12 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 13

Trimming13 <- allbatresults2 |> 
  filter(Bat_ID %in% c("76.2"))
summary(as.factor(Trimming13$Bat_ID))

Trimming13 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim13 <- Trimming13 |>
  filter(NumberDate >= "1749987600")

Trim13 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 14

Trimming14 <- allbatresults2 |> 
  filter(Bat_ID %in% c("78.3"))
summary(as.factor(Trimming14$Bat_ID))

Trimming14 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim14 <- Trimming14 |>
  filter(NumberDate >= "1750006800")

Trim14 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Trim 15

Trimming15 <- allbatresults2 |> 
  filter(Bat_ID %in% c("83.2"))
summary(as.factor(Trimming15$Bat_ID))

Trimming15 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim15 <- Trimming15 |>
  filter(NumberDate >= "1749920400")

Trim15 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Trim 16

Trimming16 <- allbatresults2 |> 
  filter(Bat_ID %in% c("26"))
summary(as.factor(Trimming16$Bat_ID))

Trimming16 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

Trim16 <- Trimming16 |>
  filter(NumberDate >= "1748968200")

Trim16 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))


LTrimmed <- bind_rows(Trim1, Trim2, Trim3, Trim4, Trim5, Trim6, Trim7, Trim8, Trim9, Trim10, Trim11, Trim12, Trim13, Trim14, Trim15, Trim16)



summary(as.factor(LTrimmed$Bat_ID))
summary(as.factor(allbatresults2$Bat_ID))
```


## Right trim again 

```{r}
# Right trim 1
RTrimming1 <- LTrimmed |> 
  filter(Bat_ID %in% c("64.1"))
summary(as.factor(RTrimming1$Bat_ID))

RTrimming1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim1 <- RTrimming1 |>
  filter(NumberDate <= "1749673200")

RTrim1 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right trim 2
RTrimming2 <- LTrimmed |> 
  filter(Bat_ID %in% c("68.1"))
summary(as.factor(RTrimming2$Bat_ID))

RTrimming2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim2 <- RTrimming2 |>
  filter(NumberDate <= "1748880600")

RTrim2 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right trim 3

RTrimming3 <- LTrimmed |> 
  filter(Bat_ID %in% c("78.2"))
summary(as.factor(RTrimming3$Bat_ID))

RTrimming3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim3 <- RTrimming3 |>
  filter(NumberDate <= "1749931200")

RTrim3 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# Right Trim 4 

RTrimming4 <- LTrimmed |> 
  filter(Bat_ID %in% c("22"))
summary(as.factor(RTrimming4$Bat_ID))

RTrimming4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim4 <- RTrimming4 |>
  filter(NumberDate <= "1749742800")

RTrim4 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# RTrim 5

RTrimming5 <- LTrimmed |> 
  filter(Bat_ID %in% c("21"))
summary(as.factor(RTrimming5$Bat_ID))

RTrimming5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim5 <- RTrimming5 |>
  filter(NumberDate <= "1749439200")

RTrim5 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

# RTrim 6

RTrimming6 <- LTrimmed |> 
  filter(Bat_ID %in% c("3"))
summary(as.factor(RTrimming6$Bat_ID))

RTrimming6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

RTrim6 <- RTrimming6 |>
  filter(NumberDate <= "1749276000")

RTrim6 |> summarize(minTimepoint = min(DateTime), maxTimepoint = max(DateTime))

#Keeping all results from LTrimmed besides the ones that were right trimmed again

RTrimmed_Temp <- LTrimmed |> 
  filter(!Bat_ID %in% c("64.1", "68.1", "78.2", "22", "21", "3"))

summary(as.factor(RTrimmed_Temp$Bat_ID))

RTrimmed <- bind_rows(RTrimmed_Temp, RTrim1, RTrim2, RTrim3, RTrim4, RTrim5, RTrim6)

summary(as.factor(RTrimmed$Bat_ID))

```

## Dropping bad data and gaps

```{r}

# Dropping 1
Dropping1 <- RTrimmed |> 
  filter(Bat_ID %in% c("68.2"))

summary(as.factor(Dropping1$Bat_ID))

Drop1.1 <- Dropping1 |>
  filter(NumberDate <= "1750024800")
Drop1.2 <- Dropping1 |> 
  filter(NumberDate >= "1750090800")

Drop1.final <- bind_rows(Drop1.1, Drop1.2)

#Dropping 2

Dropping2 <- RTrimmed |> 
  filter(Bat_ID %in% c("83.2"))

summary(as.factor(Dropping2$Bat_ID))

Drop2.1 <- Dropping2 |>
  filter(NumberDate <= "1750009200")
Drop2.2 <- Dropping2 |> 
  filter(NumberDate >= "1750090800")

Drop2.final <- bind_rows(Drop2.1, Drop2.2)

# Dropping bad data

Bad_Drop <- RTrimmed |> 
  filter(Bat_ID %in% "64.2") |> 
  filter(Temperature > "10") |> 
  filter(Temperature < "45")

# Removing the edited data in this chunk from RTrimmed so I can merge all the data back together

allbatresults3.temp <- RTrimmed |> 
  filter(!Bat_ID %in% c("64.2", "68.2", "83.2"))

clean_results <- bind_rows(allbatresults3.temp, Bad_Drop, Drop2.final, Drop1.final)

summary(as.factor(clean_results$Bat_ID))
```

Making a new plotly function for these results. 

```{r}
plotly_fun3 <- function(arg1) {
  plot_ly(data = clean_results[Bat_ID == arg1],
          x = ~DateTime, y = ~Temperature,
          type = 'scatter', mode = 'lines') %>%
    layout(xaxis = list(title = arg1))}
```

```{r}
lapply(batnames, plotly_fun3)

```

Consider dropping or trimming 21 and 3

Okay, that is all as expected.
Time to retrim that data.

Added the dates each logger fell off or was removed in the excel file. 



```{r}
write_tsv(clean_results, file = "TrimmedTemperatureLoggerData.long.tsv")
```

##Cold bats initial

```{r}
# Cold_Bats <- clean_results |> 
#   filter(Bat_ID %in% c("64.1", "68.1", "74.1")) |> 
#   rename(id = Bat_ID) |> 
#   rename(t = NumberDate)
# 
# Cold_initial_Results <- Cold_Bats |> 
#   filter(DateTime <= "2025-05-27 23:50:00")
#   
# 
# ColdBatNames <- unique(Cold_initial_Results$id)
# 
# Cold_met <- clean_met %>%
#   rename(id = Bat_ID) %>%
#   filter(id %in% ColdBatNames)
#   
# Cold_data <- Cold_met %>%
#   select(id) %>%
#   left_join(Cold_initial_Results, by = join_by(id))
# str(Cold_data)
# 
# cold_met2 <- data.table::data.table(Cold_met, key = "id")
# cold_data2 <- data.table::data.table(Cold_data, key = "id")
# 
# cold_behavr <- behavr(cold_data2, cold_met2)
# summary(cold_behavr)
```

### Cold Bat Behavr 

```{r}

# ticks_7am <- seq(
#   from = floor_date(as.POSIXct(min(cold_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
#   to   = ceiling_date(as.POSIXct(max(cold_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
#   by   = "1 day"
# )
# 
# 
# 
# cold_behavr |>
#   ggetho(aes(x = t, y = Temperature, colour = id)) +
#   stat_pop_etho() +
#   stat_ld_annotations(
#     ld_colours = c("white", "black"),
#        # 12h light : 12h dark, in seconds
#     offset = 7 * 3600) +           # <<< shift start by 7 hours (7 AM)
#   facet_grid(id ~ .) +
#     scale_x_continuous(
#     name = "Date/time",
#     breaks = as.numeric(ticks_7am),
#     labels = function(x)
#       format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) -> MOCO_Cold
# 
#  ggsave("MOCO_Cold.jpeg", plot = MOCO_Cold, device = jpeg(), width = 12, height = 6.67)
#  
#  
#  cold_behavr |>
#   ggetho(aes(x = t, y = Temperature, colour = id)) +
#   stat_pop_etho() +
#   stat_ld_annotations(
#     ld_colours = c( "black", "white")) +
#   facet_grid(id ~ .) +
#     scale_x_continuous(
#     name = "Date/time",
#     breaks = as.numeric(ticks_7am),
#     labels = function(x)
#       format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) +  theme(
#     axis.text = element_text(size = 14),       # axis tick labels
#     axis.title = element_text(size = 16),      # axis titles
#     legend.text = element_text(size = 13),     # legend labels
#     legend.title = element_text(size = 15),    # legend title
#     strip.text = element_text(size = 12)       # facet labels
#   )  -> MOCO_Cold
# 
# ggsave("MOCO_Cold.jpeg", plot = MOCO_Cold, device = jpeg(), width = 12, height = 6.67)
```



## Ambient Temp

```{r}
# Union_Ambient <- read_excel("MOCO Data/Roost Temps/MOCO_Union_Ambient.xlsx")
# 
# results <- clean_results |> 
#   left_join(Union_Ambient, by = join_by(DateTime))
#   left_join(clean_results, Union_Ambient, join_by(DateTime == Date/Time))

```

## Combining Data

I am going to combine data from multiple loggers for one bat into one long dataset. It looks like this only needs to be done for bat 78

```{r}
# DO LATER NOT WORKING NOW
# summary(as.factor(clean_results$Bat_ID))
# 
# Bat78_Boost <- clean_results |> 
#   filter(Bat_ID %in% c("78.2", "78.3"))
# 
# ## Bat78_Boost <- Bat78_Boost |> 
#   mutate(Bat_ID == "78")
# split_data <- clean_results |> 
#   filter(!Bat_ID %in% c("78.2", "78.3"))

```


### Shifting Data


```{r}
boostbat <- clean_results |> 
  filter(!Bat_ID %in% c("64.1", "68.1", "74.1"))

  
boostbatnames <- unique(boostbat$Bat_ID)

boost_met <- clean_met %>%
  filter(Bat_ID %in% boostbatnames)

summary(as.factor(boost_met$Treatment))

boostresultsmet <- boost_met |> 
left_join(boostbat, by = join_by(Bat_ID))


boostresultsmet %>%
  group_by(Bat_ID) %>%
  summarize(Min_DateTime = min(DateTime), Max_DateTime = max(DateTime))


boostresultsmet$time_aligned <- boostresultsmet$DateTime


boostresultsmet$time_aligned[boostresultsmet$Group == "2"] <- boostresultsmet$time_aligned[boostresultsmet$Group == "2"] - lubridate::days(2)

boostresultsmet$time_aligned[boostresultsmet$Group == "3"] <- boostresultsmet$time_aligned[boostresultsmet$Group == "3"] - lubridate::days(11)

boostresultsmet$shift_time_num <- boostresultsmet$time_aligned

boostresultsmet$shift_time_num <- as.numeric(boostresultsmet$shift_time_num)

str(boostresultsmet)

boostresultsmet |> 
  group_by(Bat_ID) |> 
  summarize(Min_date = min(DateTime), Min_shift = min(time_aligned),
            min_daten =min(NumberDate, Min_shiftn = min(shift_time_num)))

boosted <- boostresultsmet %>%
  select(id = Bat_ID, t = shift_time_num, Temperature, )
head(boosted)

boosted %>%
  group_by(id) %>%
  summarize(Min_DateTime = min(t), Max_DateTime = max(t))

```

I shifted the data so now data all shows d21 on June 5th, the same day as D21 for Group 1. 

## Behavr Analysis

```{r}

BoostBatNames <- unique(boosted$id)

Boost_met <- clean_met %>%
  rename(id = Bat_ID) %>%
  filter(id %in% BoostBatNames)


Boost_data <- Boost_met %>%
  select(id) %>%
  left_join(boosted, by = join_by(id))
str(Boost_data)

boost_met2 <- data.table::data.table(Boost_met, key = "id")
boost_data2 <- data.table::data.table(Boost_data, key = "id")

boost_behavr <- behavr(boost_data2, boost_met2)
summary(boost_behavr)

```

```{r}
ggetho(boost_behavr, aes(x=t, y=id, z=Temperature)) + stat_tile_etho()
```

```{r}
ggetho(boost_behavr, aes(z = Temperature, y = Treatment)) + stat_tile_etho()
```

```{r, warning=FALSE}

ticks_7am <- seq(
  from = floor_date(as.POSIXct(min(boost_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
  to   = ceiling_date(as.POSIXct(max(boost_behavr$t), origin = "1970-01-01", tz = "UTC"), "day") + hours(7),
  by   = "1 day"
)

boost_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
  stat_pop_etho() +
  stat_ld_annotations(
    ld_colours = c( "white", "black"),
    ratio = c(12, 12)* 3600) +
  facet_grid(Treatment ~ .) +
    scale_x_continuous(
    name = "Date/time",
    labels = function(x)
      format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M"))


cols <- c(
  "eVLP alone" = "#F8776D",
  "eVLP+cGAMP" = "#ABA300",
  "eVLP+LPS"    = "#00C19A",
  "eVLP+POLY I:C" = "#00A9FF",
  "SHAM"        = "#ED68ED",
  "TC"          = "#E68613")

boost_behavr |>
  ggetho(aes(x = t, y = Temperature, colour = Treatment)) +
  stat_pop_etho() +
  stat_ld_annotations(
    ld_colours = c("white", "black"),
       # 12h light : 12h dark, in seconds
    offset = 7 * 3600) +           # <<< shift start by 7 hours (7 AM)
  facet_grid(Treatment ~ .) +
    scale_x_continuous(
    name = "Date/time",
    breaks = as.numeric(ticks_7am),
    labels = function(x)
      format(as.POSIXct(x, origin = "1970-01-01", tz = "UTC"), "%Y-%m-%d %H:%M")) +
   scale_colour_manual(name = "Treatment", values = cols) +
  scale_fill_manual(name = "Treatment", values = cols) +
  theme(
    axis.text = element_text(size = 14),       # axis tick labels
    axis.title = element_text(size = 16),      # axis titles
    legend.text = element_text(size = 13),     # legend labels
    legend.title = element_text(size = 15),    # legend title
    strip.text = element_text(size = 12)       # facet labels
  )  -> MOCO_Boost

unique(boost_behavr$Treatment)

ggsave("MOCO_Boost.jpeg", plot = MOCO_Boost, device = jpeg(), width = 16.4, height = 9.5)

```

```{r, warning=FALSE}
#ggetho(bat_behavr, aes(y=Temperature, colour=Treatment), time_wrap = hours(24)) +
#      stat_pop_etho() +
#      stat_ld_annotations() +
#      facet_grid(Treatment ~ .) +
#      scale_y_continuous(name= "Temperature")
```

```{r Sham, warning=FALSE}
bat_met_sham <- bat_met2 %>%
  filter(Treatment == "SHAM")
bat_data_sham <- bat_data2 %>%
  filter(id %in% bat_met_sham$id)
bat_behavr_sham <- behavr(bat_data_sham, bat_met_sham)

bat_behavr_sham %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP, warning=FALSE}
bat_met_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP")
bat_data_eVLP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP$id)
bat_behavr_eVLP <- behavr(bat_data_eVLP, bat_met_eVLP)

bat_behavr_eVLP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_LPS, warning=FALSE}
bat_met_eVLP_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS")
bat_data_eVLP_LPS <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_LPS$id)
bat_behavr_eVLP_LPS <- behavr(bat_data_eVLP_LPS, bat_met_eVLP_LPS)

bat_behavr_eVLP_LPS %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_POLY, warning=FALSE}
bat_met_eVLP_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY")
bat_data_eVLP_POLY <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_POLY$id)
bat_behavr_eVLP_POLY <- behavr(bat_data_eVLP_POLY, bat_met_eVLP_POLY)

bat_behavr_eVLP_POLY %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r eVLP_cGAMP, warning=FALSE}
bat_met_eVLP_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP")
bat_data_eVLP_cGAMP <- bat_data2 %>%
  filter(id %in% bat_met_eVLP_cGAMP$id)
bat_behavr_eVLP_cGAMP <- behavr(bat_data_eVLP_cGAMP, bat_met_eVLP_cGAMP)

bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(y=Temperature, colour=id)) +
      stat_pop_etho() +
      stat_ld_annotations() +
      facet_grid(id ~ .)
```

```{r, warning=FALSE}
ggetho(bat_behavr, aes(y=Temperature, colour=Treatment)) + 
  stat_pop_etho()
```

```{r, warning=FALSE}
bat_behavr_eVLP_cGAMP %>%
  ggetho(aes(x=t, z=Temperature), multiplot = 2) +
      stat_tile_etho() +
      facet_wrap( ~ id)
```

## Subsetting the data to pre-boost and post-boost

Before I proceed with subsetting, I am going to make sure that I understand the data.

```{r}
allbatresults3 %>%
  group_by(batnames) %>%
  summarize(Min_timepoint = min(Timepoint), Min_DateTime = min(DateTime), Max_DateTime = max(DateTime),
            Min_Temp = min(Temperature), Max_Temp = max(Temperature))
```

We can tell that the loggers on 2023-04-16 were all programmed uniformly and collected 42 timepoints before the new start time of midnight on 2023-04-16. 
While the 2023-04-16 loggers have two different programming regimens. Some collected 186 and others collected 330 earlier datapoints. 

Bottom line, there are three different starting timepoints: 43, 187, and 331.
I will substract the appropriate value from the Timepoint column after I separate the data.


```{r}
left_trimmed_results %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint))
allbatresults %>%
  group_by(batnames) %>%
  summarize(minTimepoint = min(Timepoint), maxTimepoint = max(Timepoint), 
            minTemp = min(Temperature), maxTemp = max(Temperature)) %>%
  arrange(maxTimepoint)
```


There are three different batches of treatment, as expected:
2023-03-27 2023-03-29 2023-03-31 were the initial immunization dates (prime)
Day 21 for each batch was, therefore:
2023-04-17 2023-04-19 2023-04-21

```{r}
summary(as.factor(bat_met2$Treatment_date))
first_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-27"))
second_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-29"))
third_treatment <- bat_met2 %>%
  filter(Treatment_date == ymd("2023-03-31"))

first_results <- allbatresults3 %>%
  filter(batnames %in% first_treatment$id)
second_results <- allbatresults3 %>%
  filter(batnames %in% second_treatment$id)
third_results <- allbatresults3 %>%
  filter(batnames %in% third_treatment$id)

first_results_trim <- first_results %>%
  filter(DateTime <= ymd("2023-04-17")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_trim <- second_results %>%
  filter(DateTime <= ymd("2023-04-19")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_trim <- third_results %>%
  filter(DateTime <= ymd("2023-04-21")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

pretreatment_results <- bind_rows(first_results_trim, second_results_trim) %>%
  bind_rows(third_results_trim)

select_bat_results <- pretreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(select_bat_results)
  
bat_data_preboost <- clean_met3 %>%
  select(id) %>%
  left_join(select_bat_results, by = join_by(id))
str(bat_data_preboost)

bat_data_preboost2 <- data.table::data.table(bat_data_preboost, key = "id")

bat_behavr_preboost <- behavr(bat_data_preboost2, bat_met2)
```

Note that I think that the last day shown on these plots is the booster immunization day.

```{r, warning=FALSE}
# summary(bat_behavr_preboost)
# bat_behavr_preboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_preboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

And now for the post-boost results. 
For starters, I will trim the 4 days after the injections.
That didn't work because ggetho kept saying that there was less than 3 days of data :/

Secondly just kept the above format but used more data.

```{r}
# first_results_post <- first_results %>%
#   filter(DateTime >= ymd("2023-04-17")) %>%
#   filter(DateTime <= ymd("2023-04-22")) %>%
#   mutate(Timepoint = Timepoint - 474) ## My attempt to standardize the start time. 
# second_results_post <- second_results %>%
#   filter(DateTime >= ymd("2023-04-19")) %>%
#   filter(DateTime <= ymd("2023-04-24")) %>%
#   mutate(Timepoint = Timepoint - 762)
# third_results_post <- third_results %>%
#   filter(DateTime >= ymd("2023-04-21")) %>%
#   filter(DateTime <= ymd("2023-04-26")) %>%
#   mutate(Timepoint = Timepoint - 906)

first_results_post <- first_results %>%
  filter(DateTime <= ymd("2023-04-22")) %>%
  mutate(Timepoint = Timepoint - 42) ## My attempt to standardize the start time. 
second_results_post <- second_results %>%
  filter(DateTime <= ymd("2023-04-24")) %>%
  mutate(Timepoint = Timepoint - 330) ## My attempt to standardize the start time. 
third_results_post <- third_results %>%
  filter(DateTime <= ymd("2023-04-26")) %>%
  mutate(Timepoint = Timepoint - 186) ## My attempt to standardize the start time. 

posttreatment_results <- bind_rows(first_results_post, second_results_post) %>%
  bind_rows(third_results_post)

post_bat_results <- posttreatment_results %>%
  mutate(Timepoint = Timepoint*10*60) %>%
  select(id = batnames, t = Timepoint, Temperature) %>%
  filter(id != "3N8245")
head(post_bat_results)
  
bat_data_postboost <- clean_met3 %>%
  select(id) %>%
  left_join(post_bat_results, by = join_by(id))
str(bat_data_postboost)

bat_data_postboost2 <- data.table::data.table(bat_data_postboost, key = "id")

bat_behavr_postboost <- behavr(bat_data_postboost2, bat_met2)
summary(bat_behavr_postboost)
```
```{r Postboost_Summary, warning=FALSE}
# summary(bat_behavr_postboost)
# bat_behavr_postboost %>%
#   ggetho(aes(y=Temperature, colour=Treatment)) +
#       stat_pop_etho() +
#       stat_ld_annotations() +
#       facet_grid(Treatment ~ .)
# ggetho(bat_behavr_postboost, aes(y=Temperature, colour=Treatment)) + 
#   stat_pop_etho()
```

So I think that the booster happened on day 4 on these graphs, but I need to verify.

## Booster Analysis

To be certain of using the correct timestamps, I am going to analyze with the openair package, as I did for the squirrels.

This is done one bat at a time...

https://bookdown.org/david_carslaw/openair/sections/trend-analysis/time-variation.html

First, comparing all pre-boost data to (boost day + 1 day following)

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8476" # A SHAM bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

The results look very similar.


```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + BoostWindow) %>%
  mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

Now let's look at comparing (boost day + 1 day following) to all other days

```{r}
mybatID <- "1N8475" # A eVLP cGAMP bat
singlebatresults <- allbatresults3 %>%
  filter(batnames == mybatID)

Immunization <- bat_met2 %>%
  filter(id == mybatID) %>%
  select(Treatment_date) %>%
  pull
Boost <- Immunization + days(21)
BoostWindow <- days(2)

singlebatresults2 <- singlebatresults %>%
  filter(DateTime <= Boost + days(BoostWindow)) %>%
  mutate(Boosted = if_else((DateTime >= Boost ) & (DateTime <= (Boost + BoostWindow)), "Postboost", "Other")) %>%
  mutate(Boosted = factor(Boosted, levels = c("Other", "Postboost"))) %>%
  rename(date = DateTime)

newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", group = "Boosted",  xlab = "Hour",
              difference = TRUE)
plot(newplot, subset="hour")
```

To keep the comparison cleaner, I will continue with pre versus post. 
The only downside is that the confidence interval is a little wider on the "pre" data than the "other" data.

## Comparison of all bats

```{r}
bat_met2 %>%
  arrange(Treatment) %>%
  select(id, Treatment)
```

Let's make a function for the timeVariation plot

```{r}
timeVariation_fun <- function(arg1) {
  mybatID <- arg1

  
  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(2)

  singlebatresults <- allbatresults3 %>%
    filter(batnames == mybatID)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant ="Temperature", 
                           group = "Boosted",  
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

```{r}
timeVariation_fun("1N8476")
```

Now let's look at the groups one at a time.

```{r SHAM}
bat_sham <- bat_met2 %>%
  filter(Treatment == "SHAM") %>%
  select(id) %>%
  pull
lapply(bat_sham, timeVariation_fun)
```

```{r eVLP only}
bat_eVLP <- bat_met2 %>%
  filter(Treatment == "eVLP") %>%
  select(id) %>%
  pull
lapply(bat_eVLP, timeVariation_fun)
```


```{r eVLP LPS}
bat_LPS <- bat_met2 %>%
  filter(Treatment == "eVLP LPS") %>%
  select(id) %>%
  pull
lapply(bat_LPS, timeVariation_fun)
```

```{r eVLP POLY}
bat_POLY <- bat_met2 %>%
  filter(Treatment == "eVLP POLY") %>%
  select(id) %>%
  pull
lapply(bat_POLY, timeVariation_fun)
```

```{r eVLP cGAMP}
bat_cGAMP <- bat_met2 %>%
  filter(Treatment == "eVLP cGAMP") %>%
  select(id) %>%
  pull
lapply(bat_cGAMP, timeVariation_fun)
```

## Tent Logger Data

Two possible ways to incorporate this data. 
First, I will add it as a separate "pollutant".
Second, I will show it as a separate "feature".

```{r}
alltentresults <- fread("TemperatureLoggers/tentloggerdata_long.tsv")
alltentresults <- alltentresults %>%
  mutate(Tent = gsub("Tent ", "", Tent))
str(alltentresults)

tent_met <- Metadata %>%
  drop_na(Filename) %>%
  select(Filename, DMR_ID, Tag, Treatment, Age, Sex,
         Attachment_date = "Date Attached", 
         Treatment_date = Day_00, 
         RIP_date,
         Tent = Tent.x) %>%
  mutate(id = gsub(".DAT", "", Filename)) %>%
  relocate(id)

tentbatresults <- tent_met %>%
  select(id, Tent) %>%
  right_join(allbatresults3, by = join_by(id == batnames))

tentbatresults2 <- tentbatresults %>%
  left_join(alltentresults, by = join_by(Tent, DateTime)) %>%
  rename(`Bat T` = Temperature.x) %>%
  rename(`Tent T` = Temperature.y)

write_csv(tentbatresults2, file = "allbatandtentresults.csv")
```

A new plot function:

```{r}
timeVariation_fun2 <- function(arg1) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  # Immunization <- bat_met2 %>%
  #   filter(id == mybatID) %>%
  #   select(Treatment_date) %>%
  #   pull
  # Boost <- Immunization + days(21)
  # BoostWindow <- days(2)

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID)

  # singlebatresults2 <- singlebatresults %>%
  #   filter(DateTime <= Boost + BoostWindow) %>%
  #   mutate(Boosted = if_else((DateTime >= Boost ), "Postboost", "Preboost")) %>%
  #   mutate(Boosted = factor(Boosted, levels = c("Preboost", "Postboost"))) %>%
  #   rename(date = DateTime)
  
  singlebatresults2 <- singlebatresults %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant =c("Tent T", "Bat T"), 
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun2("1N8476")
```
That looks good. Let's plot each group at a time.

```{r Sham-Tent}
lapply(bat_sham, timeVariation_fun2)
```

```{r eVLP-Tent}
lapply(bat_eVLP, timeVariation_fun2)
```

```{r LPS-Tent}
lapply(bat_LPS, timeVariation_fun2)
```

```{r POLY-Tent}
lapply(bat_POLY, timeVariation_fun2)
```

```{r cGAMP-Tent}
lapply(bat_cGAMP, timeVariation_fun2)
```

Okay, because the "difference" function worked fine, I don't think that I need to reorganize the data. 
Let's see if I can also use the boost "group" at the same time.

Dammit: "Can only have one pollutant with a grouping variable, or several pollutants and no grouping variable."

Okay, I think I will make a new variable called delta T and use that as the only pollutant.

I will also test different "postboost" windows.

A new plot function:

```{r}
timeVariation_fun3 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime)

  newplot <- timeVariation(singlebatresults2, pollutant = "delta T", 
                           group = "Boosted",
                           xlab = c(paste(mybatID, Group), paste(mybatID, Group), "Month", "Weekday"),
                           difference = TRUE, plot = FALSE)
  plot(newplot, subset="day.hour")
  plot(newplot, subset="hour")
}
```

Test with one bat:

```{r}
timeVariation_fun3("1N8476", 5)
```

```{r Sham-deltaT}
lapply(bat_sham, timeVariation_fun3, 3)
```

```{r eVLP-deltaT}
lapply(bat_eVLP, timeVariation_fun3, 3)
```

```{r LPS-deltaT}
lapply(bat_LPS, timeVariation_fun3, 3)
```

```{r POLY-deltaT}
lapply(bat_POLY, timeVariation_fun3, 3)
```

```{r cGAMP-deltaT}
lapply(bat_cGAMP, timeVariation_fun3, 3)
```

## Comparing by Treatment Group

Using the "group" parameter in timeVariation and to make separate plots for pre-boost and then each day post-boost.


```{r Group Comparisons}
tentbatresults3 <- tentbatresults2 %>%
  mutate(`delta T` = `Bat T` - `Tent T`) %>%
  rename(date = DateTime)

bat_boost_date <- bat_met2 %>%
  select(id, Treatment, Treatment_date) %>%
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP",
                                                  "eVLP LPS", "eVLP POLY", "eVLP cGAMP")))

tentbatresults4 <- tentbatresults3 %>%
  left_join(bat_boost_date, by = join_by(id))

str(tentbatresults4)

preboost_tentbatresults <- tentbatresults4 %>%
  filter(date < (Treatment_date + days(21)))

boostday_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(21)) & date <= (Treatment_date + days(22)))

postboostday1_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(22)) & date <= (Treatment_date + days(23)))

postboostday2_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(23)) & date <= (Treatment_date + days(24)))

postboostday3_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday4_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(24)) & date <= (Treatment_date + days(25)))

postboostday5_tentbatresults <- tentbatresults4 %>%
  filter(date >= (Treatment_date + days(25)) & date <= (Treatment_date + days(26)))


preboost_plot <- timeVariation(preboost_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Preboost", "Preboost", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(preboost_plot, subset="hour")

boostday_plot <- timeVariation(boostday_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Boost Day", "Boost Day", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(boostday_plot, subset="hour")

postboostday1_plot <- timeVariation(postboostday1_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 1", "Postboost Day 1", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday1_plot, subset="hour")

postboostday2_plot <- timeVariation(postboostday2_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 2", "Postboost Day 2", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday2_plot, subset="hour")

postboostday3_plot <- timeVariation(postboostday3_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 3", "Postboost Day 3", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday3_plot, subset="hour")

postboostday4_plot <- timeVariation(postboostday4_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 4", "Postboost Day 4", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday4_plot, subset="hour")

postboostday5_plot <- timeVariation(postboostday5_tentbatresults, pollutant = "delta T", 
                               group = "Treatment",
                               xlab = c("Postboost Day 5", "Postboost Day 5", "Month", "Weekday"),
                               difference = FALSE, plot = FALSE, 
                               key.columns = 3)
plot(postboostday5_plot, subset="hour")

```


## Quantifying Temperature Differences

Designing a function to measure the daily difference in Bat temperature compared to the tent temperature.

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
```


A function to quantify this delta T by day:

```{r}
tempDifference_fun <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  DMR_Id <- NewTempLoggerMetadata %>%
    filter(Id == mybatID) %>%
    select(DMR_ID) %>%
    pull
  
  Tent <- Metadata %>%
    filter(DMR_ID == DMR_Id) %>%
    select(Tent.x) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`)/144, 
                      daily_batT = mean(`Bat T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 Tent = Tent,
                                 `DeltaT d-3` = daily_summary$daily_delta[1],
                                 `DeltaT d-2` = daily_summary$daily_delta[2],
                                 `DeltaT d-1` = daily_summary$daily_delta[3],
                                 `DeltaT d0` = daily_summary$daily_delta[4],
                                 `DeltaT d1` = daily_summary$daily_delta[5],
                                 `DeltaT d2` = daily_summary$daily_delta[6],
                                 `BatT d-3` = daily_summary$daily_batT[1],
                                 `BatT d-2` = daily_summary$daily_batT[2],
                                 `BatT d-1` = daily_summary$daily_batT[3],
                                 `BatT d0` = daily_summary$daily_batT[4],
                                 `BatT d1` = daily_summary$daily_batT[5],
                                 `BatT d2` = daily_summary$daily_batT[6])
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
tempDifference_fun("1N8476", 3)
bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
bat_delta <- bind_rows(bat_delta_list)
bat_delta_full <- bat_delta |>
  mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
  rowwise() |>
  mutate(`Preboost Mean dT` = mean(c(`DeltaT d-3`,`DeltaT d-2`,`DeltaT d-1`))) |>
  mutate(`Postboost Mean dT` = mean(c(`DeltaT d0`,`DeltaT d1`,`DeltaT d2`))) |>
  mutate(`Preboost Bat T` = mean (c(`BatT d-3`,`BatT d-2`,`BatT d-1`))) |>
  mutate(`Postboost Bat T` = mean(c(`BatT d0`,`BatT d1`,`BatT d2`)))
bat_delta_full
```

And let's try to look at the various possible models.

```{r Delta T Models}
first_model <- glm(`Postboost Mean dT` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Mean dT` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Mean dT` ~ `Preboost Mean dT` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

```{r Delta T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Mean dT`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day Postboost ',Delta,'T'))) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)
bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Mean dT`:`Postboost Mean dT`, names_to = "Boost")
ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab(expression(paste('3-day ',Delta,'T')))

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Mean dT`, y=`Postboost Mean dT`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


Well that is quite disappointing. 
Tent is clearly the biggest factor driving this.
Delta T is not the best way to analyze this phenomenon.

First, we need to look at Bat T instead of delta T.


```{r Bat T Models}
first_model <- glm(`Postboost Bat T` ~ Treatment, data = bat_delta_full)
summary(first_model)
second_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(second_model)
Day0_model <- glm(`DeltaT d0` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day0_model)
Day1_model <- glm(`DeltaT d1` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day1_model)
Day2_model <- glm(`DeltaT d2` ~ `Preboost Bat T` + Treatment, data = bat_delta_full)
summary(Day2_model)
bigger_model <- glm(`Postboost Bat T` ~ `Preboost Bat T` + Tent + Age + Sex + Treatment, data = bat_delta_full)
summary(bigger_model)
```

This is the ideal situation for a mixed effects model!

Let's see how it works.

Following guidance at https://m-clark.github.io/mixed-models-with-R/random_intercepts.html

```{r}
bat_delta_T <- bat_delta_full |>
  mutate(`Boost dT` = `Postboost Bat T` - `Preboost Bat T`)

mixed_model <- lmer(`Boost dT` ~ Treatment + (1|Tent), data = bat_delta_T)
summary(mixed_model)
confint(mixed_model)
ranef(mixed_model)$Tent
coef(mixed_model)$Tent

predictInterval(mixed_model)   # for various model predictions, possibly with new data

REsim(mixed_model)             # mean, median and sd of the random effect estimates

plotREsim(REsim(mixed_model))  # plot the interval estimates

```

It looks like even after incorporating Tent as a random effect, the Treatment effect is not significant.
This may be because I am looking at the 3-day average and I need a more focused approach.

```{r Bat T Graphs}
stat_summary <- bat_delta_full |> 
  group_by(Treatment) |>
  tally()
stat_summary
ggplot(bat_delta_full, aes(x = Treatment, y = `Postboost Bat T`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme_cowplot() +
  theme(legend.position="none") +
  theme(axis.title.x = element_blank()) +
  geom_text(data = stat_summary, aes(Treatment, Inf, label = paste0("n=",n)), vjust = 1)

bat_delta_pivot <- bat_delta_full |>
  pivot_longer(cols = `Preboost Bat T`:`Postboost Bat T`, names_to = "Boost") |>
  mutate(Boost = factor(Boost, levels = c("Preboost Bat T", "Postboost Bat T")))

ggplot(bat_delta_pivot, aes(x = Treatment, y = value, fill = Boost)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5,  position = position_jitterdodge()) +
  theme(axis.title.x = element_blank()) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_pivot, aes(x = Boost, y = value)) +
  geom_boxplot(outlier.shape = NA) +
  theme_cowplot() +
  geom_point(aes(fill = Boost), size = 2, shape = 21, alpha = 0.5, position = position_nudge()) +
  geom_line(aes(group=BatID), colour="red", linetype="11") +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Treatment) +
  ylab("3-day Bat T Mean")

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, color=Treatment)) + 
    geom_point() +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Age)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Tent, shape = Treatment)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_full, aes(x=`Preboost Bat T`, y=`Postboost Bat T`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

```


### Day 1 Post-boost

Reviewing some of the data from the sham group, I think it would be better to look only at post-boost day 1 or day 2.
This should avoid some of the impact of handling. 

```{r}
bat_delta_T_day1 <- bat_delta_full |>
  mutate(`Boost dT day1` = `BatT d1` - `Preboost Bat T`) |>
  mutate(`Boost dT day2` = `BatT d2` - `Preboost Bat T`)

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d1`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

ggplot(bat_delta_T_day1, aes(x=`Preboost Bat T`, y=`BatT d2`, 
                           color=Treatment, shape = Sex)) + 
    geom_point() +
    geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "red") +
    theme_cowplot()

mixed_model_d1 <- lmer(`Boost dT day1` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d1)
confint(mixed_model_d1)
ranef(mixed_model_d1)$Tent
coef(mixed_model_d1)$Tent
predictInterval(mixed_model_d1)   # for various model predictions, possibly with new data
REsim(mixed_model_d1)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d1))  # plot the interval estimates

mixed_model_d2 <- lmer(`Boost dT day2` ~ Treatment + (1|Tent), data = bat_delta_T_day1)
summary(mixed_model_d2)
confint(mixed_model_d2)
ranef(mixed_model_d2)$Tent
coef(mixed_model_d2)$Tent
predictInterval(mixed_model_d2)   # for various model predictions, possibly with new data
REsim(mixed_model_d2)             # mean, median and sd of the random effect estimates
plotREsim(REsim(mixed_model_d2))  # plot the interval estimates
```

The d2 model might show something interesting for the Poly(I:C) group, but it is a small effect.


## Hourly Temperature Differences

```{r}
mybatID <- "1N8476" # A SHAM bat

Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull

Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
Boost <- Immunization + days(21)
BoostWindow <- days(3)
BoostText <- paste0("Postboost(", as.character(3), ")")

singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
singlebatresults3 <- singlebatresults2 %>%
  group_by(Boosted) %>%
  summarise_by_time(.date_var = date, 
                    .by = "hour", 
                    hourly_delta = sum(`delta T`)/6)


ggplot(singlebatresults3, aes(x = date, y = hourly_delta, color = Boosted)) + 
  geom_line() +
  theme_cowplot()
```
A function to quantify this delta T by hour:

```{r}
tempDifference_fun2 <- function(arg1, arg2) {
  mybatID <- arg1

  Group <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment) %>%
    pull
  
  Sex <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Sex) %>%
    pull
  
  Age <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Age) %>%
    pull
  
  Immunization <- bat_met2 %>%
    filter(id == mybatID) %>%
    select(Treatment_date) %>%
    pull
  Boost <- Immunization + days(21)
  BoostWindow <- days(arg2)
  BoostText <- paste0("Postboost(", as.character(arg2), ")")

  singlebatresults <- tentbatresults2 %>%
    filter(id == mybatID) %>%
    mutate(`delta T` = `Bat T` - `Tent T`)

  singlebatresults2 <- singlebatresults %>%
    filter(DateTime <= Boost + BoostWindow) %>%
    mutate(Boosted = if_else((DateTime >= Boost ), BoostText, "Preboost")) %>%
    mutate(Boosted = factor(Boosted, levels = c("Preboost", BoostText))) %>%
    rename(date = DateTime) %>%
    filter(row_number() <= n()-1)
  
  daily_summary <- singlebatresults2 %>%
    group_by(Boosted) %>%
    summarise_by_time(.date_var = date, 
                      .by = "day",
                      daily_delta = sum(`delta T`))
  
  daily_summary_tibble <- tibble(BatID = mybatID, 
                                 Treatment = Group,
                                 Age = Age,
                                 Sex = Sex,
                                 `Preboost d-3` = daily_summary$daily_delta[1]/144,
                                 `Preboost d-2` = daily_summary$daily_delta[2]/144,
                                 `Preboost d-1` = daily_summary$daily_delta[3]/144,
                                 `Boost d0` = daily_summary$daily_delta[4]/144,
                                 `Boost d1` = daily_summary$daily_delta[5]/144,
                                 `Boost d2` = daily_summary$daily_delta[6]/144)
  
  return(daily_summary_tibble)
}
```

Now let's try it!

```{r}
# tempDifference_fun2("1N8476", 3)
# bat_delta_list <- lapply(batnames, tempDifference_fun, 3)
# bat_delta <- bind_rows(bat_delta_list)
# bat_delta_full <- bat_delta |>
#   mutate(Treatment = factor(Treatment, levels = c("SHAM", "eVLP", "eVLP LPS", "eVLP POLY", "eVLP cGAMP"))) |>
#   rowwise() |>
#   mutate(`Preboost Mean dT` = mean(c(`Preboost d-3`,`Preboost d-2`,`Preboost d-1`))) |>
#   mutate(`Postboost Mean dT` = mean(c(`Boost d0`,`Boost d1`,`Boost d2`)))
# bat_delta_full
```
